### 2020-03-05 ManuscriptTime.tex
Initial commit

\documentclass[
%     paper,
    manuscript,
%     twocolumn,
%     twoside,
%     revised,
  ]{geophysics}

% Remove once done
\usepackage[textsize=footnotesize, textwidth=2.8cm]{todonotes}

% Additional packages to geophysics.cls
\usepackage[UKenglish]{babel}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[T1]{fontenc}

\usepackage{amssymb, amsmath, amsfonts}
\usepackage{upquote}
\usepackage[strings]{underscore}
\usepackage{siunitx}                % SI conform commands (eg \num)
\usepackage{tabularx}
\usepackage{colortbl}
\usepackage{booktabs}
\usepackage{xspace}

\usepackage[pdftex, final]{hyperref}
% \usepackage[pdftex, hidelinks]{hyperref}
\hypersetup{allcolors=blue, allbordercolors={0 0 .5}, colorlinks=true}

% Figures
\DeclareGraphicsExtensions{.pdf,.png,.jpg}
\renewcommand{\figdir}{./figures}
\ifthenelse{\boolean{@twoc}}{%
  \newcommand{\cwidth}{250pt}}{%
  \newcommand{\cwidth}{240pt}}

\ifthenelse{\boolean{@twoc}}{%
  \newcommand{\acwidth}{250pt}%
  \newcommand{\tcwidth}{250pt}}{%
  \newcommand{\acwidth}{240pt}%
  \newcommand{\tcwidth}{312pt}}

% Own commands
\newcommand{\mr}[1]{\mathrm{#1}}
\newcommand{\emg}[2]{\texttt{emg#1#2}\xspace}
\newcommand{\empymod}{\texttt{empymod}\xspace}
\newcommand{\simpeg}{\texttt{SimPEG}\xspace}

% Remove once done
\newcommand{\itodo}[1]{\todo[color=blue!40!white, inline]{\sffamily #1}}
\newcommand{\mtodo}[1]{\todo[inline]{\sffamily #1}}
\newcommand{\etodo}[1]{\todo[color=blue!40!white]{#1}}
\newcommand{\dul}[1]{\underline{\underline{#1}}}
\newcommand{\ul}[1]{\underline{#1}}
\newcommand{\rmk}[1]{{\color{red}{#1}}}

\begin{document}

\title{Fast time-domain electromagnetic modelling in the frequency domain}

\renewcommand{\thefootnote}{\fnsymbol{footnote}}

\ms{}  % ARTICLE-ID

\address{
\footnotemark[1]TU Delft,
Building 23,
Stevinweg 1 / PO-box 5048,
2628 CN Delft;\\
\footnotemark[2]Shell Global Solutions International BV,
Grasweg 31, 1031 HW Amsterdam;\\
E-mail: \href{mailto:Dieter@Werthmuller.org}{Dieter@Werthmuller.org};\\
\textbf{Keywords}: CSEM, Fourier transforms, 3D modelling.
}

\author{%
Dieter Werthmüller\footnotemark[1], %            orcid: 0000-0002-8575-2484
Wim A.\ Mulder\footnotemark[1]\footnotemark[2], % orcid: 0000-0001-7020-9297
and Evert C.\ Slob\footnotemark[1] %             orcid: 0000-0002-4529-1134
}

\footer{}
\lefthead{Werthmüller et al.}
\righthead{Time-domain CSEM}

\maketitle

%%fakesection ===    ABSTRACT    ===
\begin{abstract}
%
Modelling three-dimensional time-domain controlled-source electromagnetic data
with frequency-domain solvers is an alternative to time-domain solvers. It
requires the computation of many frequencies if standard Fourier transforms are
used, which can make it prohibitively expensive in comparison with a
time-domain computation. The speed of time-domain modelling with
frequency-domain computation is defined by three key points: solver, method and
implementation of the Fourier transformation, and gridding.
%
The faster the frequency-domain solver, the faster time-domain modelling will
be. It is important that the solver is robust over a wide range of frequencies.
A three-dimensional, iterative, frequency-domain multigrid solver satisfies
this requirement, but any other robust solver will do.
%
The method should require as few frequencies as possible while remaining
robust. As the frequency range spans many orders of magnitude, the required
frequencies are ideally equally spaced on a logarithmic scale, as is the case
for, e.g., the digital linear filter method or the logarithmic fast Fourier
transform. With careful selection of frequencies and interpolation, 15 to 20
frequencies are enough to cover a wide range of offsets in one-dimensional test
models.
%
The gridding should be frequency-dependent, which is generally accomplished by
making it a function of skin depth. Optimising for the least number of required
cells for a given frequency should be combined with optimising for
computational speed.
%
Looking carefully at these three points results in much faster computation
times with speedup-factors of ten and more over previous results. A
frequency-domain computation of time-domain data can therefore be competitive
to time-domain solvers if the required frequencies and their corresponding
grids are carefully chosen.
%
\end{abstract}

\section{Introduction}

The controlled-source electromagnetic (CSEM) method is one of the common
non-seismic tools in exploration geophysics, not just in hydrocarbon
exploration \citep{GEO.10.Constable}, but also in the search for sulfides
\citep{GRL.19.Gehrmann}, water \citep{GEO.05.Pedersen}, or geothermal sources
\citep{WGC.15.Girard}. While in the deep marine environment current sources
with few frequencies are used, transient measurements are more common in the
shallow marine environment and on land. One of the main reasons is the
dominance of the airwave in shallow marine and terrestrial measurements, which
can be better separated in the time domain.

Acquired CSEM data are often analysed (modelled and inverted for) in their
respective domain, either the frequency or time domain. Modellers of layered
media usually exploit the horizontal shift-invariance by computing the
responses in the wavenumber-frequency domain followed by a 2D inverse spatial
Fourier transform, also called Hankel transform, to the space-frequency domain,
and a regular inverse Fourier transform if time-domain data are required
\citep[e.g., ][]{GEO.15.Hunziker}.

CSEM codes for 2D and 3D computations, on the other hand, mostly compute their
responses directly in the required domain, either frequency or time. However,
under certain conditions it can be competitive to model time-domain data with a
frequency-domain code, as shown by \cite{GEO.08.Mulder}. The conditions are: a
sufficiently powerful solver, appropriate frequency selection and
interpolation, and an automated gridding, for which they used the
multi-frequency CSEM approach presented by \cite{GEO.07.Plessix}. We build upon
these results but improve the run time from hours to minutes. The main reasons
for this significant speed-up are an adaptive, frequency-dependent gridding
scheme that minimizes the required cells in each dimension, and a logarithmic
Fourier transform such as digital linear filters \citep[DLF, ][]{GP.71.Ghosh}
or the logarithmic fast Fourier transform \citep[FFTLog, ][]{RAS.00.Hamilton}
to go from the frequency to the time domain. The latter makes it also possible
to only use the imaginary part of the frequency-domain response, which has
advantages when it comes to interpolation.

In the next section, we briefly review the methods and highlight their
advantages and shortcomings. This is followed by an outline of our changes to
the Fourier transform and the adaptive gridding. Finally, we demonstrate the
efficiency of the approach with some numerical results.

\section{Motivation}

Time-domain modelling with a frequency-domain code requires the computation of
a relatively wide range of frequencies for the Fourier transform. However, the
required computation grids for low frequencies (in our case in the order of
0.001\,Hz) and high frequencies (around 100\,Hz) are hugely different. Low
frequencies can be computed on a coarser grid, but they require a much larger
domain in order to avoid boundary effects. High frequencies, on the other hand,
require denser gridding, but they are much more limited in reach. An adaptive
gridding scheme is therefore required. We build our approach upon
\cite{GEO.07.Plessix}, who presented such an adaptive gridding for
multi-frequency (and multi-source) CSEM modelling.

An adaptive grid in the frequency domain is naturally based on the skin depth,
which is the distance after with the amplitude of the electromagnetic field is
decayed by $1/e\approx 37\,\%$. The skin depth $\delta$ is a function of
conductivity and frequency, and for the diffusive approximation of Maxwell's
equation in an isotropic, homogeneous medium is given by \citep[e.g.,][
equation~1.53]{B.SEG.88.Ward}
%
\begin{equation}
  \delta = \sqrt{\frac{2}{\omega\mu\sigma}}
  \ \approx \
  % \frac{503.3}{\sqrt{f\sigma}} \, ,
  % 503.3(f\sigma)^{-1/2} \, ,
  503.3/\sqrt{f\sigma} \, ,
  \label{eq:skindepth}
\end{equation}
%
where $\sigma$ is conductivity (S/m), $\omega=2\pi f$ is angular frequency of
frequency $f$ (Hz), and $\mu$ is magnetic permeability (H/m). The approximation
is obtained by using the free-space value of magnetic permeability,
$\mu_0=4\pi\times10^{-7}\,$H/m.

\cite{GEO.07.Plessix} define the minimum cell width $\Delta_\mathrm{min}$ as a
fraction of the minimum skin depth $\delta_\mathrm{min}$, where
$\delta_\mathrm{min}$ should be $2-3$ times bigger than $\Delta_\mathrm{min}$.
The smallest cells have to be around the source; in the marine case the minimum
skin depth is therefore defined by the conductivity of seawater. However, this
yields quite large cells for low frequencies, so special care has to be taken
around the source. The grid dimension, on the other hand, is defined as a
function of the skin depth for the average conductivity of the background,
$\delta_\mathrm{ave}$. They use four times $\delta_\mathrm{ave}$ for the $x$-,
$y$-, and downward $z$-directions, and a fixed 50\,km for the upward
$z$-direction to account for the airwave. To reduce the number of cells, it is
desirable to introduce stretching, at least in the boundary zone outside of the
area where source and receivers are located.

\cite{GEO.08.Mulder} provide a computational complexity analysis of various
methods to model transient electromagnetic responses directly in the
time-domain, and compare it to the computation of transient EM responses in the
frequency-domain with a subsequent Fourier transform. They conclude the review
by stating \emph{\guillemotleft Although it remains to be seen which of the
four methods requires the least computer time for a given accuracy, the
frequency-domain approach appears to be attractive.\guillemotright} Their
approach is to minimize the computation time by having, in addition to the just
introduced adaptive gridding, an adaptive frequency selection scheme. This
scheme starts with computing the responses for a set of just five frequencies,
regularly sampled on a log-scale, from minimum to maximum required frequency.
All the other frequencies are interpolated with a shape-preserving
piecewise-cubic Hermite interpolation \citep[PCHIP, ][]{SIAM.80.Fritsch}.
Testing the stability of the obtained response by removing a single
frequency-value at a time their scheme decides if more frequencies in-between
the already computed ones are required. In this way frequencies are only added
if required, hence if certain criteria of response stability are not met. While
this method is good and effective for a single offset, it looses all its
advantages if one tries to compute different offsets within one computation, as
each offset requires a different set of adaptive frequencies. Additionally, it
hampers the parallelization over frequencies.

We present improvements to both the adaptive gridding and the transform from
frequency domain to time domain, which makes time-domain modelling with a
frequency-domain code even more competitive in comparison with time-domain
codes.

\section{Methodology}

For the numerical computations we use the open-source (Apache License 2.0) code
\emg3d \citep{JOSS.19.Werthmuller}, a multigrid solver based on
\cite{GP.06.Mulder}, which can be used as a preconditioner for Krylov subspace
solvers or as a solver on its own. The multigrid approach works fine for the
diffusive approximation of Maxwell's equation, which assumes that
$\omega\varepsilon \ll \sigma$, where $\varepsilon$ is electric permittivity
(F/m). The remaining system to solve in the frequency domain is then given by
the second-order differential equation of the electric field,
%
\begin{equation}
    \mathrm{i}\omega\sigma \mathbf{E} +
    \nabla \times \mu^{-1} \nabla \times \mathbf{E}
    = -\mathrm{i}\omega\mathbf{J}_\mathrm{s} \, ,
  \label{eq:maxwell}
\end{equation}
%
where $\mathbf{E}$ is the electric field (V/m) and $\mathbf{J}_\mathrm{s}$ the
current source (A/m$^2$); time convention is $\exp(\mathrm{i}\omega t)$. The
standard multigrid approach fails for severe stretching or strong anisotropy,
for which \emg3d has implemented known improvements such as line-relaxation and
semicoarsening \citep{ECCFD.06.Jonsthovel}, with a non-standard Cholesky
decomposition to speed up the computation \citep{GEO.08.Mulder}. One of the big
advantages of the multigrid method is that it scales linearly (optimal) with
the grid size in both CPU and RAM usage \citep{B.Springer.20.Mulder}. This
makes it feasible to run even big models on standard computers, without the
need for big clusters. All the examples in this article are run on a laptop
with an i7-6600U\,CPU\,@\,2.6\,GHz (x4) and 16\,GB of memory, using Ubuntu
18.04 and Python 3.7. For the semi-analytical computations of layered models we
use the open-source (Apache License 2.0) code \empymod
\citep{GEO.17.Werthmuller}.

\section{Frequency selection}

An important factor in terms of speed and accuracy for time-domain responses
obtained from frequency-domain computations is the selection of the required
frequencies. The fewer frequencies required, the faster we obtain the
time-domain result. We decided to use a regular spacing of frequencies on a
log-scale, rather than an adaptive scheme. This approach is favourable if a
wide range of offsets is needed, as the required frequencies change with offset
and an adaptive frequency selection is therefore often tailored to a single
offset. Also, it allows for straightforward parallelization over frequencies,
which is not completely possible with an adaptive scheme.

For the actual transform we use either the digital linear filter (DLF) method
or the logarithmic fast Fourier transform (FFTLog). The DLF method was
introduced to geophysics by \cite{GP.71.Ghosh}, and is arguably the most common
method in EM geophysics for both its simplicity and its speed. It is
implemented for the Hankel and Fourier transforms in most EM modelling codes,
e.g., \cite{GEO.09.Key}. A simple tool to design digital linear filters was
recently presented by \cite{GEO.19.Werthmuller}, together with a comprehensive
overview of the history and development of DLF in geophysics. FFTLog,
introduced by \cite{RAS.00.Hamilton}, is another transform algorithm which
proved to be powerful for the frequency-to-time domain transformation of EM
responses, e.g., \cite{INT.14.Werthmuller}. In our tests they are both about
equal in speed and accuracy. DLF requires a wider range and many more
frequencies than the FFTLog. Both methods share some important characteristics
in comparison with the standard FFT: the required input frequencies are equally
spaced on a logarithmic scale, and they only require either the real or the
imaginary part of the frequency-domain response. We can take advantage of that
by using only the imaginary part of the frequency-domain response. The
imaginary part has the characteristics that its value goes to zero when the
frequency goes to zero or to infinity, with the advantage that knowing the
endpoints makes it possible to convert the extrapolation of missing frequencies
into interpolation.

There are three key-parameters that have to be defined: Minimum and maximum
frequency ($f_\mr{min}, f_\mr{max}$) and the number of frequencies per decade
($n_\mr{dec}^f$). Trial-and-error with a 3D code is very time-intensive.
However, a simplified, layered model for the required survey setup and a fast
1D modeller makes it possible to estimate these parameters easily. To this end,
we created a graphical, interactive tool using \empymod, as shown in
Figure~\ref{fig:GUI-FFTLog-lin}.
%
\plot{GUI-FFTLog-lin}{width=\tcwidth}{
  Interactive frequency selection for a user-provided layered model (the shown
  model parameters are described in the text). This example shows the impulse
  response at an offset of 5\,km, for which it uses FFTlog from 0.001\,Hz to
  10\,Hz with five frequencies per decade. Note that the error is clipped for
  values smaller than 0.01\,\% and bigger than 100\,\%.
}
%
The example model is a marine scenario with 1\,km water depth of resistivity
$\rho=0.3\,\Omega$\,m ($\rho = \sigma^{-1}$) and a 100\,m thick target of
100\,$\Omega$\,m 1\,km below the seafloor in a background of 1\,$\Omega$\,m.
The source is 50\,m above the seafloor, the receiver is on the seafloor, and
the response is the inline $x$-directed $E$-field. The left subplot shows the
imaginary part in the frequency domain and the right plot the corresponding
impulse response in the time domain. The red lines are the precise results
obtained from \empymod. The blue circles indicate the actually computed
responses and the black dots the frequencies which are interpolated or set to
zero. The resulting time-domain response has a relative error of less than
1\,\% everywhere except for very early times.

Figure \ref{fig:GUI-FFTLog-log} shows exactly the same on a logarithmic scale,
without the interactive GUI-related elements. It can be seen that with the
chosen frequencies the time-domain response starts to divert above about 100\,s
and below 0.3\,s. It also shows the oscillating high-frequency part, which is
hard to interpolate. Figure \ref{fig:GUI-DLF-lin} is the same as
\ref{fig:GUI-FFTLog-lin}, but transformed with the DLF method applying the
81-point sine-cosine filter from \cite{GEO.09.Key}. The same frequencies were
computed as in the FFTLog case and the missing ones interpolated. The error of
the corresponding time-domain response is comparable, so either FFTLog or DLF
can be used.
%
\multiplot{2}{GUI-FFTLog-log,GUI-DLF-lin}{width=\tcwidth}
{(a) Same as Figure \ref{fig:GUI-FFTLog-lin}, but on a logarithmic scale.
(b) Using DLF instead of FFTLog, but computing the same frequencies as for the
FFTLog. The resulting time domain response has comparable accuracy (compare the
error to the error in Figure \ref{fig:GUI-FFTLog-lin}).
}
%
In the above example we used 20 frequencies, but that many would not be
required for the shown offset of 5\,km, a few of the lower and higher
frequencies could be left out. However, with this frequency-selection we can
model a wide range of offsets. This is shown in Figure \ref{fig:multi-offset},
where the above parameters were used to model the responses at offsets
$r=1.5,3,6$, and 12\,km.
%
\plot{multi-offset}{width=\tcwidth}{
  Normalized (a) frequency- and (b) time-domain responses for offsets
  $r=1.5,3,6$, and 12\,km using the parameters defined in Figure
  \ref{fig:GUI-FFTLog-lin}. The coloured circles are the actually computed
  responses, the black dots are the responses which are set to zero or
  interpolated. The black curves are the analytical responses.
}
%

The shortest offset defines the highest required frequency, and the largest
offset the lowest required frequency. So the more one can restrict the
necessary offset range, the less frequencies are needed. Another important
factor is how to interpolate and extrapolate from the computed frequencies to
the frequencies required for the Fourier transform. For the FFTLog only
extrapolation for higher and lower frequencies is required. The EM response
becomes highly oscillatory for high frequencies, which makes it very hard to
extrapolate the response to frequencies $f>f_\mathrm{max}$. However, if
$f_\mathrm{max}$ is chosen appropriately, the importance of higher frequencies
for the Fourier transform can be neglected and we can set those responses to
zero. The extrapolation of frequencies $f<f_\mathrm{min}$ can be changed to an
interpolation by assuming a zero imaginary response at a very low frequency,
and we then use PCHIP to interpolate the missing frequencies. In the case of
the DLF method we also have to interpolate in between the computed frequencies,
for which we found it better to use a cubic spline. As can be seen from Figures
\ref{fig:GUI-FFTLog-lin} and \ref{fig:GUI-DLF-lin}, using the FFTLog or the DLF
with the same actually computed frequencies results in very similar responses.

The changes to the Fourier transform in comparison with \cite{GEO.08.Mulder}
can be summarized in three points: (1) regular log-scale spacing for the
frequency selection; (2) DLF or FFTLog instead of FFT; and (3) using only the
imaginary part of the frequency-domain response. The actual speed of the
transform is unimportant, as the computation of the frequency-domain responses
takes much longer than the transform itself. What matters is solely how many
frequencies are required by it in order to be stable, and how long it takes to
compute the responses for these frequencies.

\section{Gridding}

The computation grid consists of a core or survey domain $D_\mr{s}$ which
should contain all source and receiver positions. The survey domain usually has
no or very little cell stretching $\alpha_\mr{s}$. The minimum cell width is
defined as
%
\begin{equation}
  \Delta_\mr{min}=\delta(f, \sigma_\mr{src})/n_\delta \, ,
  \label{eq:minwidth}
\end{equation}
%
where $\sigma_\mr{src}$ is the conductivity of the media in which the source
resides, and $n_\delta$ is a positive number that defines how many cells there
should be per skin depth. The actual computation domain $D_\mr{c}$ is usually
much bigger than $D_\mr{s}$ in order to avoid artefacts from the perfectly
electrically conducting boundary condition. It can also have a much higher
stretching $\alpha_\mr{c}$. In our scheme we have chosen $D_\mr{c}$ such that
the distance for the signal diffusing from the source to the boundary and back
to the receiver closest to the boundary is at least two wavelengths, after
which the initial signal is reduced to a millionth of its strength. The
wavelength to compute $D_\mr{c}$ is given by
%
\begin{equation}
  2\lambda = 4\pi\delta(f, \sigma_\mr{ave})
  \ \approx \ 6324.7/\sqrt{f\sigma_\mr{ave}}\, ,
 \label{eq:lambda}
\end{equation}
%
where $\sigma_\mr{ave}$ is the average conductivity, which can vary for
different directions. However, the skin-depth approach fails for air, in which
the EM field propagates as a wave at the speed of light. A largest
computational domain is therefore enforced, defining the maximum distance from
the source to the boundary; this distance is by default 100\,km, but this can
be reduced in the marine case with increasing water depth. Note that for
shallow marine and land cases this also applies to the horizontal dimensions,
not only to the upward $z$-direction and similarly for very resistive
basements, even in deep water. One way to circumvent this difficulty is the use
of a primary-secondary formulation, where the primary field, including the air
wave, is computed with a semi-analytical code for layered media. We do not
consider this approach here.

In summary, the adaptive gridding takes $f$, $D_\mr{s}$, $\sigma_\mr{src}$,
$\sigma_\mr{ave}$, $n_\delta$, and ranges for $\alpha_\mr{s}$, $\alpha_\mr{c}$,
where we usually fix $\alpha_\mr{s}=1$ or keep it at least below 1.05, and let
$\alpha_\mr{c}$ be anything between $[1, 1.5]$. The minimum cell width
$\Delta_\mr{min}$ can further be restricted by a user-defined range. Given
these inputs the adaptive gridding will search for the smallest possible number
of cells which fulfils these criteria. The multigrid method implemented in
\emg3d puts some constraints on the number of cells, of which the adaptive
gridding takes care (the number of cells have to be powers of two multiplied by
a low prime, e.g., $\{2,3,5\}\cdot2^n$).

The main difference with \cite{GEO.08.Mulder} is that their adaptive gridding
searches for the optimal stretching factor $\alpha$ fulfilling certain
criteria, for a fixed number of cells. Our adaptive gridding, on the other
hand, searches for the smallest number of cells that still fulfil the given
criteria. The number of cells becomes therefore also a function of frequency.
To go from the model grid to the computational grid, we use the
volume-averaging technique on logarithmic resistivities, as used in
\cite{GEO.07.Plessix}. While this technique ensures that the total resistivity
in the subsurface remains the same, it does not consider effective-medium
theory \citep{GEO.03.Davydycheva}, for instance, the apparent anisotropy from a
stack of finely layered formations of varying resistivity.

\section{Numerical examples}

\subsection{Homogeneous space}

The first example is the inline electric field from a source at the origin
measured by an inline receiver with an offset of 900\,m in a homogeneous space
of $1\,\Omega\,$m. We chose this simple example to compare it with the
analytical solution and with previously published results. We used the
following values to define the required frequencies: $f_\mr{min}=0.05\,$Hz,
$f_\mr{max}=21\,$Hz, using FFTLog with 5 frequencies per decade. This results
in 14 frequencies to compute from 0.05\,Hz to 20.0\,Hz. The complete frequency
range for the transform includes 30 frequencies from 0.0002\,Hz to 126.4\,Hz.
For the adaptive gridding the following inputs were used: $n_\delta=12$,
minimum cell width must be between 20 and 40\,m, and $\alpha_\mr{s}=1$,
$\alpha_\mr{c}=[1,1.3]$. This created grids with cell numbers between
\num{46080} ($80\times24\times24$, for 20.0\,Hz) and \num{128000}
($80\times40\times40$, for 0.05\,Hz) cells. The run times for each frequency,
the corresponding number of cells, minimum cell width and computation domain
stretching factor are listed in Table~\ref{tbl:timefull}. The total run time to
compute this model was less than two minutes.
%
\tabl[btp]{timefull}{Run times per frequency for the homogeneous space example,
  with the corresponding number of cells and minimum cell width as well as the
  stretching factor in the computation domain; $\alpha_\mr{s}=1$ everywhere.}{
  \centering
\begin{tabular}{S[table-format=2.4]rrrr}
  \toprule
  %
  {freq.} & time & nx$\times$ny$\times$nz & $\Delta_\mr{min}$ & $\alpha_\mr{c}$ \\
  {(Hz)}  & (s) &  & (m) & \\
  \midrule
  %
  20.0    &  3 & 80$\times$24$\times$24 & 20 & 1.26 \\
  12.6    &  6 & 96$\times$32$\times$32 & 20 & 1.17 \\
   7.98   &  8 & 96$\times$32$\times$32 & 20 & 1.20 \\
   5.03   &  7 & 96$\times$32$\times$32 & 20 & 1.23 \\
   3.18   &  7 & 80$\times$32$\times$32 & 24 & 1.25 \\
   2.00   &  5 & 80$\times$32$\times$32 & 30 & 1.21 \\
   1.26   &  4 & 64$\times$32$\times$32 & 37 & 1.21 \\
   0.798  &  5 & 64$\times$32$\times$32 & 40 & 1.23 \\
   0.503  &  5 & 64$\times$32$\times$32 & 40 & 1.26 \\
   0.318  &  5 & 64$\times$32$\times$32 & 40 & 1.28 \\
   0.200  & 10 & 64$\times$40$\times$40 & 40 & 1.27 \\
   0.126  & 10 & 64$\times$40$\times$40 & 40 & 1.30 \\
   0.0798 & 14 & 80$\times$40$\times$40 & 40 & 1.26 \\
   0.0503 & 15 & 80$\times$40$\times$40 & 40 & 1.28 \\
  %
  \bottomrule
\end{tabular}}%
%

Figure \ref{fig:fullspace} (a) shows the frequency-domain result, where the
blue dots are the computed responses and the black dots correspond to the
interpolated values or the values set to zero.
%
\plot*{fullspace}{width=\textwidth}{
  (a) Frequency- and (b) time-domain results for the homogeneous space model.
  The red lines are the analytical solutions, the blue circles are the actually
  computed responses with \emg3d, the black dots are the interpolated
  responses, and the dashed black line the obtained time-domain response. The
  errors are clipped for values smaller than 0.1\,\% and bigger than 100\,\%.
}
%
Most of the computed values stay below a relative error of 1\,\%, our
chosen adaptive gridding only starts to generate considerable errors at higher
frequencies. Figure \ref{fig:fullspace}(b) shows the corresponding time-domain
result, where the dashed black line is the result from \emg3d, on top of the
red line which is the analytical result. The relative error is mostly below
1\,\%, except for early times. However, for practical reasons that is more than
enough. Figure~\ref{fig:fullspace-log} shows the same on a logarithmic scale,
with times up to 10 seconds. It clearly shows that if later times are required,
we would need to adjust our Fourier transform parameters.
%
\plot{fullspace-log}{width=\cwidth}{
  Same as in Figure~\ref{fig:fullspace} (b), but on a logarithmic scale. To
  improve later times we would have to compute lower frequencies; to improve
  earlier times we would have to compute more frequencies per decade to get a
  better resolution.
}
%
Note that for the gridding we chose $n_\delta=12$, which is very dense. This
was necessary because we are relatively close to the source. If the offsets of
interest are larger this factor can be lowered considerably; $3-4$ is often
enough.

This model corresponds to the one presented in Table 1 and in Figures 3 and 4
of \cite{GEO.08.Mulder}. The response here appears to be more accurate, their
reported peak-error is roughly 1\,\%, whereas we are below 0.1\,\% at the peak
(there are no error-plots presented, so visual inspection is all we have).
However, the difference in run time is dramatic. Summing the run times for the
different frequencies of the original figure comes to a total computation time
of 3\,h 47\,min 12\,s; 0.01\,Hz was the slowest run with 31\,min 19\,s, and
2.37\,Hz was the fastest run with 2\,min 54\,s. Our example, on the other hand,
took less than two minutes in total, where the individual frequencies took
between 3 and 15\,s to run.

This massive speed-up has a couple of reasons. Computers have become more
powerful in the last twelve years, and the codes were run on different
computers. A quick test with the old scripts on our test machine shows that it
would roughly run 2-3 times faster, therefore somewhere between 1 and 2 hours.
The more important facts besides different hardware are: (1) we only used 14
frequencies instead of the 26 frequencies between 0.01 and 100\,Hz of the
original; (2) our adaptive gridding used significantly less cells
($f$-dependent) in comparison to the fixed \num{2097152} cells ($128^3$) used
in the original example. We did not see a significant difference in the speed
of the actual codes, where the kernel-algorithm of the two implementations is
the same, but in the original example it is implemented in Matlab/C, whereas
\emg3d is written in Python/Numba (Numba is a just-in-time compiler for Python
code, \citealp{Numba}).

\subsection{1D Model}

The second example is a shallow marine, layered model with 200\,m of seawater
(3\,S/m) above a halfspace of 1\,S/m, an embedded target layer at 2\,km depth,
100\,m thick, with a conductivity of 0.02\,S/m. The source is located 20\,m
above the seafloor and the receivers are on the seafloor. We chose the
frequency range such that we can model offsets from 3 to 7 kilometers, with
$f_\mr{min}=0.007\,$Hz and $f_\mr{max}=32\,$Hz, using FFTLog with 5 frequencies
per decade. This results in computations for 19 frequencies from 0.008\,Hz to
31.8\,Hz. The complete frequency range for the transform includes 35
frequencies from $2\,10^{-5}$ to $126.4\,$Hz. For the adaptive gridding, we
used a cell width of 100\,m in the core domain and stretching outside up to a
factor 1.5, where the computation domain extends up to 50\,km in each
direction. This yielded grids between \num{204800} (higher frequencies) and
\num{245760} (lower frequencies) cells. The run times for each frequency and
their corresponding parameters are listed in Table~\ref{tbl:timemarine}.
%
\tabl[btp]{timemarine}{Run times per frequency for the marine 1D example,
  with the corresponding number of cells and minimum cell width as well as the
  stretching factor in the computation domain; $\alpha_\mr{s}=1$ everywhere.}{
  \centering
\begin{tabular}{S[table-format=2.5]rrrr}
  \toprule
  %
  {freq.} & time & nx$\times$ny$\times$nz & $\Delta_\mr{min}$ & $\alpha_\mr{c}$ \\
  {(Hz)}  & (s) &  & (m) &  \\
  \midrule
  %
  31.8     & 10 & 128$\times$40$\times$40 & 100 & 1.36 \\
  20.0     &  9 & 128$\times$40$\times$40 & 100 & 1.36 \\
  12.6     &  9 & 128$\times$40$\times$40 & 100 & 1.36 \\
   7.98    & 17 & 128$\times$40$\times$40 & 100 & 1.36 \\
   5.03    & 21 & 128$\times$40$\times$40 & 100 & 1.44 \\
   3.18    & 20 & 128$\times$40$\times$40 & 100 & 1.48 \\
   2.00    & 18 & 128$\times$40$\times$40 & 100 & 1.49 \\
   1.26    & 22 & 128$\times$40$\times$48 & 100 & 1.36 \\
   0.798   & 26 & 128$\times$40$\times$48 & 100 & 1.36 \\
   0.503   & 29 & 128$\times$40$\times$48 & 100 & 1.36 \\
   0.318   & 42 & 128$\times$40$\times$48 & 100 & 1.36 \\
   0.200   & 41 & 128$\times$40$\times$48 & 100 & 1.38 \\
   0.126   & 48 & 128$\times$40$\times$48 & 100 & 1.40 \\
   0.0798  & 54 & 128$\times$40$\times$48 & 100 & 1.41 \\
   0.0503  & 59 & 128$\times$40$\times$48 & 100 & 1.44 \\
   0.0318  & 74 & 128$\times$40$\times$48 & 100 & 1.44 \\
   0.0200  & 81 & 128$\times$40$\times$48 & 100 & 1.47 \\
   0.0126  & 96 & 128$\times$40$\times$48 & 100 & 1.48 \\
   0.00798 & 93 & 128$\times$40$\times$48 & 100 & 1.49 \\
  %
  \bottomrule
\end{tabular}}%
%

Figure~\ref{fig:marine} shows the result for an offset of 5\,km, in (a) the
frequency and (b) the time domain. The recovered response with the 3D code
captures the airwave (first peak) and the subsurface (second peak) very
accurately. At later times the error starts to increase. We would need to
compute a few additional lower frequencies if we want to improve it.
%
\plot*{marine}{width=\textwidth}{
  Electric inline response at an offset of 5\,km for a shallow marine, layered
  scenario. (a) Frequency-domain, where the blue circles are computed responses
  and the black dots are interpolated responses or responses set to zero. (b)
  Time-domain response.
}
%
In the frequency-domain plot, it can be seen that the high frequencies are not
computed very accurately, but without too much influence on the time-domain
response. These frequencies could be left out if an offset of 5\,km is the only
objective. However, we also want to retrieve shorter offsets from the same
computation, for which these frequencies are required.

Figure~\ref{fig:marine-multioffset} shows the time-domain responses of the same
model for offsets of 3, 5, and 7\,km, all obtained with the same
frequency-domain computations and the same frequencies for the Fourier
transform. The computation of these frequencies took less than 13 minutes and
it handles any offset between 3 and 7\,km. It can be seen that the chosen
frequency selection is sufficient for this offset range; again, more low
frequencies could be added to improve late-time values.
%
\plot{marine-multioffset}{width=\cwidth}{
  Time-domain responses for offsets of 3, 5, and 7\,km for the same model as
  shown in Figure~\ref{fig:marine}.
}
%

\subsection{Horizontal extent of the computation domain}

The skin-depth approach fails for the air layer, as explained in the Section
\emph{Gridding}. The reason is that the EM field in the air travels at the
speed of light as a wave, and its amplitude is only reduced through geometrical
spreading. On land and in shallow marine scenarios one has therefore to include
a sufficiently large computational domain, and the default in our scheme is
100\,km. The important point is that this does not only apply to the upward
$z$-direction, but also to the horizontal directions, as the airwave also
bounces back horizontally and would continuously emit energy into the
subsurface if the boundaries are not chosen far enough away from the receivers.
If models are computed with very resistive layers or models with highly
resistive basements, this can even apply to deep marine scenarios.

Figure~\ref{fig:marine-wrong-x-y} shows this effect. It is the same model as in
the previous section; however, for the adaptive gridding in the horizontal
directions $\rho_\mr{ave}=1\,\Omega\,$m was used instead of
$\rho_\mr{ave}=\num{10000}\,\Omega\,$m. Having the boundaries too near in the
horizontal directions leads to worse results for most frequencies and entirely
wrong results for high frequencies. Comparison with the 1D result in the time
domain shows that it is the airwave whose amplitude is heavily overestimated.
%
\plot*{marine-wrong-x-y}{width=\textwidth}{
  Same model as used for Figure~\ref{fig:marine}, but with the horizontal
  boundaries not far enough. Although the resulting time-domain result looks
  plausible, the comparison with the 1D result shows that it significantly
  overestimates the amplitude of the airwave.
}
%
It can be difficult to spot these errors in the time-domain result, as the
response looks plausible and only comparing it with the 1D result reveals that
it is actually wrong. A possibility to detect such problems for complicated
cases, where there is no semi-analytical result to compare with, is to compute
two or more models, moving the boundary. When the responses stop to change, one
can assume that the boundary is far enough. Another possibility is to look at
the amplitudes close to the boundaries and ensure that they are small enough.

\subsection{3D Model}

The final example consists of a resistive, three-dimensional block embedded in
the lower of two halfspaces, as depicted in Figure~\ref{fig:3d-model}. The
target has resistivity $\rho_\mr{tg}=100\,\Omega\,$m, the upper halfspace
corresponds to seawater with $\rho_\mr{sea}=0.3\,\Omega\,$m, and the lower
halfspace is the background with $\rho_\mr{bg}=1\,\Omega\,$m.
%
\plot{3d-model}{width=\cwidth}{
  Three-dimensional block embedded in the lower of two halfspaces. The 100-m
  long, $x$-directed dipole source is located 50\,m above the seafloor at the
  origin, and the receiver is on the seafloor at an offset of 2\,km.
}
%
The source is a 100-m long, $x$-directed dipole at the origin, 50\,m above the
seafloor, and we are using a step-off source function. The $x$-directed inline
receiver is at an offset of 2\,km. The dimension of the target cube is
$1.1\times1.0\times0.4$\,km, located 300\,m below the seafloor in the centre
between source and receiver.

For the comparison we use the open-source code \simpeg \citep{CAG.15.Cockett}.
\simpeg is a framework for modelling and inversion of geophysical data such as
gravity, magnetics, and CSEM. It has Maxwell's equations implemented in both
the frequency and time domain. As such we can compare our result computed in
the frequency domain followed by a Fourier transform to a result computed
directly in the time domain. A principal difference between \simpeg and \emg3d
is that the former has various direct solvers implemented, whereas the latter
is an iterative multigrid solver. The 3D model is therefore a rather small
example in order to be able to run it on our test machine, as the memory
requirement by the direct solver would otherwise be too high.

The model was discretised with $100\times100\times100\,$m cells in the survey
domain $D_\mr{s}$. For the \simpeg model, 14 cells in $x$-direction and 12
cells in $y$- and $z$-directions were used on both sides with a stretching of
1.3 for the total computation domain $D_\mr{c}$, which yields a mesh of
\num{58344} cells. The time-steps start at 0.1\,s and are: $21\times0.01\,$s,
$23\times0.03\,$s, $21\times0.1\,$s, $23\times0.3\,$s, covering exactly the
desired range of $0.1-10\,$s. For \emg3d, the mesh is generated
frequency-dependent as in the previous examples, with a maximum stretching of
$\alpha_\mr{c}=1.5$. This results in meshes between \num{18432} cells for the
highest frequencies and \num{76800} cells for the lowest frequencies. The
required frequencies were obtained by using the FFTLog with five points per
decade, which results in 20 frequencies between 0.001\,Hz and 8\,Hz. The actual
transform was carried out with the 201-point sine-cosine filter from
\cite{GEO.09.Key}.

The results are shown in Figure~\ref{fig:3d-result}: In (a) the 1D background
responses and the relative error using the semi-analytical result obtained from
\empymod, and in (b) the responses including the target.
%
\plot{3d-result}{width=\textwidth}{
  Responses for the model outlined in Figure~\ref{fig:3d-model} using \simpeg
  and \emg3d, and for the layered background also \empymod. In the lower plot
  of (a) the relative error (\%) in comparison to \empymod is shown, and in (b)
  the normalized difference (\%) between the two 3D codes. The results of these
  plots are clipped for values smaller than 0.1\,\% and bigger than 10\,\%.
}
%
The background comparison shows that both 3D codes do an acceptable job having
a relative error of a few percents at most; the result for \emg3d seems to be
better at early times. We cannot compare the errors for the response that
includes the target for lack of an analytical solution. The background model
from \empymod is only included to show that there is a significant response
from the target. We therefore show the normalized difference (ND) between the
two 3D codes as a percentage, where $\mr{ND}=
50|(\simpeg-\emg3d)/(\simpeg+\emg3d)|$. The ND between the two codes is below
1\,\% everywhere except for early times.

Both codes took a little over 7 minutes to compute the two models (single
thread). However, in this particular comparison, the main difference in runtime
is not frequency-domain computation vs. time-domain computation, but iterative
solver vs. direct solver. For a test we doubled the number of cells in each
direction by using 50\,m for the minimum cell width instead of 100\,m, and
taking also half of the stretching, which results then in a mesh of
\num{466752} cells for \simpeg and meshes between \num{102400} and \num{614400}
for \emg3d. Because the memory of our test machine was not large enough for the
direct solver of \simpeg, the last model had to be run on a cluster. In this
scenario, \emg3d was about eight times faster (single thread).

\section{Conclusions}

We have shown that time-domain CSEM modelling with a frequency-domain code can
be competitive given a robust frequency-domain solver, a frequency-dependent
gridding function that minimizes the required cells, and a Fourier transform
that works on a logarithmic scale. Fast layered modelling can be used to design
the required frequency range, as the Fourier transform does not know about the
dimensionality of the underlying model. Twenty frequencies or less are usually
sufficient for a wide range of offsets. We believe that our proposed
improvements to the previously published methods make time-domain CSEM
modelling with a frequency-domain code even more attractive than it was already
before.

\section{Acknowledgment}

This research was conducted within the Gitaro.JIM project funded through
MarTERA as part of Horizon 2020, a funding scheme of the European Research Area
(ERA-NET Cofund, \href{https://www.martera.eu}{martera.eu}).

\section{Data Availability Statement}

The data that support the findings of this study are openly available at Zenodo
at \href{https://zenodo.org/badge/DOI/10.5281/zenodo.???????}{doi:
10.5281/zenodo.???????}. The data includes the scripts and instructions to
reproduce all results and figures.

\emph{Note to editors and reviewers: The data is NOT YET on Zenodo, we will
  upload there the final versions and link the right DOI for the final
  manuscript. Until then everything can be found on GitHub at
  \href{https://github.com/empymod/article-TDEM}%
  {github.com/empymod/article-TDEM}.}

% REFERENCES

\bibliographystyle{Refs}          % Modified SEG bibliography style.
\bibliography{Refs}

\end{document}

### 2020-03-05 notebooks/freqselect.py

@@ -1,0 +1,503 @@
+import emg3d
+import empymod
+import numpy as np
+import ipywidgets as widgets
+import matplotlib.pyplot as plt
+from IPython.display import display
+
+
+class InteractiveFrequency(emg3d.utils.Fourier):
+    """App to create required frequencies for Fourier Transform."""
+
+    def __init__(self, src_z, rec_z, depth, res, time, signal=0, ab=11,
+                 aniso=None, **kwargs):
+        """App to create required frequencies for Fourier Transform.
+
+        No thorough input checks are carried out. Rubbish in, rubbish out.
+
+        See empymod.model.dipole for details regarding the modelling.
+
+
+        Parameters
+        ----------
+        src_z, rec_z : floats
+            Source and receiver depths and offset. The source is located at
+            src=(0, 0, src_z), the receiver at rec=(off, 0, rec_z).
+
+        depth : list
+            Absolute layer interfaces z (m); #depth = #res - 1
+            (excluding +/- infinity).
+
+        res : array_like
+            Horizontal resistivities rho_h (Ohm.m); #res = #depth + 1.
+
+        time : array_like
+            Times t (s).
+
+        signal : {0, 1, -1}, optional
+            Source signal, default is 0:
+                - -1 : Switch-off time-domain response
+                - 0 : Impulse time-domain response
+                - +1 : Switch-on time-domain response
+
+        ab : int, optional
+            Source-receiver configuration, defaults to 11. (See
+            empymod.model.dipole for all possibilities.)
+
+        aniso : array_like, optional
+            Anisotropies lambda = sqrt(rho_v/rho_h) (-); #aniso = #res.
+            Defaults to ones.
+
+        ft : {'sin', 'cos'}, optional
+            Flag to choose either the Sine or Cosine Digital Linear Filter
+            method.  Defaults to 'sin'.
+            If signal is < 0, it is set to cosine. If signal > 0, it is set to
+            sine. Both can be used for signal = 0.
+
+        ftarg : dict or list, optional
+            [fftfilt, pts_per_dec]:
+
+            - fftfilt: string of filter name in ``empymod.filters`` or the
+                       filter method itself. (Default:
+                       ``empymod.filters.key_201_CosSin_2012()``)
+            - pts_per_dec: points per decade; (default: 4)
+                - If 0: Standard DLF.
[…]

### 2020-06-22 ManuscriptTime.tex
Improve environment and instructions

@@
 \newcommand{\simpeg}{\texttt{SimPEG}\xspace}
+\newcommand{\custem}{\texttt{custEM}\xspace}
 
@@
 E-mail: \href{mailto:Dieter@Werthmuller.org}{Dieter@Werthmuller.org};\\
-\textbf{Keywords}: CSEM, Fourier transforms, 3D modelling.
+\textbf{Keywords}: Fourier transforms, CSEM, 3D modelling.
 }
@@
 robust. As the frequency range spans many orders of magnitude, the required
-frequencies are ideally equally spaced on a logarithmic scale, as is the case
-for, e.g., the digital linear filter method or the logarithmic fast Fourier
-transform. With careful selection of frequencies and interpolation, 15 to 20
-frequencies are enough to cover a wide range of offsets in one-dimensional test
-models.
+frequencies are ideally equally spaced on a logarithmic scale. We present a
+fast frequency-to-time domain method using either the digital linear filter
+method or the logarithmic fast Fourier transform together with careful
+selection of frequencies and interpolation. 15 to 20 frequencies are generally
+enough with this methodology to cover a wide range of offsets in
+one-dimensional test models.
 %
-The gridding should be frequency-dependent, which is generally accomplished by
-making it a function of skin depth. Optimising for the least number of required
-cells for a given frequency should be combined with optimising for
-computational speed.
+The gridding should be frequency-dependent, which is accomplished by making it
+a function of skin depth. Optimising for the least number of required cells for
+a given frequency should be combined with optimising for computational speed.
 %
@@
 
-Acquired CSEM data are often analysed (modelled and inverted for) in their
+CSEM is usually divided into frequency- and time-domain methods, depending
+whether the source signal is a continuous waveform, such as a sine, or a finite
+waveform, such as a pseudo-random binary sequence (PRBS). A numerical
+comparison of the two methods is given in \cite{GP.13.Conell}. Acquired CSEM
+data are subsequently often analysed (modelled and inverted for) in their
 respective domain, either the frequency or time domain. Modellers of layered
@@
 CSEM codes for 2D and 3D computations, on the other hand, mostly compute their
-responses directly in the required domain, either frequency or time. However,
-under certain conditions it can be competitive to model time-domain data with a
-frequency-domain code, as shown by \cite{GEO.08.Mulder}. The conditions are: a
-sufficiently powerful solver, appropriate frequency selection and
-interpolation, and an automated gridding, for which they used the
-multi-frequency CSEM approach presented by \cite{GEO.07.Plessix}. We build upon
-these results but improve the run time from hours to minutes. The main reasons
-for this significant speed-up are an adaptive, frequency-dependent gridding
-scheme that minimizes the required cells in each dimension, and a logarithmic
-Fourier transform such as digital linear filters \citep[DLF, ][]{GP.71.Ghosh}
-or the logarithmic fast Fourier transform \citep[FFTLog, ][]{RAS.00.Hamilton}
-to go from the frequency to the time domain. The latter makes it also possible
-to only use the imaginary part of the frequency-domain response, which has
-advantages when it comes to interpolation.
+responses directly in the required domain, either frequency or time (see,
+e.g., \cite{SG.10.Borner, SG.05.Avdeev} for comprehensive overviews of CSEM
+codes). However, under certain conditions it can be competitive to model
+time-domain data with a frequency-domain code, as shown by
+\cite{GEO.08.Mulder}. The conditions are: a sufficiently powerful solver,
+appropriate frequency selection and interpolation, and an automated gridding,
+for which they used the multi-frequency CSEM approach presented by
+\cite{GEO.07.Plessix}. We build upon these results but improve the run time
+from hours to minutes. The main reasons for this significant speed-up are an
+adaptive, frequency-dependent gridding scheme that minimizes the required cells
+in each dimension, and a logarithmic Fourier transform such as digital linear
+filters \citep[DLF, ][]{GP.71.Ghosh} or the logarithmic fast Fourier transform
+\citep[FFTLog, ][]{RAS.00.Hamilton} to go from the frequency to the time
+domain. The latter makes it also possible to only use the imaginary part of the
+frequency-domain response, which has advantages when it comes to interpolation.
 
@@
 
-Time-domain modelling with a frequency-domain code requires the computation of
-a relatively wide range of frequencies for the Fourier transform. However, the
-required computation grids for low frequencies (in our case in the order of
-0.001\,Hz) and high frequencies (around 100\,Hz) are hugely different. Low
-frequencies can be computed on a coarser grid, but they require a much larger
-domain in order to avoid boundary effects. High frequencies, on the other hand,
-require denser gridding, but they are much more limited in reach. An adaptive
-gridding scheme is therefore required. We build our approach upon
-\cite{GEO.07.Plessix}, who presented such an adaptive gridding for
-multi-frequency (and multi-source) CSEM modelling.
+Being able to model CSEM data both in the frequency-domain and in the
+time-domain can be desirable, as both domains have advantages and
+disadvantages. One way to achieve this is to implement Maxwell's equation in
+both domains, as it is done, for instance, in \simpeg \citep{CAG.15.Cockett}.
+Another method is to have Maxwell's equation only implemented in one domain,
+and use Fourier transforms to go to the other domain. However, this approach
+can be costly, as many frequencies over a wide range are required to go from
+the time domain to the frequency domain, or many times over a wide range for
+the opposite direction. We present a methodology which significantly reduces
+the amount and range of the required frequencies, which translates into a
+significant faster computation times. The required computation grids for low
+frequencies (in our case in the order of 0.001\,Hz) and high frequencies
+(around 100\,Hz) are hugely different. Low frequencies can be computed on a
+coarser grid, but they require a much larger domain in order to avoid boundary
+effects. High frequencies, on the other hand, require denser gridding, but they
+are much more limited in reach. An adaptive gridding scheme is therefore
+required. We build our approach upon \cite{GEO.07.Plessix}, who presented such
+an adaptive gridding for multi-frequency (and multi-source) CSEM modelling.
 
 An adaptive grid in the frequency domain is naturally based on the skin depth,
-which is the distance after with the amplitude of the electromagnetic field is
+which is the distance after which the amplitude of the electromagnetic field is
 decayed by $1/e\approx 37\,\%$. The skin depth $\delta$ is a function of
@@
 frequency-domain code even more competitive in comparison with time-domain
-codes.
+codes. It is important to note that while we show the transform from frequency
+to time domain, the conclusions regarding the Fourier transform can equally
+well be applied to the transform from time to frequency domain data for a code
+that solves Maxwell's equations in the time domain. A variation of our method
+was implemented, for instance, by \custem \cite{GEO.19.Rochlitz} to obtain
+frequency-domain data from time-domain computations (Rochlitz, personal
+communication).
 
@@
 example in order to be able to run it on our test machine, as the memory
-requirement by the direct solver would otherwise be too high.
+requirement by the direct solver would otherwise be too high. There are not
+many options out there of open-source time-domain 3D codes, \simpeg being the
+one we found to be suitable. A step-off response was chosen as this is the
+response currently implemented in \simpeg.
 
@@
 
-We have shown that time-domain CSEM modelling with a frequency-domain code can
-be competitive given a robust frequency-domain solver, a frequency-dependent
-gridding function that minimizes the required cells, and a Fourier transform
-that works on a logarithmic scale. Fast layered modelling can be used to design
-the required frequency range, as the Fourier transform does not know about the
-dimensionality of the underlying model. Twenty frequencies or less are usually
-sufficient for a wide range of offsets. We believe that our proposed
-improvements to the previously published methods make time-domain CSEM
-modelling with a frequency-domain code even more attractive than it was already
-before.
+We have shown a method to minimize the required frequencies and their range for
+the computation of time-domain CSEM data with a frequency-domain code. This can
+significantly reduce the computation time and makes time-domain CSEM modelling
+with a frequency-domain code competitive given a robust frequency-domain
+solver, a frequency-dependent gridding function that minimizes the required
+cells, and a Fourier transform that works on a logarithmic scale. Fast layered
+modelling can be used to design the required frequency range, as the Fourier
+transform does not know about the dimensionality of the underlying model.
+Twenty frequencies or less are usually sufficient for a wide range of offsets.
+We believe that our proposed improvements to the previously published methods
+make time-domain CSEM modelling with a frequency-domain code even more
+attractive than it was already before. We have demonstrated the idea of our
+Fourier transform method on CSEM data transformed from the frequency to the
+time domain. However, it could equally be applied to the transform from the
+time to the frequency domain and to other methods with similar
+characteristics.
 

### 2020-06-22 notebooks/freqselect.py

@@ -48,27 +48,6 @@ class InteractiveFrequency(emg3d.utils.Fourier):
             Anisotropies lambda = sqrt(rho_v/rho_h) (-); #aniso = #res.
             Defaults to ones.
 
-        ft : {'sin', 'cos'}, optional
-            Flag to choose either the Sine or Cosine Digital Linear Filter
-            method.  Defaults to 'sin'.
-            If signal is < 0, it is set to cosine. If signal > 0, it is set to
-            sine. Both can be used for signal = 0.
-
-        ftarg : dict or list, optional
-            [fftfilt, pts_per_dec]:
-
-            - fftfilt: string of filter name in ``empymod.filters`` or the
-                       filter method itself. (Default:
-                       ``empymod.filters.key_201_CosSin_2012()``)
-            - pts_per_dec: points per decade; (default: 4)
-                - If 0: Standard DLF.
-                - If < 0: Lagged Convolution DLF.
-                - If > 0: Splined DLF
-
-            The values can be provided as dict with the keywords, or as list.
-            However, if provided as list, you have to follow the order given
-            above.
-
         **kwargs : Optional parameters:
 
             - ``fmin`` : float
@@ -80,8 +59,8 @@ class InteractiveFrequency(emg3d.utils.Fourier):
             - ``off`` : float
               Initial offset. Default is 500.
 
-            - ``ft`` : str {'sin', 'cos', 'fftlog'}
-              Initial Fourier transform method. Default is 'sin'.
+            - ``ft`` : str {'dlf', 'fftlog'}
+              Initial Fourier transform method. Default is 'dlf'.
 
             - ``ftarg`` : dict
               Initial Fourier transform arguments corresponding to ``ft``.
@@ -104,7 +83,7 @@ class InteractiveFrequency(emg3d.utils.Fourier):
         fmin = kwargs.pop('fmin', 1e-3)
         fmax = kwargs.pop('fmax', 1e1)
         off = kwargs.pop('off', 5000)
-        ft = kwargs.pop('ft', 'sin')
+        ft = kwargs.pop('ft', 'dlf')
         ftarg = kwargs.pop('ftarg', None)
         self.pts_per_dec = kwargs.pop('pts_per_dec', 5)
         self.linlog = kwargs.pop('linlog', 'linear')
@@ -136,8 +115,8 @@ class InteractiveFrequency(emg3d.utils.Fourier):
         """Create the figure."""
 
[…]

### 2020-06-30 ManuscriptTime.tex

@@
 robust. As the frequency range spans many orders of magnitude, the required
-frequencies are ideally equally spaced on a logarithmic scale. We present a
-fast frequency-to-time domain method using either the digital linear filter
-method or the logarithmic fast Fourier transform together with careful
-selection of frequencies and interpolation. 15 to 20 frequencies are generally
-enough with this methodology to cover a wide range of offsets in
-one-dimensional test models.
+frequencies are ideally equally spaced on a logarithmic scale. The fast
+frequency-to-time domain method uses either the digital linear filter method
+or the logarithmic fast Fourier transform together with a careful selection of
+frequencies and interpolation. This methodology requires typically 15 to 20
+frequencies to cover a wide range of offsets in one-dimensional test models.
 %
@@
 %
-Looking carefully at these three points results in much faster computation
-times with speedup-factors of ten and more over previous results. A
+Looking carefully at these three points results in much smaller computation
+times with speedup factors of ten or more over previous results. A
 frequency-domain computation of time-domain data can therefore be competitive
-to time-domain solvers if the required frequencies and their corresponding
-grids are carefully chosen.
+to time-domain solvers if the required frequencies and the corresponding grids
+are carefully chosen.
 %
@@
 The controlled-source electromagnetic (CSEM) method is one of the common
-non-seismic tools in exploration geophysics, not just in hydrocarbon
+non-seismic tools in exploration geophysics, not only in hydrocarbon
 exploration \citep{GEO.10.Constable}, but also in the search for sulfides
-\citep{GRL.19.Gehrmann}, water \citep{GEO.05.Pedersen}, or geothermal sources
-\citep{WGC.15.Girard}. While in the deep marine environment current sources
-with few frequencies are used, transient measurements are more common in the
-shallow marine environment and on land. One of the main reasons is the
-dominance of the airwave in shallow marine and terrestrial measurements, which
-can be better separated in the time domain.
-
+\citep{GRL.19.Gehrmann}, water \citep{GEO.05.Pedersen}, geothermal sources
+\citep{WGC.15.Girard}, or geological purposes \citep{NAT.19.Johanson}. While
+current sources with a few frequencies are used in the deep marine environment,
+transient measurements are more common in the shallow marine environment and on
+land \citep[e.g., ][]{GEO.07.Ziolkowski, SEG.07.Andreis, SEG.07.Avdeeva}. One
+of the main reasons is the dominance of the airwave in shallow marine and
+terrestrial measurements, which can be better separated in the time domain.
 CSEM is usually divided into frequency- and time-domain methods, depending
@@
 waveform, such as a pseudo-random binary sequence (PRBS). A numerical
-comparison of the two methods is given in \cite{GP.13.Conell}. Acquired CSEM
+comparison of the two methods is given by \cite{GP.13.Conell}. Acquired CSEM
 data are subsequently often analysed (modelled and inverted for) in their
@@
 CSEM codes for 2D and 3D computations, on the other hand, mostly compute their
-responses directly in the required domain, either frequency or time (see,
-e.g., \cite{SG.10.Borner, SG.05.Avdeev} for comprehensive overviews of CSEM
-codes). However, under certain conditions it can be competitive to model
-time-domain data with a frequency-domain code, as shown by
-\cite{GEO.08.Mulder}. The conditions are: a sufficiently powerful solver,
-appropriate frequency selection and interpolation, and an automated gridding,
-for which they used the multi-frequency CSEM approach presented by
-\cite{GEO.07.Plessix}. We build upon these results but improve the run time
-from hours to minutes. The main reasons for this significant speed-up are an
-adaptive, frequency-dependent gridding scheme that minimizes the required cells
-in each dimension, and a logarithmic Fourier transform such as digital linear
-filters \citep[DLF, ][]{GP.71.Ghosh} or the logarithmic fast Fourier transform
-\citep[FFTLog, ][]{RAS.00.Hamilton} to go from the frequency to the time
-domain. The latter makes it also possible to only use the imaginary part of the
-frequency-domain response, which has advantages when it comes to interpolation.
+responses directly in the required domain, either frequency or time. There is a
+wealth of 3D electromagnetic codes, with a first boom in the 1990's, with a
+good state of the art provided by \cite{B.SEG.99.Oristaglio}. A second wave
+occurred during the CSEM boom in the hydrocarbon industry in the early 21th
+century, see, e.g., \cite{SG.05.Avdeev} and \cite{SG.10.Borner} for
+comprehensive overviews. Recent publications span the width of numerical
+methods from integral equation codes \citep{MGS.17.Kruglyakov} and differential
+equation methods such as finite differences \citep{CAG.13.Sommer}, finite
+elements \citep{GJI.13.Grayver}, and finite volumes \citep{GEO.14.Jahandari}.
+Besides the different methods there are also many different types of
+discretisation. Commonly used, besides rectilinear Yee grids
+\citep{IEEE.66.Yee} as used by our code, are Lebedev grids
+\citep{CMMP.64.Lebedev} but also unstructured grids using, for instance,
+tetrahedral \citep{CAG.17.Cai}. The computational requirement to model CSEM
+data also depends heavily on the technique which is used to solve the system,
+where there are direct solvers \citep{GEO.15.Grayver}, indirect solvers
+\citep{GJI.15.Jaysaval}, and more recently also hybrid solvers
+\citep{GEO.18.Liu}.
 
-In the next section, we briefly review the methods and highlight their
-advantages and shortcomings. This is followed by an outline of our changes to
-the Fourier transform and the adaptive gridding. Finally, we demonstrate the
-efficiency of the approach with some numerical results.
+Under certain conditions it can be competitive to model time-domain data with a
+frequency-domain code, as shown by \cite{GEO.08.Mulder}. The conditions are: a
+sufficiently powerful solver, appropriate frequency selection and
+interpolation, and an automated gridding, for which they used the
+multi-frequency CSEM approach presented by \cite{GEO.07.Plessix}. We build upon
+these results but improve the run time from hours to minutes. The main reasons
+for this significant speed-up are an adaptive, frequency-dependent gridding
+scheme that minimizes the required cells in each dimension, and a logarithmic
+Fourier transform such as digital linear filters \citep[DLF, ][]{GP.71.Ghosh}
+or the logarithmic fast Fourier transform \citep[FFTLog, ][]{RAS.00.Hamilton}
+to go from the frequency to the time domain. The latter makes it also possible
+to only use the imaginary part of the frequency-domain response, which has
+advantages when it comes to interpolation.
+
+In the next section, we briefly review the methodology as introduced by
+\cite{GEO.07.Plessix} and \cite{GEO.08.Mulder} and highlight their advantages
+and shortcomings. This is followed by an outline of our changes to the Fourier
+transform and the adaptive gridding. Finally, we demonstrate the efficiency of
+the approach with some numerical results.
 
@@
 both domains, as it is done, for instance, in \simpeg \citep{CAG.15.Cockett}.
-Another method is to have Maxwell's equation only implemented in one domain,
-and use Fourier transforms to go to the other domain. However, this approach
-can be costly, as many frequencies over a wide range are required to go from
-the time domain to the frequency domain, or many times over a wide range for
-the opposite direction. We present a methodology which significantly reduces
-the amount and range of the required frequencies, which translates into a
-significant faster computation times. The required computation grids for low
+Another approach is to have Maxwell's equation only implemented in one domain,
+and use Fourier transforms to go to the other. However, this approach can be
+costly, as many frequencies over a wide range are required to go from the time
+domain to the frequency domain, or many times over a wide range for the
+opposite direction. We present a methodology which significantly reduces the
+amount and range of the required frequencies, translating into a significant
+reduction in computation time. The required computation grids for low
 frequencies (in our case in the order of 0.001\,Hz) and high frequencies
@@
 fraction of the minimum skin depth $\delta_\mathrm{min}$, where
-$\delta_\mathrm{min}$ should be $2-3$ times bigger than $\Delta_\mathrm{min}$.
-The smallest cells have to be around the source; in the marine case the minimum
-skin depth is therefore defined by the conductivity of seawater. However, this
-yields quite large cells for low frequencies, so special care has to be taken
-around the source. The grid dimension, on the other hand, is defined as a
+$\delta_\mathrm{min}$ should be 2--3 times bigger than $\Delta_\mathrm{min}$.
+The cells have to be smallest around the source; in the marine case, the
+minimum skin depth is therefore defined by the conductivity of seawater.
+However, this yields quite large cells for low frequencies, so special care has
+to be taken around the source by defining a maximum allowed
+$\Delta_\mathrm{min}$. The grid dimension, on the other hand, is defined as a
 function of the skin depth for the average conductivity of the background,
@@
 if required, hence if certain criteria of response stability are not met. While
-this method is good and effective for a single offset, it looses all its
+this method is good and effective for a single offset, it loses all its
 advantages if one tries to compute different offsets within one computation, as
@@
 that solves Maxwell's equations in the time domain. A variation of our method
-was implemented, for instance, by \custem \cite{GEO.19.Rochlitz} to obtain
-frequency-domain data from time-domain computations (Rochlitz, personal
-communication).
+was implemented, for instance, by \custem \citep{GEO.19.Rochlitz} to obtain
+frequency-domain data from time-domain computations (Rochlitz, private
+communication, 2020).
 
@@
 where $\mathbf{E}$ is the electric field (V/m) and $\mathbf{J}_\mathrm{s}$ the
-current source (A/m$^2$); time convention is $\exp(\mathrm{i}\omega t)$. The
+current source (A/m$^2$); time dependence is $\exp(\mathrm{i}\omega t)$. The
 standard multigrid approach fails for severe stretching or strong anisotropy,
@@
 obtained from frequency-domain computations is the selection of the required
-frequencies. The fewer frequencies required, the faster we obtain the
+frequencies. The fewer frequencies required, the quicker we obtain the
 time-domain result. We decided to use a regular spacing of frequencies on a
@@
 by using only the imaginary part of the frequency-domain response. The
-imaginary part has the characteristics that its value goes to zero when the
-frequency goes to zero or to infinity, with the advantage that knowing the
-endpoints makes it possible to convert the extrapolation of missing frequencies
-into interpolation.
+imaginary part goes to zero when the frequency goes to zero or to infinity,
+with the advantage that knowing the endpoints makes it possible to convert the
+extrapolation of missing frequencies into interpolation.
 
@@
   model parameters are described in the text). This example shows the impulse
-  response at an offset of 5\,km, for which it uses FFTlog from 0.001\,Hz to
+  response at an offset of 5\,km, for which it uses FFTLog from 0.001\,Hz to
   10\,Hz with five frequencies per decade. Note that the error is clipped for
@@
 $\rho=0.3\,\Omega$\,m ($\rho = \sigma^{-1}$) and a 100\,m thick target of
-100\,$\Omega$\,m 1\,km below the seafloor in a background of 1\,$\Omega$\,m.
+100\,$\Omega$\,m at 1\,km below the seafloor in a background of 1\,$\Omega$\,m.
 The source is 50\,m above the seafloor, the receiver is on the seafloor, and
@@
 extrapolate the response to frequencies $f>f_\mathrm{max}$. However, if
-$f_\mathrm{max}$ is chosen appropriately, the importance of higher frequencies
+$f_\mathrm{max}$ is chosen judiciously, the importance of higher frequencies
 for the Fourier transform can be neglected and we can set those responses to
 zero. The extrapolation of frequencies $f<f_\mathrm{min}$ can be changed to an
-interpolation by assuming a zero imaginary response at a very low frequency,
-and we then use PCHIP to interpolate the missing frequencies. In the case of
-the DLF method we also have to interpolate in between the computed frequencies,
-for which we found it better to use a cubic spline. As can be seen from Figures
-\ref{fig:GUI-FFTLog-lin} and \ref{fig:GUI-DLF-lin}, using the FFTLog or the DLF
-with the same actually computed frequencies results in very similar responses.
+interpolation by assuming a zero imaginary response at zero frequency, and we
+then use PCHIP to interpolate the missing frequencies (as we work on a
+logarithmic scale we cannot choose 0\,Hz, but instead take $10^{-100}$\,Hz). In
+the case of the DLF method we also have to interpolate in between the computed
+frequencies, for which we found it better to use a cubic spline. As can be seen
+from Figures \ref{fig:GUI-FFTLog-lin} and \ref{fig:GUI-DLF-lin}, using the
+FFTLog or the DLF with the same actually computed frequencies results in very
+similar responses.
 
@@
 was necessary because we are relatively close to the source. If the offsets of
-interest are larger this factor can be lowered considerably; $3-4$ is often
+interest are larger this factor can be lowered considerably; 3--4 is often
 enough.
@@
 computers. A quick test with the old scripts on our test machine shows that it
-would roughly run 2-3 times faster, therefore somewhere between 1 and 2 hours.
+would roughly run 2--3 times faster, therefore somewhere between 1 and 2 hours.
 The more important facts besides different hardware are: (1) we only used 14
@@
 $23\times0.03\,$s, $21\times0.1\,$s, $23\times0.3\,$s, covering exactly the
-desired range of $0.1-10\,$s. For \emg3d, the mesh is generated
+desired range of 0.1--10\,s. For \emg3d, the mesh is generated
 frequency-dependent as in the previous examples, with a maximum stretching of
@@
 (ERA-NET Cofund, \href{https://www.martera.eu}{martera.eu}).
+An early idea of this manuscript was presented at \cite{EAGE.20.Werthmuller}.
 
@@
 
-\emph{Note to editors and reviewers: The data is NOT YET on Zenodo, we will
+\emph{Note to editors and reviewers: The data are NOT YET on Zenodo, we will
   upload there the final versions and link the right DOI for the final

### 2020-09-21 ManuscriptTime.tex
Correct NRMSD formula
start revision

@@
 %     twoside,
-%     revised,
+    revised,
   ]{geophysics}
@@
 from the target. We therefore show the normalized difference (ND) between the
-two 3D codes as a percentage, where $\mr{ND}=
-50|(\simpeg-\emg3d)/(\simpeg+\emg3d)|$. The ND between the two codes is below
-1\,\% everywhere except for early times.
+two 3D codes as a percentage, where
+\old{$\mr{ND}= 50|(\simpeg-\emg3d)/(\simpeg+\emg3d)|$.}
+\new{$\mr{ND}= 200|\simpeg-\emg3d|/(|\simpeg|+|\emg3d|)$.}
+The ND between the two codes is below 1\,\% everywhere except for early times.
 

### 2020-10-15 ManuscriptTime.tex
Update figures within revision

--- 2020-06-30 ManuscriptTime.tex
@@
 E-mail: \href{mailto:Dieter@Werthmuller.org}{Dieter@Werthmuller.org};\\
-\textbf{Keywords}: Fourier transforms, CSEM, 3D modelling.
 }
@@
 %
-Looking carefully at these three points results in much smaller computation
-times with speedup factors of ten or more over previous results. A
+Looking carefully at these three \new{key} points results in much smaller
+computation times with speedup factors of ten or more over previous results. A
 frequency-domain computation of time-domain data can therefore be competitive
@@
 %
+\\[1cm] \textbf{Keywords}: Controlled source electromagnetics (CSEM); Numerical
+modelling; Fourier analysis.
 \end{abstract}
@@
 disadvantages. One way to achieve this is to implement Maxwell's equation in
-both domains, as it is done, for instance, in \simpeg \citep{CAG.15.Cockett}.
-Another approach is to have Maxwell's equation only implemented in one domain,
-and use Fourier transforms to go to the other. However, this approach can be
-costly, as many frequencies over a wide range are required to go from the time
-domain to the frequency domain, or many times over a wide range for the
-opposite direction. We present a methodology which significantly reduces the
-amount and range of the required frequencies, translating into a significant
-reduction in computation time. The required computation grids for low
-frequencies (in our case in the order of 0.001\,Hz) and high frequencies
+both domains\old{, as it is done, for instance, in \simpeg (Cockett et al.,
+2015)}. Another approach is to have Maxwell's equation only implemented in one
+domain, and use Fourier transforms to go to the other. However, this approach
+can be costly, as many frequencies over a wide range are required to go from
+the time domain to the frequency domain, or many times over a wide range for
+the opposite direction. We present a methodology which significantly reduces
+the amount and range of the required frequencies, translating into a
+significant reduction in computation time. The required computation grids for
+low frequencies (in our case in the order of 0.001\,Hz) and high frequencies
 (around 100\,Hz) are hugely different. Low frequencies can be computed on a
@@
 standard multigrid approach fails for severe stretching or strong anisotropy,
-for which \emg3d has implemented known improvements such as line-relaxation and
-semicoarsening \citep{ECCFD.06.Jonsthovel}, with a non-standard Cholesky
-decomposition to speed up the computation \citep{GEO.08.Mulder}. One of the big
-advantages of the multigrid method is that it scales linearly (optimal) with
-the grid size in both CPU and RAM usage \citep{B.Springer.20.Mulder}. This
-makes it feasible to run even big models on standard computers, without the
-need for big clusters. All the examples in this article are run on a laptop
-with an i7-6600U\,CPU\,@\,2.6\,GHz (x4) and 16\,GB of memory, using Ubuntu
-18.04 and Python 3.7. For the semi-analytical computations of layered models we
-use the open-source (Apache License 2.0) code \empymod
-\citep{GEO.17.Werthmuller}.
+for which \old{\emg3d has implemented }known improvements such as
+line-relaxation and semicoarsening \citep{ECCFD.06.Jonsthovel} \new{are
+implemented}, with a non-standard Cholesky decomposition to speed up the
+computation \citep{GEO.08.Mulder}. One of the big advantages of the multigrid
+method is that it scales linearly (optimal) with the grid size in both CPU and
+RAM usage \citep{B.Springer.20.Mulder}. This makes it feasible to run even big
+models on standard computers, without the need for big clusters. All the
+examples in this article are run on a laptop with an i7-6600U\,CPU\,@\,2.6\,GHz
+(x4) and 16\,GB of memory, using Ubuntu 18.04 and Python 3.7. For the
+semi-analytical computations of layered models we use the open-source (Apache
+License 2.0) code \empymod \citep{GEO.17.Werthmuller}.
 
@@
 imaginary part in the frequency domain and the right plot the corresponding
-impulse response in the time domain. The red lines are the precise results
-obtained from \empymod. The blue circles indicate the actually computed
-responses and the black dots the frequencies which are interpolated or set to
-zero. The resulting time-domain response has a relative error of less than
-1\,\% everywhere except for very early times.
+impulse response in the time domain. The red lines are the \new{semi-analytical
+responses}\old{precise results obtained from \empymod}. The blue circles
+indicate the actually computed responses and the black dots the frequencies
+which are interpolated or set to zero. The resulting time-domain response has a
+relative error of less than 1\,\% everywhere except for very early times.
 
@@
 offset the lowest required frequency. So the more one can restrict the
-necessary offset range, the less frequencies are needed. Another important
-factor is how to interpolate and extrapolate from the computed frequencies to
-the frequencies required for the Fourier transform. For the FFTLog only
-extrapolation for higher and lower frequencies is required. The EM response
-becomes highly oscillatory for high frequencies, which makes it very hard to
-extrapolate the response to frequencies $f>f_\mathrm{max}$. However, if
-$f_\mathrm{max}$ is chosen judiciously, the importance of higher frequencies
-for the Fourier transform can be neglected and we can set those responses to
-zero. The extrapolation of frequencies $f<f_\mathrm{min}$ can be changed to an
-interpolation by assuming a zero imaginary response at zero frequency, and we
-then use PCHIP to interpolate the missing frequencies (as we work on a
-logarithmic scale we cannot choose 0\,Hz, but instead take $10^{-100}$\,Hz). In
-the case of the DLF method we also have to interpolate in between the computed
-frequencies, for which we found it better to use a cubic spline. As can be seen
-from Figures \ref{fig:GUI-FFTLog-lin} and \ref{fig:GUI-DLF-lin}, using the
-FFTLog or the DLF with the same actually computed frequencies results in very
-similar responses.
+necessary offset range, the \old{less}\new{fewer} frequencies are needed.
+Another important factor is how to interpolate and extrapolate from the
+computed frequencies to the frequencies required for the Fourier transform. For
+the FFTLog only extrapolation for higher and lower frequencies is required. The
+EM response becomes highly oscillatory for high frequencies, which makes it
+very hard to extrapolate the response to frequencies $f>f_\mathrm{max}$.
+However, if $f_\mathrm{max}$ is chosen judiciously, the importance of higher
+frequencies for the Fourier transform can be neglected and we can set those
+responses to zero. The extrapolation of frequencies $f<f_\mathrm{min}$ can be
+changed to an interpolation by assuming a zero imaginary response at zero
+frequency, and we then use PCHIP to interpolate the missing frequencies (as we
+work on a logarithmic scale we cannot choose 0\,Hz, but instead take
+$10^{-100}$\,Hz). In the case of the DLF method we also have to interpolate in
+between the computed frequencies, for which we found it better to use a cubic
+spline. As can be seen from Figures \ref{fig:GUI-FFTLog-lin} and
+\ref{fig:GUI-DLF-lin}, using the FFTLog or the DLF with the same actually
+computed frequencies results in very similar responses.
 
@@
 these inputs the adaptive gridding will search for the smallest possible number
-of cells which fulfils these criteria. The multigrid method implemented in
-\emg3d puts some constraints on the number of cells, of which the adaptive
-gridding takes care (the number of cells have to be powers of two multiplied by
-a low prime, e.g., $\{2,3,5\}\cdot2^n$).
+of cells which fulfils these criteria. The \new{implemented} multigrid method
+\old{implemented in \emg3d} puts some constraints on the number of cells, of
+which the adaptive gridding takes care (the number of cells have to be powers
+of two multiplied by a low prime, e.g., $\{2,3,5\}\cdot2^n$).
 
@@
   The red lines are the analytical solutions, the blue circles are the actually
-  computed responses with \emg3d, the black dots are the interpolated
+  computed responses\old{ with \emg3d}, the black dots are the interpolated
   responses, and the dashed black line the obtained time-domain response. The
@@
 %
-Most of the computed values stay below a relative error of 1\,\%, our
-chosen adaptive gridding only starts to generate considerable errors at higher
+Most of the computed values stay below a relative error of 1\,\%, our chosen
+adaptive gridding only starts to generate considerable errors at higher
 frequencies. Figure \ref{fig:fullspace}(b) shows the corresponding time-domain
-result, where the dashed black line is the result from \emg3d, on top of the
-red line which is the analytical result. The relative error is mostly below
-1\,\%, except for early times. However, for practical reasons that is more than
-enough. Figure~\ref{fig:fullspace-log} shows the same on a logarithmic scale,
-with times up to 10 seconds. It clearly shows that if later times are required,
-we would need to adjust our Fourier transform parameters.
+result, where the dashed black line is the result \old{from
+\emg3d}\new{obtained by transformation of the frequency-domain response}, on
+top of the red line which is the analytical result. The relative error is
+mostly below 1\,\%, except for early times. However, for practical reasons that
+is more than enough. Figure~\ref{fig:fullspace-log} shows the same on a
+logarithmic scale, with times up to 10 seconds. It clearly shows that if later
+times are required, we would need to adjust our Fourier transform parameters.
 %
@@
 
-For the comparison we use the open-source code \simpeg \citep{CAG.15.Cockett}.
-\simpeg is a framework for modelling and inversion of geophysical data such as
-gravity, magnetics, and CSEM. It has Maxwell's equations implemented in both
-the frequency and time domain. As such we can compare our result computed in
-the frequency domain followed by a Fourier transform to a result computed
-directly in the time domain. A principal difference between \simpeg and \emg3d
-is that the former has various direct solvers implemented, whereas the latter
-is an iterative multigrid solver. The 3D model is therefore a rather small
-example in order to be able to run it on our test machine, as the memory
-requirement by the direct solver would otherwise be too high. There are not
-many options out there of open-source time-domain 3D codes, \simpeg being the
-one we found to be suitable. A step-off response was chosen as this is the
-response currently implemented in \simpeg.
+For the comparison we use the open-source code \simpeg
+\citep{CAG.15.Cockett}\old{. \simpeg}\new{, which} is a framework for modelling
+and inversion of geophysical data such as gravity, magnetics, and CSEM. It has
+Maxwell's equations implemented in both the frequency and time domain. As such
+we can compare our result computed in the frequency domain followed by a
+Fourier transform to a result computed directly in the time domain. A principal
+difference between \simpeg and \emg3d is that the former has various direct
+solvers implemented, whereas the latter is an iterative multigrid solver. The
+3D model is therefore a rather small example in order to be able to run it on
+our test machine, as the memory requirement by the direct solver would
+otherwise be too high. There are not many options out there of open-source
+time-domain 3D codes, \simpeg being the one we found to be suitable. A step-off
+response was chosen as this is the response currently implemented in \new{the
+framework}\old{\simpeg}.
 
 The model was discretised with $100\times100\times100\,$m cells in the survey
-domain $D_\mr{s}$. For the \simpeg model, 14 cells in $x$-direction and 12
-cells in $y$- and $z$-directions were used on both sides with a stretching of
-1.3 for the total computation domain $D_\mr{c}$, which yields a mesh of
-\num{58344} cells. The time-steps start at 0.1\,s and are: $21\times0.01\,$s,
-$23\times0.03\,$s, $21\times0.1\,$s, $23\times0.3\,$s, covering exactly the
-desired range of 0.1--10\,s. For \emg3d, the mesh is generated
+domain $D_\mr{s}$. For the \old{\simpeg}\new{time-domain} model, 14 cells in
+$x$-direction and 12 cells in $y$- and $z$-directions were used on both sides
+with a stretching of 1.3 for the total computation domain $D_\mr{c}$, which
+yields a mesh of \num{58344} cells. The time-steps start at 0.1\,s and are:
+$21\times0.01\,$s, $23\times0.03\,$s, $21\times0.1\,$s, $23\times0.3\,$s,
+covering exactly the desired range of 0.1--10\,s. For
+\old{\emg3d}\new{frequency-domain model}, the mesh is generated
 frequency-dependent as in the previous examples, with a maximum stretching of
@@
 The results are shown in Figure~\ref{fig:3d-result}: In (a) the 1D background
-responses and the relative error using the semi-analytical result obtained from
-\empymod, and in (b) the responses including the target.
+responses and the relative error using the semi-analytical result\old{obtained
+from \empymod}, and in (b) the responses including the target.
 %
 \plot{3d-result}{width=\textwidth}{
-  Responses for the model outlined in Figure~\ref{fig:3d-model} using \simpeg
-  and \emg3d, and for the layered background also \empymod. In the lower plot
-  of (a) the relative error (\%) in comparison to \empymod is shown, and in (b)
-  the normalized difference (\%) between the two 3D codes. The results of these
-  plots are clipped for values smaller than 0.1\,\% and bigger than 10\,\%.
+  Responses for the model outlined in Figure~\ref{fig:3d-model} using
+  \new{time-domain}\old{\simpeg} and \new{frequency-domain
+  computations}\old{\emg3d}, and for the layered background also \new{the
+  semi-analytical result}\old{\empymod}. In the lower plot of (a) the relative
+  error (\%) \new{is shown} in comparison to \new{the semi-analytical
+  result}\old{\empymod is shown}, and in (b) the normalized difference (\%)
+  between the two 3D codes. The results of these plots are clipped for values
+  smaller than 0.1\,\% and bigger than 10\,\%.
 }
@@
 The background comparison shows that both 3D codes do an acceptable job having
-a relative error of a few percents at most; the result for \emg3d seems to be
-better at early times. We cannot compare the errors for the response that
-includes the target for lack of an analytical solution. The background model
-from \empymod is only included to show that there is a significant response
-from the target. We therefore show the normalized difference (ND) between the
-two 3D codes as a percentage, where $\mr{ND}=
-50|(\simpeg-\emg3d)/(\simpeg+\emg3d)|$. The ND between the two codes is below
-1\,\% everywhere except for early times.
+a relative error of a few percents at most; the result \new{obtained through
+transformation}\old{for \emg3d} seems to be better at early times. We cannot
+compare the errors for the response that includes the target for lack of an
+analytical solution. The \new{semi-analytical} background model \old{from
+\empymod }is only included to show that there is a significant response from
+the target. We therefore show the normalized difference (\old{ND}\new{NRMSD})
+between the two \old{3D codes}\new{responses $R_1$ and $R_2$} as a percentage,
+where \old{$\mr{ND}= 50|(\simpeg-\emg3d)/(\simpeg+\emg3d)|$.}
+\new{$\mr{NRMSD~(\%)} = 200|R_1 - R_2|/(|R_1| + |R_2|)$.} The
+\old{ND}\new{NRMSD} between the two codes is below 1\,\% everywhere except for
+early times.
 
@@
 taking also half of the stretching, which results then in a mesh of
-\num{466752} cells for \simpeg and meshes between \num{102400} and \num{614400}
-for \emg3d. Because the memory of our test machine was not large enough for the
-direct solver of \simpeg, the last model had to be run on a cluster. In this
-scenario, \emg3d was about eight times faster (single thread).
+\num{466752} cells \old{for \simpeg}\new{in the time domain} and meshes between
+\num{102400} and \num{614400} \old{for \emg3d}\new{in the frequency domain}.
+Because the memory of our test machine was not large enough for the direct
+solver of \simpeg, the last model had to be run on a cluster. In this scenario,
+\emg3d was about eight times faster (single thread).
 

### 2020-10-15 ManuscriptTime.tex
Within revision I

@@
 \multiplot{2}{GUI-FFTLog-log,GUI-DLF-lin}{width=\tcwidth}
-{(a) Same as Figure \ref{fig:GUI-FFTLog-lin}, but on a logarithmic scale.
-(b) Using DLF instead of FFTLog, but computing the same frequencies as for the
-FFTLog. The resulting time domain response has comparable accuracy (compare the
-error to the error in Figure \ref{fig:GUI-FFTLog-lin}).
+{(a) Same as Figure \ref{fig:GUI-FFTLog-lin}, but on a logarithmic scale. (b)
+Using DLF instead of FFTLog, but computing the same frequenc\old{ies}\new{y
+range} as for the FFTLog. The resulting time domain response has comparable
+accuracy (compare the error to the error in Figure \ref{fig:GUI-FFTLog-lin}).
 }
@@
 Twenty frequencies or less are usually sufficient for a wide range of offsets.
-We believe that our proposed improvements to the previously published methods
-make time-domain CSEM modelling with a frequency-domain code even more
-attractive than it was already before. We have demonstrated the idea of our
-Fourier transform method on CSEM data transformed from the frequency to the
-time domain. However, it could equally be applied to the transform from the
+\new{The values for lower frequencies can be interpolated using PCHIP knowing
+that the imaginary part goes to zero for zero frequency. The values for higher
+frequencies can be set to zero, as we can neglect their influence. And values
+for frequencies in-between the computed ones are best obtained with a spline
+interpolation. The actual transform can be made with either the DLF method or
+FFTLog, where the latter one requires usually much fewer frequencies to be
+interpolated.} We believe that our proposed improvements to the previously
+published methods make time-domain CSEM modelling with a frequency-domain code
+even more attractive than it was already before. We have demonstrated the idea
+of our Fourier transform method on CSEM data transformed from the frequency to
+the time domain. However, it could equally be applied to the transform from the
 time to the frequency domain and to other methods with similar

### 2020-10-15 notebooks/freqselect.py

@@ -319,13 +319,19 @@ class InteractiveFrequency(emg3d.utils.Fourier):
         else:
             self.axs[1].set_xlim([min(self.time), max(self.time)])
 
-    def print_suptitle(self):
+    def print_legend(self):
         """Update suptitle."""
-        plt.suptitle(
-            f"Offset = {np.squeeze(self.model['rec'][0])/1000} km;    "
-            f"No. freq. coarse: {self.freq_calc.size};    No. freq. full: "
-            f"{self.freq_req.size}  ({self.freq_req.min():.1e} $-$ "
-            f"{self.freq_req.max():.1e} Hz)")
+        # plt.suptitle(
+        #     f"Offset = {np.squeeze(self.model['rec'][0])/1000} km;    "
+        #     f"No. freq. coarse: {self.freq_calc.size};    No. freq. full: "
+        #     f"{self.freq_req.size}  ({self.freq_req.min():.1e} $-$ "
+        #     f"{self.freq_req.max():.1e} Hz)")
+        self.clear_handle(['f_legend', ])
+        self.h_f_legend = self.axs[0].legend(
+            handles=[self.h_f_inti, self.h_f_int],
+            labels=[f"req. ({self.freq_req.size})",
+                    f"comp. ({self.freq_calc.size})"]
+        )
 
     def plot_base_model(self):
         """Update smooth, 'correct' model."""
@@ -375,8 +381,8 @@ class InteractiveFrequency(emg3d.utils.Fourier):
         self.h_t_int, = self.axs[1].plot(self.time, t_int, 'k--')
         self.h_t_inte, = self.axs[3].plot(self.time, t_error, 'k.')
 
-        # Update suptitle
-        self.print_suptitle()
+        # Update legend
+        self.print_legend()
 
     # Interactive routines
     def update_off(self, off):
@@ -454,7 +460,7 @@ class InteractiveFrequency(emg3d.utils.Fourier):
         if linlog == 'log':
             sym_dec = 10  # Number of decades to show on symlog
             lty = int(max(np.log10(abs(self.reim(self.f_dense))))-sym_dec)
-            self.axs[0].set_yscale('symlog', linthresh=10**lty, linscaley=0.7)
+            self.axs[0].set_yscale('symlog', linthresh=10**lty, linscale=0.7)
 
             # Remove the zero line becouse of the overlapping ticklabels.
             nticks = len(self.axs[0].get_yticks())//2

### 2020-10-23 ManuscriptTime.tex
Add small note

@@
 frequency-domain approach appears to be attractive.\guillemotright} Their
+\todo{mention the four methods}
 approach is to minimize the computation time by having, in addition to the just

### 2020-11-24 ManuscriptTime.tex
Change from Geophysics to GJI template

@@
-\documentclass[
-%     paper,
-    manuscript,
-%     twocolumn,
-%     twoside,
-    revised,
-  ]{geophysics}
+% \documentclass[onecolumn,extra,referee,camera]{gji}
+% \documentclass[onecolumn,extra,camera]{gji}
+\documentclass[twocolumn,extra,camera]{gji}
+\usepackage{timet}
 
-% Remove once done
-\usepackage[textsize=footnotesize, textwidth=2.8cm]{todonotes}
-
-% Additional packages to geophysics.cls
-\usepackage[UKenglish]{babel}
+% Additional packages
+% \usepackage[UKenglish]{babel}
 \usepackage[utf8]{inputenc}
@@
 \usepackage[T1]{fontenc}
-
 \usepackage{amssymb, amsmath, amsfonts}
-\usepackage{upquote}
-\usepackage[strings]{underscore}
-\usepackage{siunitx}                % SI conform commands (eg \num)
-\usepackage{tabularx}
-\usepackage{colortbl}
 \usepackage{booktabs}
 \usepackage{xspace}
+\usepackage{graphicx}
+\usepackage{subfig}
+\usepackage[draft, addedmarkup=bf]{changes}
+% \usepackage[final]{changes}
+\usepackage[pdftex, final, allcolors=blue, colorlinks=true]{hyperref}
 
-\usepackage[pdftex, final]{hyperref}
-% \usepackage[pdftex, hidelinks]{hyperref}
-\hypersetup{allcolors=blue, allbordercolors={0 0 .5}, colorlinks=true}
+% workaround siunitx clashes unfortunately with gji
+% https://tex.stackexchange.com/a/412886
+\makeatletter
+\let\zz@tabular\@tabular
+\let\zzendtabular\endtabular
+\let\zz@xtabularcr\@xtabularcr
+\let\zz@tabclassz\@tabclassz
+\let\zz@tabclassiv \@tabclassiv 
+\let\zz@tabarray\@tabarray
+\makeatother
 
-% Figures
-\DeclareGraphicsExtensions{.pdf,.png,.jpg}
-\renewcommand{\figdir}{./figures}
-\ifthenelse{\boolean{@twoc}}{%
-  \newcommand{\cwidth}{250pt}}{%
-  \newcommand{\cwidth}{240pt}}
+% Change default colour of «changes» package
+\definecolor{myred}{rgb}{.8,0,0}
+\definecolor{mygreen}{rgb}{0,.5,0}
+\setaddedmarkup{\color{myred}{#1}}
+\setdeletedmarkup{\color{mygreen}{\itshape \sout{#1}}}
 
-\ifthenelse{\boolean{@twoc}}{%
-  \newcommand{\acwidth}{250pt}%
-  \newcommand{\tcwidth}{250pt}}{%
-  \newcommand{\acwidth}{240pt}%
-  \newcommand{\tcwidth}{312pt}}
+\let\tablenum\relax
+\usepackage{siunitx}
+\usepackage{tabularx}
+
+% To fix clash of gji with hyperref
+% https://tex.stackexchange.com/a/384969
+\AtBeginDocument{
+  \label{CorrectFirstPageLabel}
+  \def\fpage{\pageref{CorrectFirstPageLabel}}
+}
 
@@
 
+\title[Time-domain CSEM]{Fast time-domain electromagnetic modelling in the frequency domain}
+
+\author[D. Werthmüller \emph{et al.}]
+  {\Large Dieter Werthmüller$^1$,        % orcid: 0000-0002-8575-2484
+   Wim A. Mulder$^{1,2}$ and          % orcid: 0000-0001-7020-9297
+   Evert C. Slob$^1$              % orcid: 0000-0002-4529-1134
+   \\
+   \footnotesize
+  $^1$ TU Delft, Building 23, Stevinweg 1 / PO-box 5048, 2628 CN Delft (NL).
+  E-mail: \href{mailto:Dieter@Werthmuller.org}{Dieter@Werthmuller.org}\\[-.3em]
+   \footnotesize
+  $^2$ Shell Global Solutions International BV, Grasweg 31, 1031 HW Amsterdam(NL)
+  }
+
+\date{Received 202Y MMMM DD; in original form 2020 June 30}
+\pagerange{\pageref{firstpage}--\pageref{lastpage}}
+\volume{VVV}
+\pubyear{202Y}
+
+% Authors with AMS fonts and mssymb.tex can comment out the following
+% line to get the correct symbol for Geophysical Journal International.
+\let\leqslant=\leq
+
+\newtheorem{theorem}{Theorem}[section]
+
 \begin{document}
 
-\title{Fast time-domain electromagnetic modelling in the frequency domain}
+\label{firstpage}
 
-\renewcommand{\thefootnote}{\fnsymbol{footnote}}
+% workaround siunitx clashes unfortunately with gji
+% https://tex.stackexchange.com/a/412886
+{\makeatletter
+\let\@tabular\zz@tabular
+\let\endtabular\zzendtabular
+\let\@xtabularcr\zz@xtabularcr
+\let\@tabclassz\zz@tabclassz
+\let\@tabclassiv \zz@tabclassiv 
+\let\@tabarray\zz@tabarray
+\maketitle
+}
+% \maketitle
 
-\ms{}  % ARTICLE-ID
-
-\address{
-\footnotemark[1]TU Delft,
-Building 23,
-Stevinweg 1 / PO-box 5048,
-2628 CN Delft;\\
-\footnotemark[2]Shell Global Solutions International BV,
-Grasweg 31, 1031 HW Amsterdam;\\
-E-mail: \href{mailto:Dieter@Werthmuller.org}{Dieter@Werthmuller.org};\\
-}
-
-\author{%
-Dieter Werthmüller\footnotemark[1], %            orcid: 0000-0002-8575-2484
-Wim A.\ Mulder\footnotemark[1]\footnotemark[2], % orcid: 0000-0001-7020-9297
-and Evert C.\ Slob\footnotemark[1] %             orcid: 0000-0002-4529-1134
-}
-
-\footer{}
-\lefthead{Werthmüller et al.}
-\righthead{Time-domain CSEM}
-
-\maketitle
-
-%%fakesection ===    ABSTRACT    ===
-\begin{abstract}
+\begin{summary}
 %
@@
 %
-Looking carefully at these three \new{key} points results in much smaller
+Looking carefully at these three \added{key} points results in much smaller
 computation times with speedup factors of ten or more over previous results. A
@@
 %
-\\[1cm] \textbf{Keywords}: Controlled source electromagnetics (CSEM); Numerical
-modelling; Fourier analysis.
-\end{abstract}
+\end{summary}
 
-\section{Introduction}
+% https://academic.oup.com/DocumentLibrary/GJI/gji%20keyword%20list%20updated2017.pdf
+\begin{keywords}
+  Controlled source electromagnetics (CSEM); Numerical modelling; Fourier analysis.
+\end{keywords}
+
+\section{Introduction}  % % % % % % % % % % % % % % % % % % % % % % % % % % % %
 
@@
 
-\section{Motivation}
+\section{Motivation}  % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
 
@@
 disadvantages. One way to achieve this is to implement Maxwell's equation in
-both domains\old{, as it is done, for instance, in \simpeg (Cockett et al.,
+both domains\deleted{, as it is done, for instance, in \simpeg (Cockett et al.,
 2015)}. Another approach is to have Maxwell's equation only implemented in one
@@
 
-\section{Methodology}
+\section{Methodology} % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
 
@@
 standard multigrid approach fails for severe stretching or strong anisotropy,
-for which \old{\emg3d has implemented }known improvements such as
-line-relaxation and semicoarsening \citep{ECCFD.06.Jonsthovel} \new{are
+for which \deleted{\emg3d has implemented }known improvements such as
+line-relaxation and semicoarsening \citep{ECCFD.06.Jonsthovel} \added{are
 implemented}, with a non-standard Cholesky decomposition to speed up the
@@
 
-\section{Frequency selection}
+\section{Frequency selection} % % % % % % % % % % % % % % % % % % % % % % % % %
 
@@
 %
-\plot{GUI-FFTLog-lin}{width=\tcwidth}{
-  Interactive frequency selection for a user-provided layered model (the shown
-  model parameters are described in the text). This example shows the impulse
-  response at an offset of 5\,km, for which it uses FFTLog from 0.001\,Hz to
-  10\,Hz with five frequencies per decade. Note that the error is clipped for
-  values smaller than 0.01\,\% and bigger than 100\,\%.
-}
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{figures/GUI-FFTLog-lin}
+  \caption{Interactive frequency selection for a user-provided layered model
+    (the shown model parameters are described in the text). This example shows
+    the impulse response at an offset of 5\,km, for which it uses FFTLog from
+    0.001\,Hz to 10\,Hz with five frequencies per decade. Note that the error
+    is clipped for values smaller than 0.01\,\% and bigger than 100\,\%.}
+  \label{fig:GUI-FFTLog-lin}
+\end{figure}
 %
+
 The example model is a marine scenario with 1\,km water depth of resistivity
@@
 imaginary part in the frequency domain and the right plot the corresponding
-impulse response in the time domain. The red lines are the \new{semi-analytical
-responses}\old{precise results obtained from \empymod}. The blue circles
-indicate the actually computed responses and the black dots the frequencies
-which are interpolated or set to zero. The resulting time-domain response has a
-relative error of less than 1\,\% everywhere except for very early times.
+impulse response in the time domain. The red lines are the
+\replaced{semi-analytical responses}{precise results obtained from \empymod}.
+The blue circles indicate the actually computed responses and the black dots
+the frequencies which are interpolated or set to zero. The resulting
+time-domain response has a relative error of less than 1\,\% everywhere except
+for very early times.
 
@@
 %
-\multiplot{2}{GUI-FFTLog-log,GUI-DLF-lin}{width=\tcwidth}
-{(a) Same as Figure \ref{fig:GUI-FFTLog-lin}, but on a logarithmic scale. (b)
-Using DLF instead of FFTLog, but computing the same frequenc\old{ies}\new{y
-range} as for the FFTLog. The resulting time domain response has comparable
-accuracy (compare the error to the error in Figure \ref{fig:GUI-FFTLog-lin}).
-}
+\begin{figure*}
+  \centering
+  \subfloat[]{{\includegraphics[width=.5\linewidth]{figures/GUI-FFTLog-log}}%
+  \label{fig:GUI-FFTLog-log}}\hfill
+  \subfloat[]{{\includegraphics[width=.5\linewidth]{figures/GUI-DLF-lin}}%
+  \label{fig:GUI-DLF-lin}}\\
+  \caption{(a) Same as Figure \ref{fig:GUI-FFTLog-lin}, but on a logarithmic
+  scale. (b) Using DLF instead of FFTLog, but computing the same
+  frequenc\replaced{y range}{ies} as for the FFTLog. The resulting time domain
+  response has comparable accuracy (compare the error to the error in Figure
+  \ref{fig:GUI-FFTLog-lin}).}
+  \label{fig:GUI-log-lin}
+\end{figure*}
 %
@@
 %
-\plot{multi-offset}{width=\tcwidth}{
-  Normalized (a) frequency- and (b) time-domain responses for offsets
+\begin{figure}
+  \centering
+  \includegraphics[width=.8\linewidth]{figures/multi-offset}
+  \caption{Normalized (a) frequency- and (b) time-domain responses for offsets
     $r=1.5,3,6$, and 12\,km using the parameters defined in Figure
@@
     responses, the black dots are the responses which are set to zero or
-  interpolated. The black curves are the analytical responses.
-}
+    interpolated. The black curves are the analytical responses.}
+  \label{fig:multi-offset}
+\end{figure}
 %
@@
 offset the lowest required frequency. So the more one can restrict the
-necessary offset range, the \old{less}\new{fewer} frequencies are needed.
+necessary offset range, the \replaced{fewer}{less} frequencies are needed.
 Another important factor is how to interpolate and extrapolate from the
@@
 
-\section{Gridding}
+\section{Gridding}  % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
 
@@
 these inputs the adaptive gridding will search for the smallest possible number
-of cells which fulfils these criteria. The \new{implemented} multigrid method
-\old{implemented in \emg3d} puts some constraints on the number of cells, of
+of cells which fulfils these criteria. The \added{implemented} multigrid method
+\deleted{implemented in \emg3d} puts some constraints on the number of cells, of
 which the adaptive gridding takes care (the number of cells have to be powers
@@
 
-\section{Numerical examples}
+\section{Numerical examples}  % % % % % % % % % % % % % % % % % % % % % % % % %
 
-\subsection{Homogeneous space}
+\subsection{Homogeneous space}  % % % % % % % % % % % % % % % % % % % % % % % %
 
@@
 %
-\tabl[btp]{timefull}{Run times per frequency for the homogeneous space example,
-  with the corresponding number of cells and minimum cell width as well as the
-  stretching factor in the computation domain; $\alpha_\mr{s}=1$ everywhere.}{
+\begin{table}
+\begin{minipage}{\columnwidth}
   \centering
+  \caption{Run times per frequency for the homogeneous space example, with the
+    corresponding number of cells and minimum cell width as well as the
+    stretching factor in the computation domain; $\alpha_\mr{s}=1$ everywhere.}
+  \label{tbl:timefull}
   \begin{tabular}{S[table-format=2.4]rrrr}
@@
     \bottomrule
-\end{tabular}}%
+  \end{tabular}
+\end{minipage}
+\end{table}
 %
@@
 %
-\plot*{fullspace}{width=\textwidth}{
-  (a) Frequency- and (b) time-domain results for the homogeneous space model.
-  The red lines are the analytical solutions, the blue circles are the actually
-  computed responses\old{ with \emg3d}, the black dots are the interpolated
-  responses, and the dashed black line the obtained time-domain response. The
-  errors are clipped for values smaller than 0.1\,\% and bigger than 100\,\%.
-}
+\begin{figure*}
+  \centering
+  \includegraphics[width=\linewidth]{figures/fullspace}
+  \caption{(a) Frequency- and (b) time-domain results for the homogeneous space
+    model. The red lines are the analytical solutions, the blue circles are the
+    actually computed responses\deleted{ with \emg3d}, the black dots are the
+    interpolated responses, and the dashed black line the obtained time-domain
+    response. The errors are clipped for values smaller than 0.1\,\% and bigger
+    than 100\,\%.}
+  \label{fig:fullspace}
+\end{figure*}
 %
@@
 frequencies. Figure \ref{fig:fullspace}(b) shows the corresponding time-domain
-result, where the dashed black line is the result \old{from
-\emg3d}\new{obtained by transformation of the frequency-domain response}, on
+result, where the dashed black line is the result \replaced{obtained by
+transformation of the frequency-domain response}{from \emg3d}, on
 top of the red line which is the analytical result. The relative error is
@@
 %
-\plot{fullspace-log}{width=\cwidth}{
-  Same as in Figure~\ref{fig:fullspace} (b), but on a logarithmic scale. To
-  improve later times we would have to compute lower frequencies; to improve
-  earlier times we would have to compute more frequencies per decade to get a
-  better resolution.
-}
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{figures/fullspace-log}
+  \caption{Same as in Figure~\ref{fig:fullspace} (b), but on a logarithmic
+    scale. To improve later times we would have to compute lower frequencies;
+    to improve earlier times we would have to compute more frequencies per
+    decade to get a better resolution.}
+  \label{fig:fullspace-log}
+\end{figure}
 %
@@
 \emg3d is written in Python/Numba (Numba is a just-in-time compiler for Python
-code, \citealp{Numba}).
+code, \cite{Numba}).
 
-\subsection{1D Model}
+\subsection{1D Model} % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
 
@@
 %
-\tabl[btp]{timemarine}{Run times per frequency for the marine 1D example,
-  with the corresponding number of cells and minimum cell width as well as the
-  stretching factor in the computation domain; $\alpha_\mr{s}=1$ everywhere.}{
+\begin{table}
+\begin{minipage}{\columnwidth}
   \centering
+  \caption{Run times per frequency for the marine 1D example, with the
+    corresponding number of cells and minimum cell width as well as the
+    stretching factor in the computation domain; $\alpha_\mr{s}=1$ everywhere.}
+  \label{tbl:timemarine}
   \begin{tabular}{S[table-format=2.5]rrrr}
@@
 %
-\plot*{marine}{width=\textwidth}{
-  Electric inline response at an offset of 5\,km for a shallow marine, layered
-  scenario. (a) Frequency-domain, where the blue circles are computed responses
-  and the black dots are interpolated responses or responses set to zero. (b)
-  Time-domain response.
-}
+\begin{figure*}
+  \centering
+  \includegraphics[width=.8\linewidth]{figures/marine}
+  \caption{Electric inline response at an offset of 5\,km for a shallow marine,
+    layered scenario. (a) Frequency-domain, where the blue circles are computed
+    responses and the black dots are interpolated responses or responses set to
+    zero. (b) Time-domain response.}
+  \label{fig:marine}
+\end{figure*}
 %
@@
 %
-\plot{marine-multioffset}{width=\cwidth}{
-  Time-domain responses for offsets of 3, 5, and 7\,km for the same model as
-  shown in Figure~\ref{fig:marine}.
-}
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{figures/marine-multioffset}
+  \caption{Time-domain responses for offsets of 3, 5, and 7\,km for the same
+    model as shown in Figure~\ref{fig:marine}.}
+  \label{fig:marine-multioffset}
+\end{figure}
 %
 
-\subsection{Horizontal extent of the computation domain}
+\subsection{Horizontal extent of the computation domain}  % % % % % % % % % % %
 
@@
 %
-\plot*{marine-wrong-x-y}{width=\textwidth}{
-  Same model as used for Figure~\ref{fig:marine}, but with the horizontal
-  boundaries not far enough. Although the resulting time-domain result looks
-  plausible, the comparison with the 1D result shows that it significantly
-  overestimates the amplitude of the airwave.
-}
+\begin{figure*}
+  \centering
+  \includegraphics[width=.8\linewidth]{figures/marine-wrong-x-y}
+  \caption{Same model as used for Figure~\ref{fig:marine}, but with the
+    horizontal boundaries not far enough. Although the resulting time-domain
+    result looks plausible, the comparison with the 1D result shows that it
+    significantly overestimates the amplitude of the airwave.}
+  \label{fig:marine-wrong-x-y}
+\end{figure*}
 %
@@
 
-\subsection{3D Model}
+\subsection{3D Model} % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
 
@@
 %
-\plot{3d-model}{width=\cwidth}{
-  Three-dimensional block embedded in the lower of two halfspaces. The 100-m
-  long, $x$-directed dipole source is located 50\,m above the seafloor at the
-  origin, and the receiver is on the seafloor at an offset of 2\,km.
-}
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{figures/3d-model}
+  \caption{Three-dimensional block embedded in the lower of two halfspaces. The
+    100-m long, $x$-directed dipole source is located 50\,m above the seafloor
+    at the origin, and the receiver is on the seafloor at an offset of 2\,km.}
+  \label{fig:3d-model}
+\end{figure}
 %
@@
 For the comparison we use the open-source code \simpeg
-\citep{CAG.15.Cockett}\old{. \simpeg}\new{, which} is a framework for modelling
+\citep{CAG.15.Cockett}\replaced{, which}{. \simpeg} is a framework for modelling
 and inversion of geophysical data such as gravity, magnetics, and CSEM. It has
@@
 time-domain 3D codes, \simpeg being the one we found to be suitable. A step-off
-response was chosen as this is the response currently implemented in \new{the
-framework}\old{\simpeg}.
+response was chosen as this is the response currently implemented in
+\replaced{the framework}{\simpeg}.
 
 The model was discretised with $100\times100\times100\,$m cells in the survey
-domain $D_\mr{s}$. For the \old{\simpeg}\new{time-domain} model, 14 cells in
+domain $D_\mr{s}$. For the \replaced{time-domain}{\simpeg} model, 14 cells in
 $x$-direction and 12 cells in $y$- and $z$-directions were used on both sides
@@
 covering exactly the desired range of 0.1--10\,s. For
-\old{\emg3d}\new{frequency-domain model}, the mesh is generated
+\replaced{the frequency-domain model}{\emg3d}, the mesh is generated
 frequency-dependent as in the previous examples, with a maximum stretching of
@@
 The results are shown in Figure~\ref{fig:3d-result}: In (a) the 1D background
-responses and the relative error using the semi-analytical result\old{obtained
-from \empymod}, and in (b) the responses including the target.
+responses and the relative error using the semi-analytical
+result\deleted{obtained from \empymod}, and in (b) the responses including the
+target.
 %
-\plot{3d-result}{width=\textwidth}{
-  Responses for the model outlined in Figure~\ref{fig:3d-model} using
-  \new{time-domain}\old{\simpeg} and \new{frequency-domain
-  computations}\old{\emg3d}, and for the layered background also \new{the
-  semi-analytical result}\old{\empymod}. In the lower plot of (a) the relative
-  error (\%) \new{is shown} in comparison to \new{the semi-analytical
-  result}\old{\empymod is shown}, and in (b) the normalized difference (\%)
+\begin{figure*}
+  \centering
+  \includegraphics[width=.8\linewidth]{figures/3d-result}
+  \caption{Responses for the model outlined in Figure~\ref{fig:3d-model} using
+    \replaced{time-domain}{\simpeg} and \replaced{frequency-domain
+    computations}{\emg3d}, and for the layered background also \replaced{the
+    semi-analytical result}{\empymod}. In the lower plot of (a) the relative
+    error (\%) \added{is shown} in comparison to \replaced{the semi-analytical
+    result}{\empymod is shown}, and in (b) the normalized difference (\%)
     between the two 3D codes. The results of these plots are clipped for values
-  smaller than 0.1\,\% and bigger than 10\,\%.
-}
+    smaller than 0.1\,\% and bigger than 10\,\%.}
+  \label{fig:3d-result}
+\end{figure*}
 %
 The background comparison shows that both 3D codes do an acceptable job having
-a relative error of a few percents at most; the result \new{obtained through
-transformation}\old{for \emg3d} seems to be better at early times. We cannot
-compare the errors for the response that includes the target for lack of an
-analytical solution. The \new{semi-analytical} background model \old{from
-\empymod }is only included to show that there is a significant response from
-the target. We therefore show the normalized difference (\old{ND}\new{NRMSD})
-between the two \old{3D codes}\new{responses $R_1$ and $R_2$} as a percentage,
-where \old{$\mr{ND}= 50|(\simpeg-\emg3d)/(\simpeg+\emg3d)|$.}
-\new{$\mr{NRMSD~(\%)} = 200|R_1 - R_2|/(|R_1| + |R_2|)$.} The
-\old{ND}\new{NRMSD} between the two codes is below 1\,\% everywhere except for
+a relative error of a few percents at most; the result \replaced{obtained
+through transformation}{for \emg3d} seems to be better at early times. We
+cannot compare the errors for the response that includes the target for lack of
+an analytical solution. The \added{semi-analytical} background model
+\deleted{from \empymod }is only included to show that there is a significant
+response from the target. We therefore show the normalized difference
+(\replaced{NRMSD}{ND}) between the two \replaced{responses $R_1$ and $R_2$}{3D
+codes} as a percentage, where \replaced{$\mr{NRMSD~(\%)} = 200|R_1 -
+R_2|/(|R_1| + |R_2|)$.}{$\mr{ND}= 50|(\simpeg-\emg3d)/(\simpeg+\emg3d)|$.} The
+\replaced{NRMSD}{ND} between the two codes is below 1\,\% everywhere except for
 early times.
@@
 taking also half of the stretching, which results then in a mesh of
-\num{466752} cells \old{for \simpeg}\new{in the time domain} and meshes between
-\num{102400} and \num{614400} \old{for \emg3d}\new{in the frequency domain}.
-Because the memory of our test machine was not large enough for the direct
-solver of \simpeg, the last model had to be run on a cluster. In this scenario,
-\emg3d was about eight times faster (single thread).
+\num{466752} cells \replaced{in the time domain}{for \simpeg} and meshes
+between \num{102400} and \num{614400} \replaced{in the frequency domain}{for
+\emg3d}. Because the memory of our test machine was not large enough for the
+direct solver of \simpeg, the last model had to be run on a cluster. In this
+scenario, \emg3d was about eight times faster (single thread).
 
-\section{Conclusions}
+\section{Conclusions} % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
 
@@
 Twenty frequencies or less are usually sufficient for a wide range of offsets.
-\new{The values for lower frequencies can be interpolated using PCHIP knowing
+\added{The values for lower frequencies can be interpolated using PCHIP knowing
 that the imaginary part goes to zero for zero frequency. The values for higher
@@
 
-\section{Acknowledgment}
-
-This research was conducted within the Gitaro.JIM project funded through
-MarTERA as part of Horizon 2020, a funding scheme of the European Research Area
-(ERA-NET Cofund, \href{https://www.martera.eu}{martera.eu}).
-An early idea of this manuscript was presented at \cite{EAGE.20.Werthmuller}.
-
-\section{Data Availability Statement}
+\section{Data Availability Statement} % % % % % % % % % % % % % % % % % % % % %
 
@@
 
+\begin{acknowledgments} % % % % % % % % % % % % % % % % % % % % % % % % % % % %
+This research was conducted within the Gitaro.JIM project funded through
+MarTERA as part of Horizon 2020, a funding scheme of the European Research Area
+(ERA-NET Cofund, \href{https://www.martera.eu}{martera.eu}).
+An early idea of this manuscript was presented at \cite{EAGE.20.Werthmuller}.
+\end{acknowledgments}
+
 % REFERENCES
-
-\bibliographystyle{Refs}          % Modified SEG bibliography style.
+\bibliographystyle{gji}
 \bibliography{Refs}
 
+\label{lastpage}
+
 \end{document}

### 2020-11-24 ManuscriptTime.tex
Make figwidth fit all formats

@@
 \usepackage{subfig}
+\usepackage{ifthen}
 \usepackage[draft, addedmarkup=bf]{changes}
@@
 
+% Figures
+% To print widths put this in text:
+% \the\textwidth
+% \the\columnwidth
+%
+% Widths:   textwidth  columnwidth  tablewidth
+%       1-c   504pt        252pt       504pt
+%       2-c   504pt        240pt       240pt
+%       ref   468pt        234pt       468pt
+%
+\ifthenelse{\boolean{GJ@referee}}{%
+  \newcommand{\cwidth}{234pt}%
+  \newcommand{\fwidth}{468pt}%
+  \newcommand{\twidth}{468pt}}{%
+  \newcommand{\fwidth}{504pt}%
+  \ifthenelse{\boolean{@twocolumn}}{%
+    \newcommand{\cwidth}{240pt}%
+    \newcommand{\twidth}{240pt}}{%
+    \newcommand{\cwidth}{252pt}%
+    \newcommand{\twidth}{504pt}%
+  }
+}
+\graphicspath{{./figures/}}
+
 % Change default colour of «changes» package
@@
 \definecolor{mygreen}{rgb}{0,.5,0}
+% \setaddedmarkup{\color{myred}{#1}}
 \setaddedmarkup{\color{myred}{#1}}
-\setdeletedmarkup{\color{mygreen}{\itshape \sout{#1}}}
+\setdeletedmarkup{\color{mygreen}{\sout{#1}}}
 
@@
   \centering
-  \includegraphics[width=\linewidth]{figures/GUI-FFTLog-lin}
+  \includegraphics[width=\cwidth]{GUI-FFTLog-lin}
   \caption{Interactive frequency selection for a user-provided layered model
@@
   \centering
-  \subfloat[]{{\includegraphics[width=.5\linewidth]{figures/GUI-FFTLog-log}}%
+  \subfloat[]{{\includegraphics[width=\cwidth]{GUI-FFTLog-log}}%
   \label{fig:GUI-FFTLog-log}}\hfill
-  \subfloat[]{{\includegraphics[width=.5\linewidth]{figures/GUI-DLF-lin}}%
+  \subfloat[]{{\includegraphics[width=\cwidth]{GUI-DLF-lin}}%
   \label{fig:GUI-DLF-lin}}\\
@@
   \centering
-  \includegraphics[width=.8\linewidth]{figures/multi-offset}
+  \includegraphics[width=\cwidth]{multi-offset}
   \caption{Normalized (a) frequency- and (b) time-domain responses for offsets
@@
 \begin{table}
-\begin{minipage}{\columnwidth}
+\begin{minipage}{\twidth}
   \centering
@@
     %
-    {freq.} & time & nx$\times$ny$\times$nz & $\Delta_\mr{min}$ & $\alpha_\mr{c}$ \\
+    {Frequency} & Time & nx$\times$ny$\times$nz & $\Delta_\mr{min}$ & $\alpha_\mr{c}$ \\
     {(Hz)}  & (s) &  & (m) & \\
@@
   \centering
-  \includegraphics[width=\linewidth]{figures/fullspace}
+  \includegraphics[width=\fwidth]{fullspace}
   \caption{(a) Frequency- and (b) time-domain results for the homogeneous space
@@
   \centering
-  \includegraphics[width=\linewidth]{figures/fullspace-log}
+  \includegraphics[width=\cwidth]{fullspace-log}
   \caption{Same as in Figure~\ref{fig:fullspace} (b), but on a logarithmic
@@
     %
-    {freq.} & time & nx$\times$ny$\times$nz & $\Delta_\mr{min}$ & $\alpha_\mr{c}$ \\
+    {Frequency} & time & nx$\times$ny$\times$nz & $\Delta_\mr{min}$ & $\alpha_\mr{c}$ \\
     {(Hz)}  & (s) &  & (m) & \\
@@
   \centering
-  \includegraphics[width=.8\linewidth]{figures/marine}
+  \includegraphics[width=\fwidth]{marine}
   \caption{Electric inline response at an offset of 5\,km for a shallow marine,
@@
   \centering
-  \includegraphics[width=\linewidth]{figures/marine-multioffset}
+  \includegraphics[width=\cwidth]{marine-multioffset}
   \caption{Time-domain responses for offsets of 3, 5, and 7\,km for the same
@@
   \centering
-  \includegraphics[width=.8\linewidth]{figures/marine-wrong-x-y}
+  \includegraphics[width=\fwidth]{marine-wrong-x-y}
   \caption{Same model as used for Figure~\ref{fig:marine}, but with the
@@
   \centering
-  \includegraphics[width=\linewidth]{figures/3d-model}
+  \includegraphics[width=\cwidth]{3d-model}
   \caption{Three-dimensional block embedded in the lower of two halfspaces. The
@@
   \centering
-  \includegraphics[width=.8\linewidth]{figures/3d-result}
+  \includegraphics[width=\fwidth]{3d-result}
   \caption{Responses for the model outlined in Figure~\ref{fig:3d-model} using

### 2020-11-25 notebooks/freqselect.py
Finish Notebook 01

@@ -329,8 +329,8 @@ class InteractiveFrequency(emg3d.utils.Fourier):
         self.clear_handle(['f_legend', ])
         self.h_f_legend = self.axs[0].legend(
             handles=[self.h_f_inti, self.h_f_int],
-            labels=[f"req. ({self.freq_req.size})",
-                    f"comp. ({self.freq_calc.size})"]
+            labels=[f"required ({self.freq_req.size})",
+                    f"computed ({self.freq_calc.size})"]
         )
 
     def plot_base_model(self):

### 2020-11-26 ManuscriptTime.tex
Within revision

@@
-% \documentclass[onecolumn,extra,referee,camera]{gji}
+\documentclass[onecolumn,extra,referee,camera]{gji}
 % \documentclass[onecolumn,extra,camera]{gji}
-\documentclass[twocolumn,extra,camera]{gji}
+% \documentclass[twocolumn,extra,camera]{gji}
 \usepackage{timet}
@@
 \usepackage{graphicx}
-\usepackage{subfig}
+\usepackage[position=top]{subfig}
 \usepackage{ifthen}
@@
 %
+\newlength{\fwidth}
+\newlength{\cwidth}
+\newlength{\twidth}
 \ifthenelse{\boolean{GJ@referee}}{%
-  \newcommand{\cwidth}{234pt}%
-  \newcommand{\fwidth}{468pt}%
-  \newcommand{\twidth}{468pt}}{%
-  \newcommand{\fwidth}{504pt}%
+  \setlength{\cwidth}{234pt}%
+  \setlength{\fwidth}{468pt}%
+  \setlength{\twidth}{468pt}}{%
+  \setlength{\fwidth}{504pt}%
   \ifthenelse{\boolean{@twocolumn}}{%
-    \newcommand{\cwidth}{240pt}%
-    \newcommand{\twidth}{240pt}}{%
-    \newcommand{\cwidth}{252pt}%
-    \newcommand{\twidth}{504pt}%
+    \setlength{\cwidth}{240pt}%
+    \setlength{\twidth}{240pt}}{%
+    \setlength{\cwidth}{252pt}%
+    \setlength{\twidth}{504pt}%
   }
@@
 % Change default colour of «changes» package
-\definecolor{myred}{rgb}{.8,0,0}
-\definecolor{mygreen}{rgb}{0,.5,0}
-% \setaddedmarkup{\color{myred}{#1}}
-\setaddedmarkup{\color{myred}{#1}}
-\setdeletedmarkup{\color{mygreen}{\sout{#1}}}
+\definecolor{alizarin}{rgb}{0.82, 0.1, 0.26}
+\definecolor{mygreen}{rgb}{0, .5, 0}
+\setaddedmarkup{\color{mygreen}{\uline{#1}}}
+\setdeletedmarkup{\color{alizarin}{\uwave{#1}}}
 
@@
 \newcommand{\rmk}[1]{{\color{red}{#1}}}
-
-\title[Time-domain CSEM]{Fast time-domain electromagnetic modelling in the frequency domain}
+\newcommand{\ohmm}{\ensuremath{\Omega\,}\text{m}\xspace}
+
+\title[\replaced{Transform for expensive EM kernels}{Time-domain CSEM}]{
+  \replaced{Fast Fourier transformation of electromagnetic data for expensive kernels}{Fast time-domain electromagnetic modelling in the frequency domain}}
 
 \author[D. Werthmüller \emph{et al.}]
-  {\Large Dieter Werthmüller$^1$,        % orcid: 0000-0002-8575-2484
+  {\Large
+   Dieter Werthmüller$^1$,     % orcid: 0000-0002-8575-2484
    Wim A. Mulder$^{1,2}$ and   % orcid: 0000-0001-7020-9297
@@
    \\
-   \footnotesize
+   {\footnotesize
     $^1$ TU Delft, Building 23, Stevinweg 1 / PO-box 5048, 2628 CN Delft (NL).
-  E-mail: \href{mailto:Dieter@Werthmuller.org}{Dieter@Werthmuller.org}\\[-.3em]
-   \footnotesize
-  $^2$ Shell Global Solutions International BV, Grasweg 31, 1031 HW Amsterdam(NL)
+    E-mail: \href{mailto:Dieter@Werthmuller.org}{Dieter@Werthmuller.org}
+   }\\[-.3em]
+   {\footnotesize
+    $^2$ Shell Global Solutions International BV, Grasweg 31, 1031 HW
+    Amsterdam(NL)
+   }
   }
@@
 \begin{document}
+\large %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TODO %%%
 
@@
 frequencies and interpolation. This methodology requires typically 15 to 20
-frequencies to cover a wide range of offsets in one-dimensional test models.
+frequencies to cover a wide range of offsets\deleted{in one-dimensional test
+models}.
 %
@@
 %
-Looking carefully at these three \added{key} points results in much smaller
+Looking carefully at these \deleted{three} points results in much smaller
 computation times with speedup factors of ten or more over previous results. A
@@
 the approach with some numerical results.
+
+{\color{alizarin}
+\verb!v v v REMOVE COMMENTS v v v REMOVE COMMENTS v v v REMOVE COMMENTS v v v!
+}
+{\color{mygreen}
+\begin{itemize}
+  \item Too generic; more references
+  \item what are the methodological advantages with respect to other
+    approaches.
+\end{itemize}
+
+Redo intro: Focus on transform, not on frequency- or time-domain EM.
+
+\begin{itemize}
+  \item Wim: Include perhaps (one of) explicitly, instead of implicitly:
+    Druskin, V. L., and L. A. Knizhnerman, 1994, Spectral approach to solving
+    three-dimensional Maxwell’s diffusion equations in the time and frequency
+    domains, Radio Science, 29, 937--953. doi: 10.1029/94RS00747 Druskin, V.
+    L., L. A. Knizhnerman, and P. Lee, 1999, New spectral Lanczos decomposition
+    method for induction modeling in arbitrary 3D geometry, Geophysics, 64,
+    701--706. doi: 10.1190/1.1444579.
+  \item Evert: I also think we should respond a bit more accommodating the
+    reviewer's request. Mulder 2008 is quite old now and papers were published
+    after that dealing with solutions directly in time domain. We can write
+    that in this paper we include only papers that were not reference in Mulder
+    et al. (2008) and refer to that for earlier papers. New are Mittet 2010
+    (FDTD of waves and correspondence principle, GPY), Um et al. 2010 (Finite
+    element, SEG abstract) and Zhou et al. 2020 TGRS
+    (10.1109/TGRS.2020.2979798). In terms of number of references, you can
+    decide to not worry to much about that, or remove some others. You have
+    several references to works on the differences between measuring/analysing
+    data in time or frequency domain. That could go out, e.g.
+  \item Evert: i think you can group separate points under the recommendation
+    that comes from his first concern. I understand his recommendation as:
+    "Write an introduction that is more focused on time domain CSEM modeling."
+    That should be done in a comprehensive way and include direct time domain
+    methods such as Krylov subspace, Simpeg, FDTD of waves and use
+    correspondence principle. It can be interesting to also include more
+    arbitrary source time functions and not only step switch on or off and
+    impulse response. For the filter approach, in principle, all FD codes are
+    suitable to be used. His second concern phrased as a recommendation: "Show
+    results from a more complex model." after that the specific points come.
+  \item Evert: I think that Commer wanted to refer to Druskin and Knihznerman
+    (i am sure spelling is not correct) in radio science 1994, which is most
+    likely the older version of the code he mentions. We refer to that 1994
+    paper in Mulder 2008.
+  \item Evert: The main issue is that the Druskin work it is what Commer uses.
+    This is the principal issue. This comes back at point 9. Perhaps now you
+    can include it in the introduction as a motivation to work on this topic.
+  \item Evert: So let us make an explicit sentence in the introduction to
+    clarify this. We also may need to change the text in the introduction to
+    avoid making our life difficult by selecting some papers while not
+    selecting others. This can be done by only selecting review papers on the
+    topic and write that that is our choice here.
+  \item Wim: Is that different from doi 10.1190/1.1801936 (mentioned in
+    Mulder,Wirianto,Slob)? Then also include Haber from that paper, or just all
+    references of section 2 (complexity) in that paper?
+  \item Evert: This all belongs to one concern: The introduction is too general
+    and insufficiently addresses the actual topic of obtaining time domain data
+    by CSEM modeling. That does not mean you need to repeat everything. You
+    could start by writing that we will only discuss new developments since
+    2008, and refer to Mulder 2008 for older works that relate to the topic of
+    time domain CSEM data modeling. I don't think Commer actually wants an
+    extensive review of the literature and include his work. He writes that the
+    introduction should be more focused on the topic of the paper.
+  \item Evert: i have no problem with mentioning in the paper that our idea
+    grid coarsening can be used in the method of Commer and Newman (2005). In
+    that case no further response is necessary, we tick it off by saying, we
+    refer to Commer and Newman (2005) on grid-coarsening. i would limit
+    references to Commer to this paper then. I don't think we say anywhere that
+    our method is the best in any sense. It seems this reviewer is in
+    competition. I don't think we should be in competition. We work for the
+    public domain. We put forward our ideas. We leave to the readers to decide
+    which options they prefer. We should not be forced need to sell our method.
+    Of course, my personal opinion does not change the reality that the AE has
+    written that he wants to see the majority of the reviewers comments
+    implemented in our revision. We should accept that.
+  \item Evert: If we judge this is part of the majorty to be implemented as the
+    Associate Editor says, it means the AE agrees with Commer and we should do
+    it. No need to add references. In the introduction, just add a summary of
+    the order analysis given in Mulder 2008 where some order analysis was
+    done.
+  \item References provided by Commer:
+    \begin{itemize}
+      \item Commer \& Newman, 2005, An accelerated time domain finite
+        difference simulation scheme for three-dimensional transient
+        electromagnetic modeling using geometric multigrid concepts, Radio
+        Science, 41, RS3007.
+      \item Commer et al., 2006, New results on the resistivity structure of
+        Merapi Volcano(Indonesia), derived from three-dimensional restricted
+        inversion of long-offset transient electromagnetic data, GJI.
+      \item Druskin \& Zaslavsky, 2011, Solution of 3D time-domain
+        electromagnetic problems using optimal subspace projection,
+        Geophysics.
+      \item Weiland, T., 1996, Time domain EM field computation with finite
+        difference methods, Int. J. Num. Mod., 9, 295-319.
+      \item Yuji Mitsuhata, Toshihiro Uchida, Yutaka Murakami, Hiroshi Amano,
+        The Fourier transform of controlled-source time-domain electromagnetic
+        data by smooth spectrum inversion, Geophysical Journal International,
+        144, January 2001, 123-135.
+    \end{itemize}
+\end{itemize}
+
+}
+{\color{alizarin}
+\verb!^ ^ ^ REMOVE COMMENTS ^ ^ ^ REMOVE COMMENTS ^ ^ ^ REMOVE COMMENTS ^ ^ ^!
+}
 
@@
 examples in this article are run on a laptop with an i7-6600U\,CPU\,@\,2.6\,GHz
-(x4) and 16\,GB of memory, using Ubuntu 18.04 and Python 3.7. For the
-semi-analytical computations of layered models we use the open-source (Apache
-License 2.0) code \empymod \citep{GEO.17.Werthmuller}.
+(x4) and 16\,GB of memory, using Ubuntu \replaced{20.04}{18.04} and Python
+\replaced{3.8}{3.7}. \added{All examples were run using a single thread, but
+parallelization over frequencies is straight forward.} For the semi-analytical
+computations of layered models we use the open-source (Apache License 2.0) code
+\empymod \citep{GEO.17.Werthmuller}.
 
@@
 frequencies. The fewer frequencies required, the quicker we obtain the
-time-domain result. We decided to use a regular spacing of frequencies on a
+time-domain result. \added{The same applies for the reverse operation, the
+fewer times required, the quicker we obtain frequency-domain results. Even
+though we limit our analysis here to frequency-to-time transformation, most of
+the arguments are reversible and applicable to time-to-frequency
+transformations.} We decided to use a regular spacing of frequencies on a
 log-scale, rather than an adaptive scheme. This approach is favourable if a
@@
 in comparison with the standard FFT: the required input frequencies are equally
-spaced on a logarithmic scale, and they only require either the real or the
-imaginary part of the frequency-domain response. We can take advantage of that
-by using only the imaginary part of the frequency-domain response. The
-imaginary part goes to zero when the frequency goes to zero or to infinity,
-with the advantage that knowing the endpoints makes it possible to convert the
-extrapolation of missing frequencies into interpolation.
+spaced on a logarithmic scale\added{ (natural logarithm in the case of the DLF
+and decimal logarithm in the case of the FFTLog)}, and they only require either
+the real or the imaginary part of the frequency-domain response. We can take
+advantage of that by using only the imaginary part of the frequency-domain
+response. The imaginary part goes to zero when the frequency goes to zero or to
+infinity, with the advantage that knowing the endpoints makes it possible to
+convert the extrapolation of missing frequencies into \added{an }interpolation.
 
@@
   \centering
-  \includegraphics[width=\cwidth]{GUI-FFTLog-lin}
+  \includegraphics[width=\cwidth]{01-GUI-FFTLog-lin}
   \caption{Interactive frequency selection for a user-provided layered model
@@
 The example model is a marine scenario with 1\,km water depth of resistivity
-$\rho=0.3\,\Omega$\,m ($\rho = \sigma^{-1}$) and a 100\,m thick target of
-100\,$\Omega$\,m at 1\,km below the seafloor in a background of 1\,$\Omega$\,m.
-The source is 50\,m above the seafloor, the receiver is on the seafloor, and
-the response is the inline $x$-directed $E$-field. The left subplot shows the
-imaginary part in the frequency domain and the right plot the corresponding
-impulse response in the time domain. The red lines are the
-\replaced{semi-analytical responses}{precise results obtained from \empymod}.
-The blue circles indicate the actually computed responses and the black dots
-the frequencies which are interpolated or set to zero. The resulting
-time-domain response has a relative error of less than 1\,\% everywhere except
-for very early times.
+$\rho=0.3\,\ohmm$ ($\rho = \sigma^{-1}$) and a 100\,m thick target of
+100\,\ohmm at 1\,km below the seafloor in a background of 1\,\ohmm. The source
+is 50\,m above the seafloor, the receiver is on the seafloor, and the response
+is the inline $x$-directed $E$-field. The left subplot shows the imaginary part
+in the frequency domain and the right plot the corresponding impulse response
+in the time domain. The red lines are the \replaced{semi-analytical
+responses}{precise results obtained from \empymod}. The blue circles indicate
+the actually computed responses and the black dots the frequencies which are
+interpolated or set to zero. The resulting time-domain response has a relative
+error of less than 1\,\% everywhere except for very early times.
 
@@
   \centering
-  \subfloat[]{{\includegraphics[width=\cwidth]{GUI-FFTLog-log}}%
+  \subfloat[]{{\includegraphics[width=0.5\fwidth]{02a-GUI-FFTLog-log}}%
   \label{fig:GUI-FFTLog-log}}\hfill
-  \subfloat[]{{\includegraphics[width=\cwidth]{GUI-DLF-lin}}%
+  \subfloat[]{{\includegraphics[width=0.5\fwidth]{02b-GUI-DLF-lin}}%
   \label{fig:GUI-DLF-lin}}\\
@@
 required for the shown offset of 5\,km, a few of the lower and higher
-frequencies could be left out. However, with this frequency-selection we can
-model a wide range of offsets. This is shown in Figure \ref{fig:multi-offset},
-where the above parameters were used to model the responses at offsets
-$r=1.5,3,6$, and 12\,km.
-%
-\begin{figure}
+frequencies could be left out. \added{But even so we only need to compute
+50\,\% of the required frequencies in the case of the FFTLog, and only 14\,\%
+in the case of the DLF.} \replaced{W}{However, w}ith this frequency-selection
+we can model a wide range of offsets. This is shown in Figure
+\ref{fig:multi-offset}, where the above parameters were used to model the
+responses at offsets $r=1.5,3,6$, and 12\,km.
+%
+\begin{figure*}
   \centering
-  \includegraphics[width=\cwidth]{multi-offset}
+  \includegraphics[width=0.75\fwidth]{03-multi-offset}
   \caption{Normalized (a) frequency- and (b) time-domain responses for offsets
@@
     responses, the black dots are the responses which are set to zero or
-    interpolated. The black curves are the analytical responses.}
+    interpolated. The \replaced{grey}{black} curves are the analytical
+    responses.}
   \label{fig:multi-offset}
-\end{figure}
+\end{figure*}
 %
@@
 frequency, and we then use PCHIP to interpolate the missing frequencies (as we
-work on a logarithmic scale we cannot choose 0\,Hz, but instead take
-$10^{-100}$\,Hz). In the case of the DLF method we also have to interpolate in
-between the computed frequencies, for which we found it better to use a cubic
-spline. As can be seen from Figures \ref{fig:GUI-FFTLog-lin} and
-\ref{fig:GUI-DLF-lin}, using the FFTLog or the DLF with the same actually
-computed frequencies results in very similar responses.
+work on a logarithmic scale we cannot choose 0\,Hz \added{nor 0\,V/m}, but
+instead take $10^{-100}$\,Hz \added{and $10^{-100}$\,V/m}). In the case of the
+DLF method we also have to interpolate in between the computed frequencies, for
+which we found it better to use a cubic spline. As can be seen from Figures
+\ref{fig:GUI-FFTLog-lin} and \ref{fig:GUI-DLF-lin}, using the FFTLog or the DLF
+with the same actually computed frequencies results in very similar responses.
 
@@
 electrically conducting boundary condition. It can also have a much higher
-stretching $\alpha_\mr{c}$. In our scheme we have chosen $D_\mr{c}$ such that
-the distance for the signal diffusing from the source to the boundary and back
-to the receiver closest to the boundary is at least two wavelengths, after
-which the initial signal is reduced to a millionth of its strength. The
-wavelength to compute $D_\mr{c}$ is given by
+stretching $\alpha_\mr{c}$. In \replaced{the presented examples}{our scheme} we
+have chosen $D_\mr{c}$ such that the distance for the signal diffusing from the
+source to the boundary and back to the receiver closest to the boundary is at
+least two wavelengths, after which the initial signal is reduced to a millionth
+of its strength. The wavelength \added{$\lambda$} to compute $D_\mr{c}$ is
+given by
 %
 \begin{equation}
-  2\lambda = 4\pi\delta(f, \sigma_\mr{ave})
-  \ \approx \ 6324.7/\sqrt{f\sigma_\mr{ave}}\, ,
+  \deleted{2}\lambda = \replaced{2}{4}\pi\delta(f, \sigma_\mr{ave})
+  \ \approx \ \replaced{3162}{6324.7}/\sqrt{f\sigma_\mr{ave}}\, ,
  \label{eq:lambda}
@@
 criteria. The number of cells becomes therefore also a function of frequency.
-To go from the model grid to the computational grid, we use the
-volume-averaging technique on logarithmic resistivities, as used in
-\cite{GEO.07.Plessix}. While this technique ensures that the total resistivity
-in the subsurface remains the same, it does not consider effective-medium
-theory \citep{GEO.03.Davydycheva}, for instance, the apparent anisotropy from a
-stack of finely layered formations of varying resistivity.
+\added{It is important to note that this is our implementation of an adaptive
+grid, but there are certainly other possibilities. The relevant point for fast
+computations is that the adaptive gridding tries to minimize the number of
+required cells while being fine enough and at the same time large enough to
+give accurate results. This is generally best done in a frequency and
+conductivity dependent manner.} To go from the model grid to the computational
+grid, we use the volume-averaging technique on logarithmic resistivities, as
+used in \cite{GEO.07.Plessix}. While this technique ensures that the total
+resistivity in the subsurface remains the same, it does not consider
+effective-medium theory \citep{GEO.03.Davydycheva}, for instance, the apparent
+anisotropy from a stack of finely layered formations of varying resistivity.
 
@@
 measured by an inline receiver with an offset of 900\,m in a homogeneous space
-of $1\,\Omega\,$m. We chose this simple example to compare it with the
-analytical solution and with previously published results. We used the
-following values to define the required frequencies: $f_\mr{min}=0.05\,$Hz,
-$f_\mr{max}=21\,$Hz, using FFTLog with 5 frequencies per decade. This results
-in 14 frequencies to compute from 0.05\,Hz to 20.0\,Hz. The complete frequency
-range for the transform includes 30 frequencies from 0.0002\,Hz to 126.4\,Hz.
-For the adaptive gridding the following inputs were used: $n_\delta=12$,
-minimum cell width must be between 20 and 40\,m, and $\alpha_\mr{s}=1$,
+of 1\,\ohmm. We chose this simple example to compare it with the analytical
+solution and with previously published results. We used the following values to
+define the required frequencies: $f_\mr{min}=0.05\,$Hz, $f_\mr{max}=21\,$Hz,
+using FFTLog with 5 frequencies per decade. This results in 14 frequencies to
+compute from 0.05\,Hz to 20.0\,Hz. The complete frequency range for the
+transform includes 30 frequencies from 0.0002\,Hz to 126.4\,Hz. For the
+adaptive gridding the following inputs were used: $n_\delta=12$, minimum cell
+width must be between 20 and 40\,m, and $\alpha_\mr{s}=1$,
 $\alpha_\mr{c}=[1,1.3]$. This created grids with cell numbers between
@@
     %
-    20.0    &  3 & 80$\times$24$\times$24 & 20 & 1.26 \\
-    12.6    &  6 & 96$\times$32$\times$32 & 20 & 1.17 \\
-     7.98   &  8 & 96$\times$32$\times$32 & 20 & 1.20 \\
-     5.03   &  7 & 96$\times$32$\times$32 & 20 & 1.23 \\
-     3.18   &  7 & 80$\times$32$\times$32 & 24 & 1.25 \\
-     2.00   &  5 & 80$\times$32$\times$32 & 30 & 1.21 \\
-     1.26   &  4 & 64$\times$32$\times$32 & 37 & 1.21 \\
-     0.798  &  5 & 64$\times$32$\times$32 & 40 & 1.23 \\
-     0.503  &  5 & 64$\times$32$\times$32 & 40 & 1.26 \\
-     0.318  &  5 & 64$\times$32$\times$32 & 40 & 1.28 \\
-     0.200  & 10 & 64$\times$40$\times$40 & 40 & 1.27 \\
-     0.126  & 10 & 64$\times$40$\times$40 & 40 & 1.30 \\
-     0.0798 & 14 & 80$\times$40$\times$40 & 40 & 1.26 \\
-     0.0503 & 15 & 80$\times$40$\times$40 & 40 & 1.28 \\
+    20.0    &\replaced{ 4}{ 3} & 80$\times$24$\times$24 & 20 & 1.26 \\
+    12.6    &\replaced{ 8}{ 6} & 96$\times$32$\times$32 & 20 & 1.17 \\
+     7.98   &\replaced{ 8}{ 8} & 96$\times$32$\times$32 & 20 & 1.20 \\
+     5.03   &\replaced{ 8}{ 7} & 96$\times$32$\times$32 & 20 & 1.23 \\
+     3.18   &\replaced{ 7}{ 7} & 80$\times$32$\times$32 & 24 & 1.23 \\
+     2.00   &\replaced{ 7}{ 5} & 80$\times$32$\times$32 & 30 & 1.21 \\
+     1.26   &\replaced{ 5}{ 4} & 64$\times$32$\times$32 & 37 & 1.21 \\
+     0.798  &\replaced{ 5}{ 5} & 64$\times$32$\times$32 & 40 & 1.23 \\
+     0.503  &\replaced{ 5}{ 5} & 64$\times$32$\times$32 & 40 & 1.26 \\
+     0.318  &\replaced{ 5}{ 5} & 64$\times$32$\times$32 & 40 & 1.28 \\
+     0.200  &\replaced{ 8}{10} & 64$\times$40$\times$40 & 40 & 1.27 \\
+     0.126  &\replaced{ 8}{10} & 64$\times$40$\times$40 & 40 & 1.30 \\
+     0.0798 &\replaced{10}{14} & 80$\times$40$\times$40 & 40 & 1.26 \\
+     0.0503 &\replaced{11}{15} & 80$\times$40$\times$40 & 40 & 1.28 \\
     %
@@
   \centering
-  \includegraphics[width=\fwidth]{fullspace}
+  \includegraphics[width=\fwidth]{04-fullspace}
   \caption{(a) Frequency- and (b) time-domain results for the homogeneous space
@@
   \centering
-  \includegraphics[width=\cwidth]{fullspace-log}
+  \includegraphics[width=\cwidth]{05-fullspace-log}
   \caption{Same as in Figure~\ref{fig:fullspace} (b), but on a logarithmic
@@
 took less than two minutes in total, where the individual frequencies took
-between 3 and 15\,s to run.
+between \replaced{4}{3} and \replaced{11}{15}\,s to run.
 
@@
     %
-    31.8     & 10 & 128$\times$40$\times$40 & 100 & 1.36 \\
-    20.0     &  9 & 128$\times$40$\times$40 & 100 & 1.36 \\
-    12.6     &  9 & 128$\times$40$\times$40 & 100 & 1.36 \\
-     7.98    & 17 & 128$\times$40$\times$40 & 100 & 1.36 \\
-     5.03    & 21 & 128$\times$40$\times$40 & 100 & 1.44 \\
-     3.18    & 20 & 128$\times$40$\times$40 & 100 & 1.48 \\
-     2.00    & 18 & 128$\times$40$\times$40 & 100 & 1.49 \\
-     1.26    & 22 & 128$\times$40$\times$48 & 100 & 1.36 \\
-     0.798   & 26 & 128$\times$40$\times$48 & 100 & 1.36 \\
-     0.503   & 29 & 128$\times$40$\times$48 & 100 & 1.36 \\
-     0.318   & 42 & 128$\times$40$\times$48 & 100 & 1.36 \\
-     0.200   & 41 & 128$\times$40$\times$48 & 100 & 1.38 \\
-     0.126   & 48 & 128$\times$40$\times$48 & 100 & 1.40 \\
-     0.0798  & 54 & 128$\times$40$\times$48 & 100 & 1.41 \\
-     0.0503  & 59 & 128$\times$40$\times$48 & 100 & 1.44 \\
-     0.0318  & 74 & 128$\times$40$\times$48 & 100 & 1.44 \\
-     0.0200  & 81 & 128$\times$40$\times$48 & 100 & 1.47 \\
-     0.0126  & 96 & 128$\times$40$\times$48 & 100 & 1.48 \\
-     0.00798 & 93 & 128$\times$40$\times$48 & 100 & 1.49 \\
+    31.8     & \replaced{16}{10} & 128$\times$40$\times$40 & 100 & 1.36 \\
+    20.0     & \replaced{17}{ 9} & 128$\times$40$\times$40 & 100 & 1.36 \\
+    12.6     & \replaced{17}{ 9} & 128$\times$40$\times$40 & 100 & 1.36 \\
+     7.98    & \replaced{17}{17} & 128$\times$40$\times$40 & 100 & 1.36 \\
+     5.03    & \replaced{17}{21} & 128$\times$40$\times$40 & 100 & 1.44 \\
+     3.18    & \replaced{17}{20} & 128$\times$40$\times$40 & 100 & 1.48 \\
+     2.00    & \replaced{17}{18} & 128$\times$40$\times$40 & 100 & 1.49 \\
+     1.26    & \replaced{20}{22} & 128$\times$40$\times$48 & 100 & 1.36 \\
+     0.798   & \replaced{24}{26} & 128$\times$40$\times$48 & 100 & 1.36 \\
+     0.503   & \replaced{27}{29} & 128$\times$40$\times$48 & 100 & 1.36 \\
+     0.318   & \replaced{27}{42} & 128$\times$40$\times$48 & 100 & 1.36 \\
+     0.200   & \replaced{34}{41} & 128$\times$40$\times$48 & 100 & 1.38 \\
+     0.126   & \replaced{34}{48} & 128$\times$40$\times$48 & 100 & 1.40 \\
+     0.0798  & \replaced{37}{54} & 128$\times$40$\times$48 & 100 & 1.41 \\
+     0.0503  & \replaced{37}{59} & 128$\times$40$\times$48 & 100 & 1.44 \\
+     0.0318  & \replaced{44}{74} & 128$\times$40$\times$48 & 100 & 1.44 \\
+     0.0200  & \replaced{44}{81} & 128$\times$40$\times$48 & 100 & 1.47 \\
+     0.0126  & \replaced{37}{96} & 128$\times$40$\times$48 & 100 & 1.48 \\
+     0.00798 & \replaced{43}{93} & 128$\times$40$\times$48 & 100 & 1.49 \\
     %
@@
   \centering
-  \includegraphics[width=\fwidth]{marine}
+  \includegraphics[width=\fwidth]{06-marine}
   \caption{Electric inline response at an offset of 5\,km for a shallow marine,
@@
   \centering
-  \includegraphics[width=\cwidth]{marine-multioffset}
+  \includegraphics[width=\cwidth]{07-marine-multioffset}
   \caption{Time-domain responses for offsets of 3, 5, and 7\,km for the same
@@
 the previous section; however, for the adaptive gridding in the horizontal
-directions $\rho_\mr{ave}=1\,\Omega\,$m was used instead of
-$\rho_\mr{ave}=\num{10000}\,\Omega\,$m. Having the boundaries too near in the
+directions $\rho_\mr{ave}=1\,\ohmm$ was used instead of
+$\rho_\mr{ave}=\num{10000}\,\ohmm$. Having the boundaries too near in the
 horizontal directions leads to worse results for most frequencies and entirely
@@
   \centering
-  \includegraphics[width=\fwidth]{marine-wrong-x-y}
+  \includegraphics[width=\fwidth]{08-marine-wrong-x-y}
   \caption{Same model as used for Figure~\ref{fig:marine}, but with the
@@
 the lower of two halfspaces, as depicted in Figure~\ref{fig:3d-model}. The
-target has resistivity $\rho_\mr{tg}=100\,\Omega\,$m, the upper halfspace
-corresponds to seawater with $\rho_\mr{sea}=0.3\,\Omega\,$m, and the lower
-halfspace is the background with $\rho_\mr{bg}=1\,\Omega\,$m.
+target has resistivity $\rho_\mr{tg}=100\,\ohmm$, the upper halfspace
+corresponds to seawater with $\rho_\mr{sea}=0.3\,\ohmm$, and the lower
+halfspace is the background with $\rho_\mr{bg}=1\,\ohmm$.
 %
@@
   \centering
-  \includegraphics[width=\cwidth]{3d-model}
+  \includegraphics[width=\cwidth]{09-3d-model}
   \caption{Three-dimensional block embedded in the lower of two halfspaces. The
@@
   \centering
-  \includegraphics[width=\fwidth]{3d-result}
+  \includegraphics[width=\fwidth]{10-3d-result}
   \caption{Responses for the model outlined in Figure~\ref{fig:3d-model} using
@@
 
+\added{¡TODO CHANGE THIS LAST PARAGRAPH!}
 Both codes took a little over 7 minutes to compute the two models (single

### 2020-11-27 ManuscriptTime.tex
EOD

@@
     approaches.
+  \item wants: TEM can be accelerated using similar coarse-gridding
+    considerations as mentioned in your manuscript. In Commer \& Newman (2005),
+    the FDTD method could be accelerated through grid-coarsening during
+    runtime, using the fact that the transient field-decay filters out
+    high-frequency components.
+  \item I do believe the subspace-projection method should definitely be
+    discussed in a pro/con manner.
+  \item \ldots
 \end{itemize}
+
+Dieter: One one hand we agree that the introduction could be expanded, on the
+other hand we are not fond of repeating everything again. The introduction
+contains quite some extensive, good review papers (Oristaglio and Spies, 1999;
+Avdeev, 2005; Börner, 2010; Constable, 2010). The comparison between different
+approaches for time-domain computation was done in quite some detail in Mulder
+(2008)
+
+TODO: We think that there is a misunderstanding here. First we mention Avdeev
+(2005) and Börner (2010), and say that these are very good review papers up to
+2005 and 2010, respectively. Recent publications then means publications since
+these two review papers were published. We do not want to repeat again all the
+pioneering publications, that has been done, but highlight a few recent ones.
+We changed the wording in the introduction to make this clear.
 
@@
     %
-    20.0    &\replaced{ 4}{ 3} & 80$\times$24$\times$24 & 20 & 1.26 \\
-    12.6    &\replaced{ 8}{ 6} & 96$\times$32$\times$32 & 20 & 1.17 \\
-     7.98   &\replaced{ 8}{ 8} & 96$\times$32$\times$32 & 20 & 1.20 \\
-     5.03   &\replaced{ 8}{ 7} & 96$\times$32$\times$32 & 20 & 1.23 \\
-     3.18   &\replaced{ 7}{ 7} & 80$\times$32$\times$32 & 24 & 1.23 \\
-     2.00   &\replaced{ 7}{ 5} & 80$\times$32$\times$32 & 30 & 1.21 \\
-     1.26   &\replaced{ 5}{ 4} & 64$\times$32$\times$32 & 37 & 1.21 \\
-     0.798  &\replaced{ 5}{ 5} & 64$\times$32$\times$32 & 40 & 1.23 \\
-     0.503  &\replaced{ 5}{ 5} & 64$\times$32$\times$32 & 40 & 1.26 \\
-     0.318  &\replaced{ 5}{ 5} & 64$\times$32$\times$32 & 40 & 1.28 \\
-     0.200  &\replaced{ 8}{10} & 64$\times$40$\times$40 & 40 & 1.27 \\
-     0.126  &\replaced{ 8}{10} & 64$\times$40$\times$40 & 40 & 1.30 \\
-     0.0798 &\replaced{10}{14} & 80$\times$40$\times$40 & 40 & 1.26 \\
-     0.0503 &\replaced{11}{15} & 80$\times$40$\times$40 & 40 & 1.28 \\
+    20.0    &  4 & 80$\times$24$\times$24 & 20 & 1.26 \\
+    12.6    &  8 & 96$\times$32$\times$32 & 20 & 1.17 \\
+     7.98   &  8 & 96$\times$32$\times$32 & 20 & 1.20 \\
+     5.03   &  8 & 96$\times$32$\times$32 & 20 & 1.23 \\
+     3.18   &  7 & 80$\times$32$\times$32 & 24 & 1.23 \\
+     2.00   &  7 & 80$\times$32$\times$32 & 30 & 1.21 \\
+     1.26   &  5 & 64$\times$32$\times$32 & 37 & 1.21 \\
+     0.798  &  5 & 64$\times$32$\times$32 & 40 & 1.23 \\
+     0.503  &  5 & 64$\times$32$\times$32 & 40 & 1.26 \\
+     0.318  &  5 & 64$\times$32$\times$32 & 40 & 1.28 \\
+     0.200  &  8 & 64$\times$40$\times$40 & 40 & 1.27 \\
+     0.126  &  8 & 64$\times$40$\times$40 & 40 & 1.30 \\
+     0.0798 & 10 & 80$\times$40$\times$40 & 40 & 1.26 \\
+     0.0503 & 11 & 80$\times$40$\times$40 & 40 & 1.28 \\
     %
@@
     %
-    31.8     & \replaced{16}{10} & 128$\times$40$\times$40 & 100 & 1.36 \\
-    20.0     & \replaced{17}{ 9} & 128$\times$40$\times$40 & 100 & 1.36 \\
-    12.6     & \replaced{17}{ 9} & 128$\times$40$\times$40 & 100 & 1.36 \\
-     7.98    & \replaced{17}{17} & 128$\times$40$\times$40 & 100 & 1.36 \\
-     5.03    & \replaced{17}{21} & 128$\times$40$\times$40 & 100 & 1.44 \\
-     3.18    & \replaced{17}{20} & 128$\times$40$\times$40 & 100 & 1.48 \\
-     2.00    & \replaced{17}{18} & 128$\times$40$\times$40 & 100 & 1.49 \\
-     1.26    & \replaced{20}{22} & 128$\times$40$\times$48 & 100 & 1.36 \\
-     0.798   & \replaced{24}{26} & 128$\times$40$\times$48 & 100 & 1.36 \\
-     0.503   & \replaced{27}{29} & 128$\times$40$\times$48 & 100 & 1.36 \\
-     0.318   & \replaced{27}{42} & 128$\times$40$\times$48 & 100 & 1.36 \\
-     0.200   & \replaced{34}{41} & 128$\times$40$\times$48 & 100 & 1.38 \\
-     0.126   & \replaced{34}{48} & 128$\times$40$\times$48 & 100 & 1.40 \\
-     0.0798  & \replaced{37}{54} & 128$\times$40$\times$48 & 100 & 1.41 \\
-     0.0503  & \replaced{37}{59} & 128$\times$40$\times$48 & 100 & 1.44 \\
-     0.0318  & \replaced{44}{74} & 128$\times$40$\times$48 & 100 & 1.44 \\
-     0.0200  & \replaced{44}{81} & 128$\times$40$\times$48 & 100 & 1.47 \\
-     0.0126  & \replaced{37}{96} & 128$\times$40$\times$48 & 100 & 1.48 \\
-     0.00798 & \replaced{43}{93} & 128$\times$40$\times$48 & 100 & 1.49 \\
+    31.8     & 16 & 128$\times$40$\times$40 & 100 & 1.36 \\
+    20.0     & 17 & 128$\times$40$\times$40 & 100 & 1.36 \\
+    12.6     & 17 & 128$\times$40$\times$40 & 100 & 1.36 \\
+     7.98    & 17 & 128$\times$40$\times$40 & 100 & 1.36 \\
+     5.03    & 17 & 128$\times$40$\times$40 & 100 & 1.44 \\
+     3.18    & 17 & 128$\times$40$\times$40 & 100 & 1.48 \\
+     2.00    & 17 & 128$\times$40$\times$40 & 100 & 1.49 \\
+     1.26    & 20 & 128$\times$40$\times$48 & 100 & 1.36 \\
+     0.798   & 24 & 128$\times$40$\times$48 & 100 & 1.36 \\
+     0.503   & 27 & 128$\times$40$\times$48 & 100 & 1.36 \\
+     0.318   & 27 & 128$\times$40$\times$48 & 100 & 1.36 \\
+     0.200   & 34 & 128$\times$40$\times$48 & 100 & 1.38 \\
+     0.126   & 34 & 128$\times$40$\times$48 & 100 & 1.40 \\
+     0.0798  & 37 & 128$\times$40$\times$48 & 100 & 1.41 \\
+     0.0503  & 37 & 128$\times$40$\times$48 & 100 & 1.44 \\
+     0.0318  & 44 & 128$\times$40$\times$48 & 100 & 1.44 \\
+     0.0200  & 44 & 128$\times$40$\times$48 & 100 & 1.47 \\
+     0.0126  & 37 & 128$\times$40$\times$48 & 100 & 1.48 \\
+     0.00798 & 43 & 128$\times$40$\times$48 & 100 & 1.49 \\
     %
@@
 a relative error of a few percents at most; the result \replaced{obtained
-through transformation}{for \emg3d} seems to be better at early times. We
-cannot compare the errors for the response that includes the target for lack of
-an analytical solution. The \added{semi-analytical} background model
-\deleted{from \empymod }is only included to show that there is a significant
-response from the target. We therefore show the normalized difference
-(\replaced{NRMSD}{ND}) between the two \replaced{responses $R_1$ and $R_2$}{3D
-codes} as a percentage, where \replaced{$\mr{NRMSD~(\%)} = 200|R_1 -
-R_2|/(|R_1| + |R_2|)$.}{$\mr{ND}= 50|(\simpeg-\emg3d)/(\simpeg+\emg3d)|$.} The
-\replaced{NRMSD}{ND} between the two codes is below 1\,\% everywhere except for
-early times.
-
-\added{¡TODO CHANGE THIS LAST PARAGRAPH!}
-Both codes took a little over 7 minutes to compute the two models (single
-thread). However, in this particular comparison, the main difference in runtime
-is not frequency-domain computation vs. time-domain computation, but iterative
-solver vs. direct solver. For a test we doubled the number of cells in each
-direction by using 50\,m for the minimum cell width instead of 100\,m, and
-taking also half of the stretching, which results then in a mesh of
-\num{466752} cells \replaced{in the time domain}{for \simpeg} and meshes
-between \num{102400} and \num{614400} \replaced{in the frequency domain}{for
-\emg3d}. Because the memory of our test machine was not large enough for the
-direct solver of \simpeg, the last model had to be run on a cluster. In this
-scenario, \emg3d was about eight times faster (single thread).
+through transformation}{for \emg3d} seems to be better at early times.
+\added{The reason is probably the implemented backward Euler in the time-domain
+code which has an error of order one in time.} We cannot compare the errors for
+the response that includes the target for lack of an analytical solution. The
+\added{1D} background model \deleted{from \empymod }is only included to show
+that there is a significant response from the target. We therefore show the
+normalized difference (\replaced{NRMSD}{ND}) between the two
+\replaced{responses $R_1$ and $R_2$}{3D codes} as a percentage, where
+\replaced{$\mr{NRMSD~(\%)} = 200|R_1 - R_2|/(|R_1| + |R_2|)$.}{$\mr{ND}=
+50|(\simpeg-\emg3d)/(\simpeg+\emg3d)|$.} The \replaced{NRMSD}{ND} between the
+two codes is below 1\,\% everywhere except for early times. Both codes took
+\replaced{roughly 4--5}{a little over 7} minutes to compute the two models
+(single thread). However, in this particular comparison, the main difference in
+runtime is not frequency-domain computation vs. time-domain computation, but
+iterative solver vs. direct solver.
+
+\deleted{\emph{(Deleted last three sentences, as they do not add a lot.)}For a
+test we doubled the number of cells in each direction by using 50\,m for the
+minimum cell width instead of 100\,m, and taking also half of the stretching,
+which results then in a mesh of \num{466752} cells for \simpeg and meshes
+between \num{102400} and \num{614400} for \emg3d. Because the memory of our
+test machine was not large enough for the direct solver of \simpeg, the last
+model had to be run on a cluster. In this scenario, \emg3d was about eight
+times faster (single thread).}
 

### 2020-11-27 ManuscriptTime.tex
EOD

@@
-\documentclass[onecolumn,extra,referee,camera]{gji}
-% \documentclass[onecolumn,extra,camera]{gji}
-% \documentclass[twocolumn,extra,camera]{gji}
+\documentclass[extra, camera,%
+    % onecolumn,   % comment for two columns
+    referee,     % uncomment for double lines
+    % final,       % uncomment for changes
+]{gji}
 \usepackage{timet}
@@
 \usepackage{xspace}
-\usepackage{graphicx}
+\usepackage[final]{graphicx}
 \usepackage[position=top]{subfig}
 \usepackage{ifthen}
-\usepackage[draft, addedmarkup=bf]{changes}
+\usepackage[addedmarkup=bf]{changes}
 % \usepackage[final]{changes}
@@
 
-\title[\replaced{Transform for expensive EM kernels}{Time-domain CSEM}]{
-  \replaced{Fast Fourier transformation of electromagnetic data for expensive kernels}{Fast time-domain electromagnetic modelling in the frequency domain}}
+\title[\replaced{Transforming expensive EM kernels}{Time-domain CSEM}]{
+  \replaced{Fast Fourier transformation of electromagnetic data for
+  computationally expensive kernels}{Fast time-domain electromagnetic modelling
+  in the frequency domain}}
 
@@
 \begin{document}
-\large %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% TODO %%%
 
@@
 %
-Modelling three-dimensional time-domain controlled-source electromagnetic data
-with frequency-domain solvers is an alternative to time-domain solvers. It
-requires the computation of many frequencies if standard Fourier transforms are
-used, which can make it prohibitively expensive in comparison with a
-time-domain computation. The speed of time-domain modelling with
-frequency-domain computation is defined by three key points: solver, method and
-implementation of the Fourier transformation, and gridding.
-%
-The faster the frequency-domain solver, the faster time-domain modelling will
-be. It is important that the solver is robust over a wide range of frequencies.
-A three-dimensional, iterative, frequency-domain multigrid solver satisfies
-this requirement, but any other robust solver will do.
-%
-The method should require as few frequencies as possible while remaining
-robust. As the frequency range spans many orders of magnitude, the required
-frequencies are ideally equally spaced on a logarithmic scale. The fast
-frequency-to-time domain method uses either the digital linear filter method
-or the logarithmic fast Fourier transform together with a careful selection of
-frequencies and interpolation. This methodology requires typically 15 to 20
-frequencies to cover a wide range of offsets\deleted{in one-dimensional test
-models}.
-%
-The gridding should be frequency-dependent, which is accomplished by making it
-a function of skin depth. Optimising for the least number of required cells for
-a given frequency should be combined with optimising for computational speed.
+\deleted{Modelling three-dimensional time-domain controlled-source
+electromagnetic data with frequen\-cy-domain solvers is an alternative to
+time-domain solvers.} \added{Three-dimensional controlled-source
+electromagnetic data is often computed directly in the domain of interest,
+either in the frequency domain or in the time domain. Computing it in one
+domain and transforming it via a Fourier transform to the other domain is a
+viable alternative.} It requires the \replaced{evaluation}{computation} of many
+\replaced{responses in the computational domain}{frequencies} if standard
+Fourier transforms are used, which can make it prohibitively expensive\deleted{
+in comparison with a time-domain computation}. The speed of
+\replaced{such}{time-domain} modelling \replaced{which is followed by a
+transform}{with frequency-domain computation} is defined by three key points:
+solver, method and implementation of the Fourier transformation, and gridding.
+%
+The faster the \deleted{frequency-domain} solver, the faster
+\deleted{time-domain} modelling will be. It is important that the solver is
+robust over a wide range of \added{values (}frequencies\added{ or times)}.
+\deleted{A three-dimensional, iterative, frequency-domain multigrid solver
+satisfies this re\-quirement, but any other robust solver will do.}
+%
+The method should require as few \replaced{kernel evaluations}{frequencies} as
+possible while remaining robust. As the frequency \added{and time}
+range\added{s} span\deleted{s} many orders of magnitude, the required
+\replaced{values}{frequencies} are ideally equally spaced on a logarithmic
+scale. The \added{proposed} fast \deleted{frequency-to-time domain} method uses
+either the digital linear filter method or the logarithmic fast Fourier
+transform together with a careful selection of \replaced{evaluation
+points}{frequencies} and interpolation. \replaced{In frequency-to-time domain
+tests t}{T}his methodology requires typically 15 to 20 frequencies to cover a
+wide range of offsets\deleted{in one-dimensional test models}.
+%
+The gridding should be frequency-\added{ or time-}dependent, which is
+accomplished by making it a function of skin depth. Optimising for the least
+number of required cells \deleted{for a given frequency} should be combined
+with optimising for computational speed.
 %
@@
 computation times with speedup factors of ten or more over previous results. A
-frequency-domain computation of time-domain data can therefore be competitive
-to time-domain solvers if the required frequencies and the corresponding grids
-are carefully chosen.
+\deleted{frequency-domain} computation \replaced{followed by transformation}{of
+time-domain data} can therefore be competitive to \replaced{computation in the
+required }{time-}domain \deleted{solvers} if the required \replaced{evaluation
+points}{frequencies} and the corresponding grids are carefully chosen.
 %

### 2020-11-28 ManuscriptTime.tex
Draft of revision to co-authors

@@
 % Remove once done
-\newcommand{\itodo}[1]{\todo[color=blue!40!white, inline]{\sffamily #1}}
-\newcommand{\mtodo}[1]{\todo[inline]{\sffamily #1}}
-\newcommand{\etodo}[1]{\todo[color=blue!40!white]{#1}}
+% \newcommand{\itodo}[1]{\todo[color=blue!40!white, inline]{\sffamily #1}}
+% \newcommand{\mtodo}[1]{\todo[inline]{\sffamily #1}}
+% \newcommand{\etodo}[1]{\todo[color=blue!40!white]{#1}}
 \newcommand{\dul}[1]{\underline{\underline{#1}}}
@@
 \replaced{responses in the computational domain}{frequencies} if standard
-Fourier transforms are used, which can make it prohibitively expensive\deleted{
-in comparison with a time-domain computation}. The speed of
-\replaced{such}{time-domain} modelling \replaced{which is followed by a
-transform}{with frequency-domain computation} is defined by three key points:
-solver, method and implementation of the Fourier transformation, and gridding.
+Fourier transforms are used\replaced{. This}{, which} can make it prohibitively
+expensive \replaced{if the kernel is time-consuming as is the case in
+three-dimensional electromagnetic modelling}{in comparison with a time-domain
+computation}. The speed of \replaced{such}{time-domain} modelling
+\replaced{that is obtained through a transform}{with frequency-domain
+computation} is defined by three key points: solver, method and implementation
+of the Fourier transformation, and gridding.
 %
@@
 century, see, e.g., \cite{SG.05.Avdeev} and \cite{SG.10.Borner} for
-comprehensive overviews. Recent publications span the width of numerical
+comprehensive overviews. \replaced{Focusing on r}{R}ecent publications
+\added{after these review articles} span\added{s} the width of numerical
 methods from integral equation codes \citep{MGS.17.Kruglyakov} and differential
-equation methods such as finite differences \citep{CAG.13.Sommer}, finite
-elements \citep{GJI.13.Grayver}, and finite volumes \citep{GEO.14.Jahandari}.
-Besides the different methods there are also many different types of
-discretisation. Commonly used, besides rectilinear Yee grids
-\citep{IEEE.66.Yee} as used by our code, are Lebedev grids
-\citep{CMMP.64.Lebedev} but also unstructured grids using, for instance,
+equation methods such as finite differences
+\replaced{\mbox{\citep{GEO.10.Mittet, GEO.11.Zaslavsky, CAG.13.Sommer}}}
+{\mbox{\citep{CAG.13.Sommer}}}, finite elements \citep{GJI.13.Grayver}, and
+finite volumes \citep{GEO.14.Jahandari}. Besides the different methods there
+are also many different types of discretisation. Commonly used, besides
+rectilinear Yee grids \citep{IEEE.66.Yee} as used by our code, are Lebedev
+grids \citep{CMMP.64.Lebedev} but also unstructured grids using, for instance,
 tetrahedral \citep{CAG.17.Cai}. The computational requirement to model CSEM
@@
 
-Under certain conditions it can be competitive to model time-domain data with a
-frequency-domain code, as shown by \cite{GEO.08.Mulder}. The conditions are: a
-sufficiently powerful solver, appropriate frequency selection and
+Under certain conditions it can be \replaced{a viable alternative}{competitive}
+to model time-domain data with a frequency-domain code, as shown by
+\cite{GEO.08.Mulder}. \added{They compared the computational complexity of
+explicit time-stepping schemes (Du Fort-Frankel method as used in, e.g.,
+\mbox{\cite{GEO.04.Commer, GEO.07.Maao}}); implicit schemes \mbox{\citep[e.g.,
+][]{GEO.04.Haber}}, and matrix exponentials and Lanczos reduction schemes
+\mbox{\cite[e.g., ][]{RS.94.Druskin}} with the Fourier transform of
+frequency-domain solutions. Their conclusion was that the Fourier transform
+method can be favourable if the number of required frequencies is small
+relative to the number of grid points in each coordinate.} The conditions
+\added{for fast computation of time-domain data with a frequency-domain code}
+are: a sufficiently powerful solver, appropriate frequency selection and
 interpolation, and an automated gridding, for which they used the
@@
 
-In the next section, we briefly review the methodology as introduced by
+\added{While we use the finite-integration technique
+\mbox{\citep{AEU.77.Weiland}} in the frequency domain to show our method it is
+important to note that our ideas can be used with any solver, it is not bound
+to one or another frequency- or time-domain solver. The ideas for the frequency
+selection can be applied to time selection, and time-dependent, adaptive
+gridding exists as well \mbox{\citep[e.g., ][]{RS.06.Commer}}.} In the next
+section, we briefly review the methodology as introduced by
 \cite{GEO.07.Plessix} and \cite{GEO.08.Mulder} and highlight their advantages
@@
 the approach with some numerical results.
-
-{\color{alizarin}
-\verb!v v v REMOVE COMMENTS v v v REMOVE COMMENTS v v v REMOVE COMMENTS v v v!
-}
-{\color{mygreen}
-\begin{itemize}
-  \item Too generic; more references
-  \item what are the methodological advantages with respect to other
-    approaches.
-  \item wants: TEM can be accelerated using similar coarse-gridding
-    considerations as mentioned in your manuscript. In Commer \& Newman (2005),
-    the FDTD method could be accelerated through grid-coarsening during
-    runtime, using the fact that the transient field-decay filters out
-    high-frequency components.
-  \item I do believe the subspace-projection method should definitely be
-    discussed in a pro/con manner.
-  \item \ldots
-\end{itemize}
-
-Dieter: One one hand we agree that the introduction could be expanded, on the
-other hand we are not fond of repeating everything again. The introduction
-contains quite some extensive, good review papers (Oristaglio and Spies, 1999;
-Avdeev, 2005; Börner, 2010; Constable, 2010). The comparison between different
-approaches for time-domain computation was done in quite some detail in Mulder
-(2008)
-
-TODO: We think that there is a misunderstanding here. First we mention Avdeev
-(2005) and Börner (2010), and say that these are very good review papers up to
-2005 and 2010, respectively. Recent publications then means publications since
-these two review papers were published. We do not want to repeat again all the
-pioneering publications, that has been done, but highlight a few recent ones.
-We changed the wording in the introduction to make this clear.
-
-Redo intro: Focus on transform, not on frequency- or time-domain EM.
-
-\begin{itemize}
-  \item Wim: Include perhaps (one of) explicitly, instead of implicitly:
-    Druskin, V. L., and L. A. Knizhnerman, 1994, Spectral approach to solving
-    three-dimensional Maxwell’s diffusion equations in the time and frequency
-    domains, Radio Science, 29, 937--953. doi: 10.1029/94RS00747 Druskin, V.
-    L., L. A. Knizhnerman, and P. Lee, 1999, New spectral Lanczos decomposition
-    method for induction modeling in arbitrary 3D geometry, Geophysics, 64,
-    701--706. doi: 10.1190/1.1444579.
-  \item Evert: I also think we should respond a bit more accommodating the
-    reviewer's request. Mulder 2008 is quite old now and papers were published
-    after that dealing with solutions directly in time domain. We can write
-    that in this paper we include only papers that were not reference in Mulder
-    et al. (2008) and refer to that for earlier papers. New are Mittet 2010
-    (FDTD of waves and correspondence principle, GPY), Um et al. 2010 (Finite
-    element, SEG abstract) and Zhou et al. 2020 TGRS
-    (10.1109/TGRS.2020.2979798). In terms of number of references, you can
-    decide to not worry to much about that, or remove some others. You have
-    several references to works on the differences between measuring/analysing
-    data in time or frequency domain. That could go out, e.g.
-  \item Evert: i think you can group separate points under the recommendation
-    that comes from his first concern. I understand his recommendation as:
-    "Write an introduction that is more focused on time domain CSEM modeling."
-    That should be done in a comprehensive way and include direct time domain
-    methods such as Krylov subspace, Simpeg, FDTD of waves and use
-    correspondence principle. It can be interesting to also include more
-    arbitrary source time functions and not only step switch on or off and
-    impulse response. For the filter approach, in principle, all FD codes are
-    suitable to be used. His second concern phrased as a recommendation: "Show
-    results from a more complex model." after that the specific points come.
-  \item Evert: I think that Commer wanted to refer to Druskin and Knihznerman
-    (i am sure spelling is not correct) in radio science 1994, which is most
-    likely the older version of the code he mentions. We refer to that 1994
-    paper in Mulder 2008.
-  \item Evert: The main issue is that the Druskin work it is what Commer uses.
-    This is the principal issue. This comes back at point 9. Perhaps now you
-    can include it in the introduction as a motivation to work on this topic.
-  \item Evert: So let us make an explicit sentence in the introduction to
-    clarify this. We also may need to change the text in the introduction to
-    avoid making our life difficult by selecting some papers while not
-    selecting others. This can be done by only selecting review papers on the
-    topic and write that that is our choice here.
-  \item Wim: Is that different from doi 10.1190/1.1801936 (mentioned in
-    Mulder,Wirianto,Slob)? Then also include Haber from that paper, or just all
-    references of section 2 (complexity) in that paper?
-  \item Evert: This all belongs to one concern: The introduction is too general
-    and insufficiently addresses the actual topic of obtaining time domain data
-    by CSEM modeling. That does not mean you need to repeat everything. You
-    could start by writing that we will only discuss new developments since
-    2008, and refer to Mulder 2008 for older works that relate to the topic of
-    time domain CSEM data modeling. I don't think Commer actually wants an
-    extensive review of the literature and include his work. He writes that the
-    introduction should be more focused on the topic of the paper.
-  \item Evert: i have no problem with mentioning in the paper that our idea
-    grid coarsening can be used in the method of Commer and Newman (2005). In
-    that case no further response is necessary, we tick it off by saying, we
-    refer to Commer and Newman (2005) on grid-coarsening. i would limit
-    references to Commer to this paper then. I don't think we say anywhere that
-    our method is the best in any sense. It seems this reviewer is in
-    competition. I don't think we should be in competition. We work for the
-    public domain. We put forward our ideas. We leave to the readers to decide
-    which options they prefer. We should not be forced need to sell our method.
-    Of course, my personal opinion does not change the reality that the AE has
-    written that he wants to see the majority of the reviewers comments
-    implemented in our revision. We should accept that.
-  \item Evert: If we judge this is part of the majorty to be implemented as the
-    Associate Editor says, it means the AE agrees with Commer and we should do
-    it. No need to add references. In the introduction, just add a summary of
-    the order analysis given in Mulder 2008 where some order analysis was
-    done.
-  \item References provided by Commer:
-    \begin{itemize}
-      \item Commer \& Newman, 2005, An accelerated time domain finite
-        difference simulation scheme for three-dimensional transient
-        electromagnetic modeling using geometric multigrid concepts, Radio
-        Science, 41, RS3007.
-      \item Commer et al., 2006, New results on the resistivity structure of
-        Merapi Volcano(Indonesia), derived from three-dimensional restricted
-        inversion of long-offset transient electromagnetic data, GJI.
-      \item Druskin \& Zaslavsky, 2011, Solution of 3D time-domain
-        electromagnetic problems using optimal subspace projection,
-        Geophysics.
-      \item Weiland, T., 1996, Time domain EM field computation with finite
-        difference methods, Int. J. Num. Mod., 9, 295-319.
-      \item Yuji Mitsuhata, Toshihiro Uchida, Yutaka Murakami, Hiroshi Amano,
-        The Fourier transform of controlled-source time-domain electromagnetic
-        data by smooth spectrum inversion, Geophysical Journal International,
-        144, January 2001, 123-135.
-    \end{itemize}
-\end{itemize}
-
-}
-{\color{alizarin}
-\verb!^ ^ ^ REMOVE COMMENTS ^ ^ ^ REMOVE COMMENTS ^ ^ ^ REMOVE COMMENTS ^ ^ ^!
-}
 
@@
 frequency-domain approach appears to be attractive.\guillemotright} Their
-\todo{mention the four methods}
 approach is to minimize the computation time by having, in addition to the just
@@
 \replaced{3.8}{3.7}. \added{All examples were run using a single thread, but
-parallelization over frequencies is straight forward.} For the semi-analytical
-computations of layered models we use the open-source (Apache License 2.0) code
-\empymod \citep{GEO.17.Werthmuller}.
+parallelization over frequencies is straight forward, as the computations are
+independent.} For the semi-analytical computations of layered models we use the
+open-source (Apache License 2.0) code \empymod \citep{GEO.17.Werthmuller}.
 
@@
 advantage of that by using only the imaginary part of the frequency-domain
-response. The imaginary part goes to zero when the frequency goes to zero or to
-infinity, with the advantage that knowing the endpoints makes it possible to
-convert the extrapolation of missing frequencies into \added{an }interpolation.
+response. The imaginary part goes to zero when the frequency goes
+\added{either} to zero or to infinity, with the advantage that knowing the
+endpoints makes it possible to convert the extrapolation of missing frequencies
+into \added{an }interpolation.
 
@@
 
-The example model is a marine scenario with 1\,km water depth of resistivity
-$\rho=0.3\,\ohmm$ ($\rho = \sigma^{-1}$) and a 100\,m thick target of
-100\,\ohmm at 1\,km below the seafloor in a background of 1\,\ohmm. The source
-is 50\,m above the seafloor, the receiver is on the seafloor, and the response
-is the inline $x$-directed $E$-field. The left subplot shows the imaginary part
-in the frequency domain and the right plot the corresponding impulse response
-in the time domain. The red lines are the \replaced{semi-analytical
-responses}{precise results obtained from \empymod}. The blue circles indicate
-the actually computed responses and the black dots the frequencies which are
-interpolated or set to zero. The resulting time-domain response has a relative
-error of less than 1\,\% everywhere except for very early times.
+The \added{shown} example model is a marine scenario with 1\,km water depth of
+resistivity $\rho=0.3\,\ohmm$ ($\rho = \sigma^{-1}$) and a 100\,m thick target
+of 100\,\ohmm at 1\,km below the seafloor in a background of 1\,\ohmm. The
+source is 50\,m above the seafloor, the receiver is on the seafloor, and the
+response is the inline $x$-directed $E$-field. The left subplot shows the
+imaginary part in the frequency domain and the right plot the corresponding
+impulse response in the time domain. The red lines are the
+\replaced{semi-analytical responses}{precise results obtained from \empymod}.
+The blue circles indicate the actually computed responses and the black dots
+the frequencies which are interpolated or set to zero. The resulting
+time-domain response has a relative error of less than 1\,\% everywhere except
+for very early times.
 
@@
 and below 0.3\,s. It also shows the oscillating high-frequency part, which is
-hard to interpolate. Figure \ref{fig:GUI-DLF-lin} is the same as
-\ref{fig:GUI-FFTLog-lin}, but transformed with the DLF method applying the
-81-point sine-cosine filter from \cite{GEO.09.Key}. The same frequencies were
-computed as in the FFTLog case and the missing ones interpolated. The error of
-the corresponding time-domain response is comparable, so either FFTLog or DLF
-can be used.
+hard to interpolate\added{ and we therefore set to zero}. Figure
+\ref{fig:GUI-DLF-lin} is the same as \ref{fig:GUI-FFTLog-lin}, but transformed
+with the DLF method applying the 81-point sine-cosine filter from
+\cite{GEO.09.Key}. The same frequencies were computed as in the FFTLog case and
+the missing ones interpolated. The error of the corresponding time-domain
+response is comparable, so either FFTLog or DLF can be used.
 %
@@
 required for the shown offset of 5\,km, a few of the lower and higher
-frequencies could be left out. \added{But even so we only need to compute
-50\,\% of the required frequencies in the case of the FFTLog, and only 14\,\%
-in the case of the DLF.} \replaced{W}{However, w}ith this frequency-selection
-we can model a wide range of offsets. This is shown in Figure
-\ref{fig:multi-offset}, where the above parameters were used to model the
-responses at offsets $r=1.5,3,6$, and 12\,km.
+frequencies could be left out. With this frequency-selection we can model a
+wide range of offsets. This is shown in Figure \ref{fig:multi-offset}, where
+the above parameters were used to model the responses at offsets $r=1.5,3,6$,
+and 12\,km. \added{We only need to compute 50\,\% of the required frequencies
+in the case of the FFTLog, and only 14\,\% in the case of the DLF.}
 %
@@
     $r=1.5,3,6$, and 12\,km using the parameters defined in Figure
-    \ref{fig:GUI-FFTLog-lin}. The coloured circles are the actually computed
-    responses, the black dots are the responses which are set to zero or
-    interpolated. The \replaced{grey}{black} curves are the analytical
+    \ref{fig:GUI-FFTLog-lin}. The coloured \replaced{symbols}{circles} are the
+    actually computed responses, the black dots are the responses which are set
+    to zero \added{(high frequencies)} or interpolated \added{(low
+    frequencies)}. The \replaced{grey}{black} curves are the analytical
     responses.}
@@
 computations is that the adaptive gridding tries to minimize the number of
-required cells while being fine enough and at the same time large enough to
-give accurate results. This is generally best done in a frequency and
-conductivity dependent manner.} To go from the model grid to the computational
-grid, we use the volume-averaging technique on logarithmic resistivities, as
-used in \cite{GEO.07.Plessix}. While this technique ensures that the total
-resistivity in the subsurface remains the same, it does not consider
-effective-medium theory \citep{GEO.03.Davydycheva}, for instance, the apparent
-anisotropy from a stack of finely layered formations of varying resistivity.
+required cells. This is generally best done in a frequency and conductivity
+dependent manner.} To go from the model grid to the computational grid, we use
+the volume-averaging technique on logarithmic resistivities, as used in
+\cite{GEO.07.Plessix}. While this technique ensures that the total resistivity
+in the subsurface remains the same, it does not consider effective-medium
+theory \citep{GEO.03.Davydycheva}, for instance, the apparent anisotropy from a
+stack of finely layered formations of varying resistivity.
 
@@
 response was chosen as this is the response currently implemented in
-\replaced{the framework}{\simpeg}.
+\replaced{it}{\simpeg}.
 
@@
 FFTLog, where the latter one requires usually much fewer frequencies to be
-interpolated.} We believe that our proposed improvements to the previously
-published methods make time-domain CSEM modelling with a frequency-domain code
-even more attractive than it was already before. We have demonstrated the idea
-of our Fourier transform method on CSEM data transformed from the frequency to
-the time domain. However, it could equally be applied to the transform from the
-time to the frequency domain and to other methods with similar
-characteristics.
+interpolated.} \deleted{We believe that our proposed improvements to the
+previously published methods make time-domain CSEM modelling with a
+frequency-domain code even more attractive than it was already before.} We have
+demonstrated the idea of our Fourier transform method on CSEM data transformed
+from the frequency to the time domain. However, it could equally be applied to
+the transform from the time to the frequency domain and to other methods with
+similar characteristics. \added{We believe that our proposed improvements to
+the previously published methods makes simulating results in one domain
+obtained through computations in the other domain followed by a transformation
+a viable alternative. The methodology is relatively simple to implement in any
+code.}
 

### 2020-12-01 ManuscriptTime.tex
Revision-I (#2)
* Revision I
* Update for matplotlib 3.3.0
* Change from Geophysics to GJI template

@@
     % onecolumn,   % comment for two columns
-    referee,     % uncomment for double lines
-    % final,       % uncomment for changes
+    % referee,     % uncomment for double lines
+    final,       % uncomment for changes
 ]{gji}
@@
 Looking carefully at these \deleted{three} points results in much smaller
-computation times with speedup factors of ten or more over previous results. A
-\deleted{frequency-domain} computation \replaced{followed by transformation}{of
-time-domain data} can therefore be competitive to \replaced{computation in the
-required }{time-}domain \deleted{solvers} if the required \replaced{evaluation
-points}{frequencies} and the corresponding grids are carefully chosen.
+computation times with speedup factors of ten or more over previous
+\replaced{methods}{results}. A \deleted{frequency-domain} computation
+\replaced{in one domain followed by transformation}{of time-domain data} can
+therefore be \replaced{an alternative}{competitive} to \replaced{computation in
+the other domain }{time-}domain \deleted{solvers} if the required
+\replaced{evaluation points}{frequencies} and the corresponding grids are
+carefully chosen.
 %
@@
 \citep{GRL.19.Gehrmann}, water \citep{GEO.05.Pedersen}, geothermal sources
-\citep{WGC.15.Girard}, or geological purposes \citep{NAT.19.Johanson}. While
-current sources with a few frequencies are used in the deep marine environment,
-transient measurements are more common in the shallow marine environment and on
-land \citep[e.g., ][]{GEO.07.Ziolkowski, SEG.07.Andreis, SEG.07.Avdeeva}. One
-of the main reasons is the dominance of the airwave in shallow marine and
-terrestrial measurements, which can be better separated in the time domain.
-CSEM is usually divided into frequency- and time-domain methods, depending
-whether the source signal is a continuous waveform, such as a sine, or a finite
-waveform, such as a pseudo-random binary sequence (PRBS). A numerical
-comparison of the two methods is given by \cite{GP.13.Conell}. Acquired CSEM
-data are subsequently often analysed (modelled and inverted for) in their
-respective domain, either the frequency or time domain. Modellers of layered
-media usually exploit the horizontal shift-invariance by computing the
-responses in the wavenumber-frequency domain followed by a 2D inverse spatial
-Fourier transform, also called Hankel transform, to the space-frequency domain,
-and a regular inverse Fourier transform if time-domain data are required
-\citep[e.g., ][]{GEO.15.Hunziker}.
-
+\citep{WGC.15.Girard}, or \added{for} geological purposes
+\citep{NAT.19.Johanson}. While current sources with a few frequencies are used
+in the deep marine environment, transient measurements are more common in the
+shallow marine environment and on land \citep[e.g., ][]{GEO.07.Ziolkowski,
+SEG.07.Andreis, SEG.07.Avdeeva}. One of the main reasons is the dominance of
+the airwave in shallow marine and terrestrial measurements, which can be better
+separated in the time domain. CSEM is usually divided into frequency- and
+time-domain methods, depending \added{on} whether the source signal is a
+continuous waveform, such as a sine, or a finite waveform, such as a
+pseudo-random binary sequence (PRBS). A numerical comparison of the two methods
+is given by \cite{GP.13.Conell}. Acquired CSEM data are subsequently often
+analysed (modelled and inverted for) in their respective domain, either the
+frequency or time domain. Modellers of layered media usually exploit the
+horizontal shift-invariance by computing the responses in the
+wavenumber-frequency domain followed by a 2D inverse spatial Fourier transform,
+also called Hankel transform, to the space-frequency domain, and a regular
+inverse Fourier transform if time-domain data are required \citep[e.g.,
+][]{GEO.15.Hunziker}.
+
+% a good review of the state of art 
 CSEM codes for 2D and 3D computations, on the other hand, mostly compute their
 responses directly in the required domain, either frequency or time. There is a
-wealth of 3D electromagnetic codes, with a first boom in the 1990's, with a
-good state of the art provided by \cite{B.SEG.99.Oristaglio}. A second wave
-occurred during the CSEM boom in the hydrocarbon industry in the early 21th
-century, see, e.g., \cite{SG.05.Avdeev} and \cite{SG.10.Borner} for
-comprehensive overviews. \replaced{Focusing on r}{R}ecent publications
-\added{after these review articles} span\added{s} the width of numerical
+wealth of 3D electromagnetic codes, with a first boom in the 1990's\replaced{;
+  \mbox{\cite{B.SEG.99.Oristaglio}} provide a good review of the state of
+art at that time}{, with a good state of the art provided by
+\mbox{\cite{B.SEG.99.Oristaglio}}}. A second wave occurred during the CSEM boom
+in the hydrocarbon industry in the early 21th century, see, e.g.,
+\cite{SG.05.Avdeev} and \cite{SG.10.Borner} for comprehensive overviews. Recent
+publications \added{after these review articles} span the width of numerical
 methods from integral equation codes \citep{MGS.17.Kruglyakov} and differential
@@
 grids \citep{CMMP.64.Lebedev} but also unstructured grids using, for instance,
-tetrahedral \citep{CAG.17.Cai}. The computational requirement to model CSEM
-data also depends heavily on the technique which is used to solve the system,
-where there are direct solvers \citep{GEO.15.Grayver}, indirect solvers
-\citep{GJI.15.Jaysaval}, and more recently also hybrid solvers
-\citep{GEO.18.Liu}.
+tetrahedra\deleted{l} \citep{CAG.17.Cai}. The computational requirement to
+model CSEM data also depends heavily on the technique \deleted{which is} used
+to solve the system\replaced{. T}{, where t}here are direct solvers
+\citep{GEO.15.Grayver}, indirect solvers \citep{GJI.15.Jaysaval}, and more
+recently also hybrid solvers \citep{GEO.18.Liu}.
 
@@
 \cite{GEO.08.Mulder}. \added{They compared the computational complexity of
-explicit time-stepping schemes (Du Fort-Frankel method as used in, e.g.,
-\mbox{\cite{GEO.04.Commer, GEO.07.Maao}}); implicit schemes \mbox{\citep[e.g.,
-][]{GEO.04.Haber}}, and matrix exponentials and Lanczos reduction schemes
-\mbox{\cite[e.g., ][]{RS.94.Druskin}} with the Fourier transform of
+explicit time-stepping schemes (Du Fort-Frankel method as used in, e.g.,}
+\cite{GEO.04.Commer, GEO.07.Maao}\added{); implicit schemes \mbox{\citep[e.g.,
+][]{GEO.04.Haber}}, and matrix exponentials and Lanczos reduction schemes}
+\cite[e.g., ][]{RS.94.Druskin} \added{with the Fourier transform of
 frequency-domain solutions. Their conclusion was that the Fourier transform
@@
 \added{While we use the finite-integration technique
-\mbox{\citep{AEU.77.Weiland}} in the frequency domain to show our method it is
-important to note that our ideas can be used with any solver, it is not bound
-to one or another frequency- or time-domain solver. The ideas for the frequency
-selection can be applied to time selection, and time-dependent, adaptive
-gridding exists as well \mbox{\citep[e.g., ][]{RS.06.Commer}}.} In the next
-section, we briefly review the methodology as introduced by
+\mbox{\citep{AEU.77.Weiland}} in the frequency domain to demonstrate our
+method, it is important to note that our ideas can be used with any solver and
+is not bound to one or another frequency- or time-domain solver. The ideas for
+the frequency selection can be applied to time selection, and time-dependent,
+adaptive gridding exists as well \mbox{\citep[e.g., ][]{RS.06.Commer}}.} In the
+next section, we briefly review the methodology as introduced by
 \cite{GEO.07.Plessix} and \cite{GEO.08.Mulder} and highlight their advantages
@@
 
-Being able to model CSEM data both in the frequency-domain and in the
-time-domain can be desirable, as both domains have advantages and
-disadvantages. One way to achieve this is to implement Maxwell's equation in
-both domains\deleted{, as it is done, for instance, in \simpeg (Cockett et al.,
-2015)}. Another approach is to have Maxwell's equation only implemented in one
-domain, and use Fourier transforms to go to the other. However, this approach
-can be costly, as many frequencies over a wide range are required to go from
-the time domain to the frequency domain, or many times over a wide range for
-the opposite direction. We present a methodology which significantly reduces
-the amount and range of the required frequencies, translating into a
-significant reduction in computation time. The required computation grids for
-low frequencies (in our case in the order of 0.001\,Hz) and high frequencies
+Being able to model CSEM data both in the frequency\replaced{ }{-}domain and in
+the time\replaced{ }{-}domain can be desirable, as both domains have advantages
+and disadvantages. One way to achieve this is to implement Maxwell's
+equation\added{s} in both domains\deleted{, as it is done, for instance, in
+\simpeg (Cockett et al., 2015)}. Another approach is to have Maxwell's
+equation\added{s} only implemented in one domain, and use Fourier transforms to
+go to the other. However, this approach can be costly, as many frequencies over
+a wide range are required to go from the \replaced{frequency}{time} domain to
+the \replaced{time}{frequency} domain, or many times over a wide range for the
+opposite direction. We present a methodology which significantly reduces the
+amount and range of the required frequencies, translating into a significant
+reduction in computation time. The required computation grids for low
+frequencies (in our case in the order of 0.001\,Hz) and high frequencies
 (around 100\,Hz) are hugely different. Low frequencies can be computed on a
@@
 conductivity and frequency, and for the diffusive approximation of Maxwell's
-equation in an isotropic, homogeneous medium is given by \citep[e.g.,][
-equation~1.53]{B.SEG.88.Ward}
+equation\added{s} in an isotropic, homogeneous medium is given by
+\citep[e.g.,][ equation~1.53]{B.SEG.88.Ward}
 %
@@
 $z$-direction to account for the airwave. To reduce the number of cells, it is
-desirable to introduce stretching, at least in the boundary zone outside of the
-area where source and receivers are located.
+desirable to introduce stretching, at least in the \replaced{buffer}{boundary}
+zone outside of the area where source and receivers are located.
 
@@
 methods to model transient electromagnetic responses directly in the
-time-domain, and compare it to the computation of transient EM responses in the
-frequency-domain with a subsequent Fourier transform. They conclude the review
-by stating \emph{\guillemotleft Although it remains to be seen which of the
-four methods requires the least computer time for a given accuracy, the
-frequency-domain approach appears to be attractive.\guillemotright} Their
-approach is to minimize the computation time by having, in addition to the just
-introduced adaptive gridding, an adaptive frequency selection scheme. This
-scheme starts with computing the responses for a set of just five frequencies,
-regularly sampled on a log-scale, from minimum to maximum required frequency.
-All the other frequencies are interpolated with a shape-preserving
-piecewise-cubic Hermite interpolation \citep[PCHIP, ][]{SIAM.80.Fritsch}.
-Testing the stability of the obtained response by removing a single
-frequency-value at a time their scheme decides if more frequencies in-between
-the already computed ones are required. In this way frequencies are only added
-if required, hence if certain criteria of response stability are not met. While
-this method is good and effective for a single offset, it loses all its
-advantages if one tries to compute different offsets within one computation, as
-each offset requires a different set of adaptive frequencies. Additionally, it
-hampers the parallelization over frequencies.
+time\replaced{ }{-}domain, and compare it to the computation of transient EM
+responses in the frequency\replaced{ }{-}domain with a subsequent Fourier
+transform. They conclude the review by stating \emph{\guillemotleft Although it
+remains to be seen which of the four methods requires the least computer time
+for a given accuracy, the frequency-domain approach appears to be
+attractive.\guillemotright} Their approach is to minimize the computation time
+by having, in addition to the just introduced adaptive gridding, an adaptive
+frequency selection scheme. This scheme starts with computing the responses for
+a set of just five frequencies, regularly sampled on a log-scale, from minimum
+to maximum required frequency. All the other frequencies are interpolated with
+a shape-preserving piecewise-cubic Hermite interpolation \citep[PCHIP,
+][]{SIAM.80.Fritsch}. Testing the stability of the obtained response by
+removing a single frequency-value at a time their scheme decides if more
+frequencies in-between the already computed ones are required. In this way
+frequencies are only added if required, hence if certain criteria of response
+stability are not met. While this method is good and effective for a single
+offset, it loses all its advantages if one tries to compute different offsets
+within one computation, as each offset requires a different set of adaptive
+frequencies. Additionally, it hampers the parallelization over frequencies.
 
@@
 frequency domain to time domain, which makes time-domain modelling with a
-frequency-domain code even more competitive in comparison with time-domain
-codes. It is important to note that while we show the transform from frequency
-to time domain, the conclusions regarding the Fourier transform can equally
-well be applied to the transform from time to frequency domain data for a code
-that solves Maxwell's equations in the time domain. A variation of our method
-was implemented, for instance, by \custem \citep{GEO.19.Rochlitz} to obtain
-frequency-domain data from time-domain computations (Rochlitz, private
-communication, 2020).
+frequency-domain code \replaced{viable}{even more competitive} in comparison
+with time-domain codes. It is important to note that while we show the
+transform from frequency to time domain, the conclusions regarding the Fourier
+transform can equally well be applied to the transform from time\added{-domain}
+to frequency\replaced{-}{ }domain data for a code that solves Maxwell's
+equations in the time domain. A variation of our method was implemented, for
+instance, by \custem \citep{GEO.19.Rochlitz} to obtain frequency-domain data
+from time-domain computations (Rochlitz, private communication, 2020).
 
@@
 solvers or as a solver on its own. The multigrid approach works fine for the
-diffusive approximation of Maxwell's equation, which assumes that
+diffusive approximation of Maxwell's equation\added{s}, which assumes that
 $\omega\varepsilon \ll \sigma$, where $\varepsilon$ is electric permittivity
@@
 
-\section{Frequency selection} % % % % % % % % % % % % % % % % % % % % % % % % %
+\subsection{Frequency selection}  % % % % % % % % % % % % % % % % % % % % % % %
 
@@
 fewer times required, the quicker we obtain frequency-domain results. Even
-though we limit our analysis here to frequency-to-time transformation, most of
-the arguments are reversible and applicable to time-to-frequency
+though we limit our analysis here to frequency-to-time transformation\added{s},
+most of the arguments are reversible and applicable to time-to-frequency
 transformations.} We decided to use a regular spacing of frequencies on a
@@
 and below 0.3\,s. It also shows the oscillating high-frequency part, which is
-hard to interpolate\added{ and we therefore set to zero}. Figure
-\ref{fig:GUI-DLF-lin} is the same as \ref{fig:GUI-FFTLog-lin}, but transformed
-with the DLF method applying the 81-point sine-cosine filter from
-\cite{GEO.09.Key}. The same frequencies were computed as in the FFTLog case and
-the missing ones interpolated. The error of the corresponding time-domain
-response is comparable, so either FFTLog or DLF can be used.
+hard to interpolate\added{ and we therefore set it to zero}. Figure
+\ref{fig:GUI-DLF-lin} is the same as \added{Figure} \ref{fig:GUI-FFTLog-lin},
+but transformed with the DLF method applying the 81-point sine-cosine filter
+from \cite{GEO.09.Key}. The same frequencies were computed as in the FFTLog
+case and the missing ones interpolated. The error of the corresponding
+time-domain response is comparable, so either FFTLog or DLF can be used.
 %
@@
   scale. (b) Using DLF instead of FFTLog, but computing the same
-  frequenc\replaced{y range}{ies} as for the FFTLog. The resulting time domain
-  response has comparable accuracy (compare the error to the error in Figure
-  \ref{fig:GUI-FFTLog-lin}).}
+  frequenc\replaced{y range}{ies} as for the FFTLog. The resulting
+  time\replaced{-}{ }domain response has comparable accuracy (compare the error
+  to the error in Figure \ref{fig:GUI-FFTLog-lin}).}
   \label{fig:GUI-log-lin}
@@
 wide range of offsets. This is shown in Figure \ref{fig:multi-offset}, where
-the above parameters were used to model the responses at offsets $r=1.5,3,6$,
-and 12\,km. \added{We only need to compute 50\,\% of the required frequencies
-in the case of the FFTLog, and only 14\,\% in the case of the DLF.}
+the \replaced{same kernels were used to yield}{above parameters were used to
+model} the responses at offsets $r=1.5,3,6$, and 12\,km. \added{We only need to
+compute 50\,\% of the required frequencies in the case of the FFTLog, and
+only 14\,\% in the case of the DLF.}
 %
@@
 DLF method we also have to interpolate in between the computed frequencies, for
-which we found it better to use a cubic spline. As can be seen from Figures
-\ref{fig:GUI-FFTLog-lin} and \ref{fig:GUI-DLF-lin}, using the FFTLog or the DLF
-with the same actually computed frequencies results in very similar responses.
+which we found it better to use a cubic spline. As can be seen \replaced{by
+comparing}{from} Figure\deleted{s} \ref{fig:GUI-FFTLog-lin} \replaced{with
+Figure}{and} \ref{fig:GUI-DLF-lin}, using the FFTLog or the DLF with the same
+actually computed frequencies results in very similar responses.
 
@@
 takes much longer than the transform itself. What matters is solely how many
-frequencies are required by it in order to be stable, and how long it takes to
-compute the responses for these frequencies.
-
-\section{Gridding}  % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
-
-The computation grid consists of a core or survey domain $D_\mr{s}$ which
-should contain all source and receiver positions. The survey domain usually has
-no or very little cell stretching $\alpha_\mr{s}$. The minimum cell width is
-defined as
+frequencies are required by it \replaced{to achieve the desired precision}{in
+order to be stable}, and how long it takes to compute the responses for these
+frequencies.
+
+\subsection{Gridding}   % % % % % % % % % % % % % % % % % % % % % % % % % % % %
+
+The computation grid consists of a core or survey domain $D_\mr{s}$
+\replaced{that}{which} should contain all source and receiver positions. The
+survey domain usually has no or very little cell stretching \added{factor}
+$\alpha_\mr{s}$. The minimum cell width is defined as
 %
@@
 resides, and $n_\delta$ is a positive number that defines how many cells there
-should be per skin depth. The actual computation domain $D_\mr{c}$ is usually
-much bigger than $D_\mr{s}$ in order to avoid artefacts from the perfectly
-electrically conducting boundary condition. It can also have a much higher
-stretching $\alpha_\mr{c}$. In \replaced{the presented examples}{our scheme} we
-have chosen $D_\mr{c}$ such that the distance for the signal diffusing from the
-source to the boundary and back to the receiver closest to the boundary is at
-least two wavelengths, after which the initial signal is reduced to a millionth
-of its strength. The wavelength \added{$\lambda$} to compute $D_\mr{c}$ is
-given by
+should be per skin depth. The actual computation\added{al} domain $D_\mr{c}$ is
+usually much bigger than $D_\mr{s}$ in order to avoid artefacts from the
+perfectly electrically conducting boundary condition. It can also have a much
+higher stretching \added{factor} $\alpha_\mr{c}$. In \replaced{the presented
+examples}{our scheme} we have chosen $D_\mr{c}$ such that the distance for the
+signal diffusing from the source to the boundary and back to the receiver
+closest to the boundary is at least two wavelengths, after which the
+\replaced{amplitude of the}{initial} signal is reduced to a millionth of its
+\added{initial} strength. The wavelength \added{$\lambda$} to compute
+$D_\mr{c}$ is given by
 %
@@
 computational domain is therefore enforced, defining the maximum distance from
-the source to the boundary; this distance is by default 100\,km, but this can
-be reduced in the marine case with increasing water depth. Note that for
-shallow marine and land cases this also applies to the horizontal dimensions,
-not only to the upward $z$-direction and similarly for very resistive
-basements, even in deep water. One way to circumvent this difficulty is the use
-of a primary-secondary formulation, where the primary field, including the air
-wave, is computed with a semi-analytical code for layered media. We do not
-consider this approach here.
+the source to the boundary; this distance is by default \added{set to} 100\,km,
+but this can be reduced in the marine case with increasing water depth. Note
+that for shallow marine and land cases this also applies to the horizontal
+dimensions, not only to the upward $z$-direction and similarly for very
+resistive basements, even in deep water. One way to circumvent this difficulty
+is the use of a primary-secondary formulation, where the primary field,
+including the air wave, is computed with a semi-analytical code for layered
+media. We do not consider this approach here.
 
@@
 $\Delta_\mr{min}$ can further be restricted by a user-defined range. Given
-these inputs the adaptive gridding will search for the smallest possible number
-of cells which fulfils these criteria. The \added{implemented} multigrid method
-\deleted{implemented in \emg3d} puts some constraints on the number of cells, of
-which the adaptive gridding takes care (the number of cells have to be powers
-of two multiplied by a low prime, e.g., $\{2,3,5\}\cdot2^n$).
+these inputs\added{,} the adaptive gridding will search for the smallest
+possible number of cells which fulfils these criteria. The \added{implemented}
+multigrid method \deleted{implemented in \emg3d} puts some constraints on the
+number of cells, of which the adaptive gridding takes care (the number of cells
+have to be powers of two multiplied by a low prime, e.g., $\{2,3,5\}\cdot2^n$).
 
@@
 criteria, for a fixed number of cells. Our adaptive gridding, on the other
-hand, searches for the smallest number of cells that still fulfil the given
-criteria. The number of cells becomes therefore also a function of frequency.
-\added{It is important to note that this is our implementation of an adaptive
-grid, but there are certainly other possibilities. The relevant point for fast
-computations is that the adaptive gridding tries to minimize the number of
-required cells. This is generally best done in a frequency and conductivity
-dependent manner.} To go from the model grid to the computational grid, we use
-the volume-averaging technique on logarithmic resistivities, as used in
-\cite{GEO.07.Plessix}. While this technique ensures that the total resistivity
-in the subsurface remains the same, it does not consider effective-medium
-theory \citep{GEO.03.Davydycheva}, for instance, the apparent anisotropy from a
-stack of finely layered formations of varying resistivity.
+hand, searches for the smallest number of cells that still fulfil\added{s} the
+given criteria. The number of cells becomes therefore also a function of
+frequency. \added{It is important to note that this is our implementation of an
+adaptive grid, but there are certainly other possibilities. The relevant point
+for fast computations is that the adaptive gridding tries to minimize the
+number of required cells. This is generally best done in a frequency\added{-}
+and conductivity\replaced{-}{ }dependent manner.} To go from the model grid to
+the computational grid, we use the volume-averaging technique on logarithmic
+resistivities, as used in \cite{GEO.07.Plessix}. While this technique ensures
+that the total resistivity in the subsurface remains the same, it does not
+consider effective-medium theory \citep{GEO.03.Davydycheva}, for instance, the
+apparent anisotropy from a stack of finely layered formations of varying
+resistivity.
 
@@
 compute from 0.05\,Hz to 20.0\,Hz. The complete frequency range for the
-transform includes 30 frequencies from 0.0002\,Hz to 126.4\,Hz. For the
-adaptive gridding the following inputs were used: $n_\delta=12$, minimum cell
-width must be between 20 and 40\,m, and $\alpha_\mr{s}=1$,
-$\alpha_\mr{c}=[1,1.3]$. This created grids with cell numbers between
-\num{46080} ($80\times24\times24$, for 20.0\,Hz) and \num{128000}
-($80\times40\times40$, for 0.05\,Hz) cells. The run times for each frequency,
-the corresponding number of cells, minimum cell width and computation domain
-stretching factor are listed in Table~\ref{tbl:timefull}. The total run time to
-compute this model was less than two minutes.
+transform\added{, including the frequencies for which we use interpolation,}
+includes 30 frequencies from 0.0002\,Hz to 126.4\,Hz. For the adaptive gridding
+the following inputs were used: $n_\delta=12$, minimum cell width must be
+between 20 and 40\,m, and $\alpha_\mr{s}=1$, $\alpha_\mr{c}=[1,1.3]$. This
+\replaced{yields}{created} grids with cell numbers between \num{46080}
+($80\times24\times24$, for 20.0\,Hz) and \num{128000} ($80\times40\times40$,
+for 0.05\,Hz) cells. The run times for each frequency, the corresponding number
+of cells, minimum cell width\added{,} and computation domain stretching factor
+are listed in Table~\ref{tbl:timefull}. The total run time to compute this
+model was less than two minutes.
 %
@@
   \caption{Electric inline response at an offset of 5\,km for a shallow marine,
-    layered scenario. (a) Frequency-domain, where the blue circles are computed
-    responses and the black dots are interpolated responses or responses set to
-    zero. (b) Time-domain response.}
+  layered scenario. (a) Frequency-domain\added{ response}, where the blue
+  circles \replaced{denote}{are} computed responses and the black dots
+  \deleted{are} interpolated responses or responses set to zero. (b)
+  Time-domain response.}
   \label{fig:marine}
@@
 frequency-domain computations and the same frequencies for the Fourier
-transform. The computation of these frequencies took less than 13 minutes and
-it handles any offset between 3 and 7\,km. It can be seen that the chosen
-frequency selection is sufficient for this offset range; again, more low
+transform. The computation of these frequencies took less than \replaced{9}{13}
+minutes and it handles any offset between 3 and 7\,km. It can be seen that the
+chosen frequency selection is sufficient for this offset range; again, more low
 frequencies could be added to improve late-time values.
@@
 spreading. On land and in shallow marine scenarios one has therefore to include
-a sufficiently large computational domain, and the default in our scheme is
-100\,km. The important point is that this does not only apply to the upward
-$z$-direction, but also to the horizontal directions, as the airwave also
-bounces back horizontally and would continuously emit energy into the
-subsurface if the boundaries are not chosen far enough away from the receivers.
-If models are computed with very resistive layers or models with highly
-resistive basements, this can even apply to deep marine scenarios.
+a sufficiently large computational domain\replaced{. T}{, and t}he default in
+our scheme is 100\,km. The important point is that this does not only apply to
+the upward $z$-direction, but also to the horizontal directions, as the airwave
+\deleted{also} bounces back horizontally and would continuously emit energy
+into the subsurface if the boundaries are not chosen far enough away from the
+receivers. If models are computed with very resistive layers or models with
+highly resistive basements, this can even apply to deep marine scenarios.
 
@@
 the previous section; however, for the adaptive gridding in the horizontal
-directions $\rho_\mr{ave}=1\,\ohmm$ was used instead of
+directions\added{,} $\rho_\mr{ave}=1\,\ohmm$ was used instead of
 $\rho_\mr{ave}=\num{10000}\,\ohmm$. Having the boundaries too near in the
@@
 It can be difficult to spot these errors in the time-domain result, as the
-response looks plausible and only comparing it with the 1D result reveals that
-it is actually wrong. A possibility to detect such problems for complicated
-cases, where there is no semi-analytical result to compare with, is to compute
-two or more models, moving the boundary. When the responses stop to change, one
-can assume that the boundary is far enough. Another possibility is to look at
-the amplitudes close to the boundaries and ensure that they are small enough.
+response looks plausible\replaced{. Only a comparison}{ and only comparing it}
+with the 1D result reveals that it is actually wrong. A possibility to detect
+such problems for complicated cases, where there is no semi-analytical result
+to compare with, is to compute two or more models, moving the boundary. When
+the responses stop to change, one can assume that the boundary is far enough.
+Another possibility is to look at the amplitudes close to the boundaries and
+ensure that they are small enough.
 
@@
 otherwise be too high. There are not many options out there of open-source
-time-domain 3D codes, \simpeg being the one we found to be suitable. A step-off
-response was chosen as this is the response currently implemented in
-\replaced{it}{\simpeg}.
+time-domain 3D \added{CSEM} codes, \simpeg being the one we found to be
+suitable. A step-off response was chosen as this is the response currently
+implemented in \replaced{it}{\simpeg}.
 
@@
 %
-The background comparison shows that both 3D codes do an acceptable job having
-a relative error of a few percents at most; the result \replaced{obtained
-through transformation}{for \emg3d} seems to be better at early times.
-\added{The reason is probably the implemented backward Euler in the time-domain
-code which has an error of order one in time.} We cannot compare the errors for
-the response that includes the target for lack of an analytical solution. The
-\added{1D} background model \deleted{from \empymod }is only included to show
-that there is a significant response from the target. We therefore show the
-normalized difference (\replaced{NRMSD}{ND}) between the two
-\replaced{responses $R_1$ and $R_2$}{3D codes} as a percentage, where
+The background comparison shows that both 3D codes do an acceptable job
+\replaced{with}{having} a relative error of a few percents at most; the result
+\replaced{obtained through transformation}{for \emg3d} seems to be better at
+early times. \added{The reason is probably the implemented backward Euler
+scheme in the time-domain code that has an error of order one in time.} We
+cannot compare the errors for the response that includes the target for lack of
+an analytical solution. The \added{1D} background model \deleted{from \empymod
+}is only included to show that there is a significant response from the target.
+We therefore show the normalized difference (\replaced{NRMSD}{ND}) between the
+two \replaced{responses $R_1$ and $R_2$}{3D codes} as a percentage, where
 \replaced{$\mr{NRMSD~(\%)} = 200|R_1 - R_2|/(|R_1| + |R_2|)$.}{$\mr{ND}=
@@
 for frequencies in-between the computed ones are best obtained with a spline
-interpolation. The actual transform can be made with either the DLF method or
-FFTLog, where the latter one requires usually much fewer frequencies to be
-interpolated.} \deleted{We believe that our proposed improvements to the
+interpolation. The actual transform can be carried out with either the DLF
+method or FFTLog, where the latter one requires usually much fewer frequencies
+to be interpolated.} \deleted{We believe that our proposed improvements to the
 previously published methods make time-domain CSEM modelling with a
@@
 demonstrated the idea of our Fourier transform method on CSEM data transformed
-from the frequency to the time domain. However, it could equally be applied to
-the transform from the time to the frequency domain and to other methods with
-similar characteristics. \added{We believe that our proposed improvements to
-the previously published methods makes simulating results in one domain
-obtained through computations in the other domain followed by a transformation
-a viable alternative. The methodology is relatively simple to implement in any
-code.}
+from the frequency \added{domain} to the time domain. However, it could equally
+be applied to the transform from the time \added{domain} to the frequency
+domain and to other methods with similar characteristics. \added{We believe
+that our proposed improvements to the previously published methods makes
+simulating results in one domain obtained through computations in the other
+domain followed by a transformation a viable alternative. The methodology is
+relatively simple to implement in any code.}
 
@@
 (ERA-NET Cofund, \href{https://www.martera.eu}{martera.eu}).
-An early idea of this manuscript was presented at \cite{EAGE.20.Werthmuller}.
+An early idea of this manuscript was presented \replaced{by}{at}
+\cite{EAGE.20.Werthmuller}.
 \end{acknowledgments}

### 2020-12-01 notebooks/freqselect.py

@@ -319,13 +319,19 @@ class InteractiveFrequency(emg3d.utils.Fourier):
         else:
             self.axs[1].set_xlim([min(self.time), max(self.time)])
 
-    def print_suptitle(self):
+    def print_legend(self):
         """Update suptitle."""
-        plt.suptitle(
-            f"Offset = {np.squeeze(self.model['rec'][0])/1000} km;    "
-            f"No. freq. coarse: {self.freq_calc.size};    No. freq. full: "
-            f"{self.freq_req.size}  ({self.freq_req.min():.1e} $-$ "
-            f"{self.freq_req.max():.1e} Hz)")
+        # plt.suptitle(
+        #     f"Offset = {np.squeeze(self.model['rec'][0])/1000} km;    "
+        #     f"No. freq. coarse: {self.freq_calc.size};    No. freq. full: "
+        #     f"{self.freq_req.size}  ({self.freq_req.min():.1e} $-$ "
+        #     f"{self.freq_req.max():.1e} Hz)")
+        self.clear_handle(['f_legend', ])
+        self.h_f_legend = self.axs[0].legend(
+            handles=[self.h_f_inti, self.h_f_int],
+            labels=[f"required ({self.freq_req.size})",
+                    f"computed ({self.freq_calc.size})"]
+        )
 
     def plot_base_model(self):
         """Update smooth, 'correct' model."""
@@ -375,8 +381,8 @@ class InteractiveFrequency(emg3d.utils.Fourier):
         self.h_t_int, = self.axs[1].plot(self.time, t_int, 'k--')
         self.h_t_inte, = self.axs[3].plot(self.time, t_error, 'k.')
 
-        # Update suptitle
-        self.print_suptitle()
+        # Update legend
+        self.print_legend()
 
     # Interactive routines
     def update_off(self, off):
@@ -454,7 +460,7 @@ class InteractiveFrequency(emg3d.utils.Fourier):
         if linlog == 'log':
             sym_dec = 10  # Number of decades to show on symlog
             lty = int(max(np.log10(abs(self.reim(self.f_dense))))-sym_dec)
-            self.axs[0].set_yscale('symlog', linthreshy=10**lty, linscaley=0.7)
+            self.axs[0].set_yscale('symlog', linthresh=10**lty, linscale=0.7)
 
             # Remove the zero line becouse of the overlapping ticklabels.
             nticks = len(self.axs[0].get_yticks())//2

### 2020-12-01 ManuscriptTime.tex
Clean diff

--- 2020-11-28 ManuscriptTime.tex
@@
 
-\title[\replaced{Transforming expensive EM kernels}{Time-domain CSEM}]{
-  \replaced{Fast Fourier transformation of electromagnetic data for
-  computationally expensive kernels}{Fast time-domain electromagnetic modelling
-  in the frequency domain}}
+\title[Transforming expensive EM kernels]{Fast Fourier transformation of
+electromagnetic data for computationally expensive kernels}
 
@@
 %
-\deleted{Modelling three-dimensional time-domain controlled-source
-electromagnetic data with frequen\-cy-domain solvers is an alternative to
-time-domain solvers.} \added{Three-dimensional controlled-source
-electromagnetic data is often computed directly in the domain of interest,
-either in the frequency domain or in the time domain. Computing it in one
-domain and transforming it via a Fourier transform to the other domain is a
-viable alternative.} It requires the \replaced{evaluation}{computation} of many
-\replaced{responses in the computational domain}{frequencies} if standard
-Fourier transforms are used\replaced{. This}{, which} can make it prohibitively
-expensive \replaced{if the kernel is time-consuming as is the case in
-three-dimensional electromagnetic modelling}{in comparison with a time-domain
-computation}. The speed of \replaced{such}{time-domain} modelling
-\replaced{that is obtained through a transform}{with frequency-domain
-computation} is defined by three key points: solver, method and implementation
-of the Fourier transformation, and gridding.
-%
-The faster the \deleted{frequency-domain} solver, the faster
-\deleted{time-domain} modelling will be. It is important that the solver is
-robust over a wide range of \added{values (}frequencies\added{ or times)}.
-\deleted{A three-dimensional, iterative, frequency-domain multigrid solver
-satisfies this re\-quirement, but any other robust solver will do.}
-%
-The method should require as few \replaced{kernel evaluations}{frequencies} as
-possible while remaining robust. As the frequency \added{and time}
-range\added{s} span\deleted{s} many orders of magnitude, the required
-\replaced{values}{frequencies} are ideally equally spaced on a logarithmic
-scale. The \added{proposed} fast \deleted{frequency-to-time domain} method uses
-either the digital linear filter method or the logarithmic fast Fourier
-transform together with a careful selection of \replaced{evaluation
-points}{frequencies} and interpolation. \replaced{In frequency-to-time domain
-tests t}{T}his methodology requires typically 15 to 20 frequencies to cover a
-wide range of offsets\deleted{in one-dimensional test models}.
-%
-The gridding should be frequency-\added{ or time-}dependent, which is
-accomplished by making it a function of skin depth. Optimising for the least
-number of required cells \deleted{for a given frequency} should be combined
-with optimising for computational speed.
-%
-Looking carefully at these \deleted{three} points results in much smaller
-computation times with speedup factors of ten or more over previous results. A
-\deleted{frequency-domain} computation \replaced{followed by transformation}{of
-time-domain data} can therefore be competitive to \replaced{computation in the
-required }{time-}domain \deleted{solvers} if the required \replaced{evaluation
-points}{frequencies} and the corresponding grids are carefully chosen.
+Three-dimensional controlled-source electromagnetic data is often computed
+directly in the domain of interest, either in the frequency domain or in the
+time domain. Computing it in one domain and transforming it via a Fourier
+transform to the other domain is a viable alternative. It requires the
+evaluation of many responses in the computational domain if standard Fourier
+transforms are used. This can make it prohibitively expensive if the kernel is
+time-consuming as is the case in three-dimensional electromagnetic modelling.
+The speed of such modelling that is obtained through a transform is defined by
+three key points: solver, method and implementation of the Fourier
+transformation, and gridding.
+%
+The faster the solver, the faster modelling will be. It is important that the
+solver is robust over a wide range of values (frequencies or times). The method
+should require as few kernel evaluations as possible while remaining robust. As
+the frequency and time ranges span many orders of magnitude, the required
+values are ideally equally spaced on a logarithmic scale. The proposed fast
+method uses either the digital linear filter method or the logarithmic fast
+Fourier transform together with a careful selection of evaluation points and
+interpolation. In frequency-to-time domain tests this methodology requires
+typically 15 to 20 frequencies to cover a wide range of offsets.
+%
+The gridding should be frequency- or time-dependent, which is accomplished by
+making it a function of skin depth. Optimising for the least number of required
+cells should be combined with optimising for computational speed. Looking
+carefully at these points results in much smaller computation times with
+speedup factors of ten or more over previous methods. A computation in one
+domain followed by transformation can therefore be an alternative to
+computation in the other domain domain if the required evaluation points and
+the corresponding grids are carefully chosen.
 %
@@
 \begin{keywords}
-  Controlled source electromagnetics (CSEM); Numerical modelling; Fourier analysis.
+  Controlled source electromagnetics (CSEM); Numerical modelling; Fourier
+  analysis.
 \end{keywords}
@@
 \citep{GRL.19.Gehrmann}, water \citep{GEO.05.Pedersen}, geothermal sources
-\citep{WGC.15.Girard}, or geological purposes \citep{NAT.19.Johanson}. While
-current sources with a few frequencies are used in the deep marine environment,
-transient measurements are more common in the shallow marine environment and on
-land \citep[e.g., ][]{GEO.07.Ziolkowski, SEG.07.Andreis, SEG.07.Avdeeva}. One
-of the main reasons is the dominance of the airwave in shallow marine and
-terrestrial measurements, which can be better separated in the time domain.
-CSEM is usually divided into frequency- and time-domain methods, depending
-whether the source signal is a continuous waveform, such as a sine, or a finite
-waveform, such as a pseudo-random binary sequence (PRBS). A numerical
-comparison of the two methods is given by \cite{GP.13.Conell}. Acquired CSEM
-data are subsequently often analysed (modelled and inverted for) in their
-respective domain, either the frequency or time domain. Modellers of layered
-media usually exploit the horizontal shift-invariance by computing the
-responses in the wavenumber-frequency domain followed by a 2D inverse spatial
-Fourier transform, also called Hankel transform, to the space-frequency domain,
-and a regular inverse Fourier transform if time-domain data are required
-\citep[e.g., ][]{GEO.15.Hunziker}.
+\citep{WGC.15.Girard}, or for geological purposes \citep{NAT.19.Johanson}.
+While current sources with a few frequencies are used in the deep marine
+environment, transient measurements are more common in the shallow marine
+environment and on land \citep[e.g., ][]{GEO.07.Ziolkowski, SEG.07.Andreis,
+SEG.07.Avdeeva}. One of the main reasons is the dominance of the airwave in
+shallow marine and terrestrial measurements, which can be better separated in
+the time domain. CSEM is usually divided into frequency- and time-domain
+methods, depending on whether the source signal is a continuous waveform, such
+as a sine, or a finite waveform, such as a pseudo-random binary sequence
+(PRBS). A numerical comparison of the two methods is given by
+\cite{GP.13.Conell}. Acquired CSEM data are subsequently often analysed
+(modelled and inverted for) in their respective domain, either the frequency or
+time domain. Modellers of layered media usually exploit the horizontal
+shift-invariance by computing the responses in the wavenumber-frequency domain
+followed by a 2D inverse spatial Fourier transform, also called Hankel
+transform, to the space-frequency domain, and a regular inverse Fourier
+transform if time-domain data are required \citep[e.g., ][]{GEO.15.Hunziker}.
 
@@
 responses directly in the required domain, either frequency or time. There is a
-wealth of 3D electromagnetic codes, with a first boom in the 1990's, with a
-good state of the art provided by \cite{B.SEG.99.Oristaglio}. A second wave
-occurred during the CSEM boom in the hydrocarbon industry in the early 21th
-century, see, e.g., \cite{SG.05.Avdeev} and \cite{SG.10.Borner} for
-comprehensive overviews. \replaced{Focusing on r}{R}ecent publications
-\added{after these review articles} span\added{s} the width of numerical
-methods from integral equation codes \citep{MGS.17.Kruglyakov} and differential
-equation methods such as finite differences
-\replaced{\mbox{\citep{GEO.10.Mittet, GEO.11.Zaslavsky, CAG.13.Sommer}}}
-{\mbox{\citep{CAG.13.Sommer}}}, finite elements \citep{GJI.13.Grayver}, and
-finite volumes \citep{GEO.14.Jahandari}. Besides the different methods there
-are also many different types of discretisation. Commonly used, besides
-rectilinear Yee grids \citep{IEEE.66.Yee} as used by our code, are Lebedev
-grids \citep{CMMP.64.Lebedev} but also unstructured grids using, for instance,
-tetrahedral \citep{CAG.17.Cai}. The computational requirement to model CSEM
-data also depends heavily on the technique which is used to solve the system,
-where there are direct solvers \citep{GEO.15.Grayver}, indirect solvers
+wealth of 3D electromagnetic codes, with a first boom in the 1990's;
+\cite{B.SEG.99.Oristaglio} provide a good review of the state of art at that
+time. A second wave occurred during the CSEM boom in the hydrocarbon industry
+in the early 21th century, see, e.g., \cite{SG.05.Avdeev} and
+\cite{SG.10.Borner} for comprehensive overviews. Recent publications after
+these review articles span the width of numerical methods from integral
+equation codes \citep{MGS.17.Kruglyakov} and differential equation methods such
+as finite differences \citep{GEO.10.Mittet, GEO.11.Zaslavsky, CAG.13.Sommer},
+finite elements \citep{GJI.13.Grayver}, and finite volumes
+\citep{GEO.14.Jahandari}. Besides the different methods there are also many
+different types of discretisation. Commonly used, besides rectilinear Yee grids
+\citep{IEEE.66.Yee} as used by our code, are Lebedev grids
+\citep{CMMP.64.Lebedev} but also unstructured grids using, for instance,
+tetrahedra \citep{CAG.17.Cai}. The computational requirement to model CSEM data
+also depends heavily on the technique used to solve the system. There are
+direct solvers \citep{GEO.15.Grayver}, indirect solvers
 \citep{GJI.15.Jaysaval}, and more recently also hybrid solvers
@@
 
-Under certain conditions it can be \replaced{a viable alternative}{competitive}
-to model time-domain data with a frequency-domain code, as shown by
-\cite{GEO.08.Mulder}. \added{They compared the computational complexity of
-explicit time-stepping schemes (Du Fort-Frankel method as used in, e.g.,
-\mbox{\cite{GEO.04.Commer, GEO.07.Maao}}); implicit schemes \mbox{\citep[e.g.,
-][]{GEO.04.Haber}}, and matrix exponentials and Lanczos reduction schemes
-\mbox{\cite[e.g., ][]{RS.94.Druskin}} with the Fourier transform of
-frequency-domain solutions. Their conclusion was that the Fourier transform
-method can be favourable if the number of required frequencies is small
-relative to the number of grid points in each coordinate.} The conditions
-\added{for fast computation of time-domain data with a frequency-domain code}
-are: a sufficiently powerful solver, appropriate frequency selection and
+Under certain conditions it can be a viable alternative to model time-domain
+data with a frequency-domain code, as shown by \cite{GEO.08.Mulder}. They
+compared the computational complexity of explicit time-stepping schemes (Du
+Fort-Frankel method as used in, e.g., \cite{GEO.04.Commer, GEO.07.Maao});
+implicit schemes \citep[e.g., ][]{GEO.04.Haber}, and matrix exponentials and
+Lanczos reduction schemes \cite[e.g., ][]{RS.94.Druskin} with the Fourier
+transform of frequency-domain solutions. Their conclusion was that the Fourier
+transform method can be favourable if the number of required frequencies is
+small relative to the number of grid points in each coordinate. The conditions
+for fast computation of time-domain data with a frequency-domain code are: a
+sufficiently powerful solver, appropriate frequency selection and
 interpolation, and an automated gridding, for which they used the
@@
 
-\added{While we use the finite-integration technique
-\mbox{\citep{AEU.77.Weiland}} in the frequency domain to show our method it is
-important to note that our ideas can be used with any solver, it is not bound
-to one or another frequency- or time-domain solver. The ideas for the frequency
-selection can be applied to time selection, and time-dependent, adaptive
-gridding exists as well \mbox{\citep[e.g., ][]{RS.06.Commer}}.} In the next
-section, we briefly review the methodology as introduced by
-\cite{GEO.07.Plessix} and \cite{GEO.08.Mulder} and highlight their advantages
-and shortcomings. This is followed by an outline of our changes to the Fourier
-transform and the adaptive gridding. Finally, we demonstrate the efficiency of
-the approach with some numerical results.
+While we use the finite-integration technique \citep{AEU.77.Weiland} in the
+frequency domain to demonstrate our method, it is important to note that our
+ideas can be used with any solver and is not bound to one or another frequency-
+or time-domain solver. The ideas for the frequency selection can be applied to
+time selection, and time-dependent, adaptive gridding exists as well
+\citep[e.g., ][]{RS.06.Commer}. In the next section, we briefly review the
+methodology as introduced by \cite{GEO.07.Plessix} and \cite{GEO.08.Mulder} and
+highlight their advantages and shortcomings. This is followed by an outline of
+our changes to the Fourier transform and the adaptive gridding. Finally, we
+demonstrate the efficiency of the approach with some numerical results.
 
@@
 
-Being able to model CSEM data both in the frequency-domain and in the
-time-domain can be desirable, as both domains have advantages and
-disadvantages. One way to achieve this is to implement Maxwell's equation in
-both domains\deleted{, as it is done, for instance, in \simpeg (Cockett et al.,
-2015)}. Another approach is to have Maxwell's equation only implemented in one
-domain, and use Fourier transforms to go to the other. However, this approach
-can be costly, as many frequencies over a wide range are required to go from
-the time domain to the frequency domain, or many times over a wide range for
-the opposite direction. We present a methodology which significantly reduces
-the amount and range of the required frequencies, translating into a
-significant reduction in computation time. The required computation grids for
-low frequencies (in our case in the order of 0.001\,Hz) and high frequencies
+Being able to model CSEM data both in the frequency domain and in the time
+domain can be desirable, as both domains have advantages and disadvantages. One
+way to achieve this is to implement Maxwell's equations in both domains.
+Another approach is to have Maxwell's equations only implemented in one domain,
+and use Fourier transforms to go to the other. However, this approach can be
+costly, as many frequencies over a wide range are required to go from the
+frequency domain to the time domain, or many times over a wide range for the
+opposite direction. We present a methodology which significantly reduces the
+amount and range of the required frequencies, translating into a significant
+reduction in computation time. The required computation grids for low
+frequencies (in our case in the order of 0.001\,Hz) and high frequencies
 (around 100\,Hz) are hugely different. Low frequencies can be computed on a
@@
 conductivity and frequency, and for the diffusive approximation of Maxwell's
-equation in an isotropic, homogeneous medium is given by \citep[e.g.,][
+equations in an isotropic, homogeneous medium is given by \citep[e.g.,][
 equation~1.53]{B.SEG.88.Ward}
@@
   \delta = \sqrt{\frac{2}{\omega\mu\sigma}}
-  \ \approx \
-  % \frac{503.3}{\sqrt{f\sigma}} \, ,
-  % 503.3(f\sigma)^{-1/2} \, ,
-  503.3/\sqrt{f\sigma} \, ,
+  \ \approx \ 503.3/\sqrt{f\sigma} \, ,
   \label{eq:skindepth}
@@
 $z$-direction to account for the airwave. To reduce the number of cells, it is
-desirable to introduce stretching, at least in the boundary zone outside of the
+desirable to introduce stretching, at least in the buffer zone outside of the
 area where source and receivers are located.
@@
 \cite{GEO.08.Mulder} provide a computational complexity analysis of various
-methods to model transient electromagnetic responses directly in the
-time-domain, and compare it to the computation of transient EM responses in the
-frequency-domain with a subsequent Fourier transform. They conclude the review
+methods to model transient electromagnetic responses directly in the time
+domain, and compare it to the computation of transient EM responses in the
+frequency domain with a subsequent Fourier transform. They conclude the review
 by stating \emph{\guillemotleft Although it remains to be seen which of the
@@
 frequency domain to time domain, which makes time-domain modelling with a
-frequency-domain code even more competitive in comparison with time-domain
-codes. It is important to note that while we show the transform from frequency
-to time domain, the conclusions regarding the Fourier transform can equally
-well be applied to the transform from time to frequency domain data for a code
+frequency-domain code viable in comparison with time-domain codes. It is
+important to note that while we show the transform from frequency to time
+domain, the conclusions regarding the Fourier transform can equally well be
+applied to the transform from time-domain to frequency-domain data for a code
 that solves Maxwell's equations in the time domain. A variation of our method
@@
 solvers or as a solver on its own. The multigrid approach works fine for the
-diffusive approximation of Maxwell's equation, which assumes that
+diffusive approximation of Maxwell's equations, which assumes that
 $\omega\varepsilon \ll \sigma$, where $\varepsilon$ is electric permittivity
@@
 standard multigrid approach fails for severe stretching or strong anisotropy,
-for which \deleted{\emg3d has implemented }known improvements such as
-line-relaxation and semicoarsening \citep{ECCFD.06.Jonsthovel} \added{are
-implemented}, with a non-standard Cholesky decomposition to speed up the
-computation \citep{GEO.08.Mulder}. One of the big advantages of the multigrid
-method is that it scales linearly (optimal) with the grid size in both CPU and
-RAM usage \citep{B.Springer.20.Mulder}. This makes it feasible to run even big
-models on standard computers, without the need for big clusters. All the
-examples in this article are run on a laptop with an i7-6600U\,CPU\,@\,2.6\,GHz
-(x4) and 16\,GB of memory, using Ubuntu \replaced{20.04}{18.04} and Python
-\replaced{3.8}{3.7}. \added{All examples were run using a single thread, but
+for which known improvements such as line-relaxation and semicoarsening
+\citep{ECCFD.06.Jonsthovel} are implemented, with a non-standard Cholesky
+decomposition to speed up the computation \citep{GEO.08.Mulder}. One of the big
+advantages of the multigrid method is that it scales linearly (optimal) with
+the grid size in both CPU and RAM usage \citep{B.Springer.20.Mulder}. This
+makes it feasible to run even big models on standard computers, without the
+need for big clusters. All the examples in this article are run on a laptop
+with an i7-6600U\,CPU\,@\,2.6\,GHz (x4) and 16\,GB of memory, using Ubuntu
+20.04 and Python 3.8. All examples were run using a single thread, but
 parallelization over frequencies is straight forward, as the computations are
-independent.} For the semi-analytical computations of layered models we use the
+independent. For the semi-analytical computations of layered models we use the
 open-source (Apache License 2.0) code \empymod \citep{GEO.17.Werthmuller}.
 
-\section{Frequency selection} % % % % % % % % % % % % % % % % % % % % % % % % %
+\subsection{Frequency selection}  % % % % % % % % % % % % % % % % % % % % % % %
 
@@
 frequencies. The fewer frequencies required, the quicker we obtain the
-time-domain result. \added{The same applies for the reverse operation, the
-fewer times required, the quicker we obtain frequency-domain results. Even
-though we limit our analysis here to frequency-to-time transformation, most of
-the arguments are reversible and applicable to time-to-frequency
-transformations.} We decided to use a regular spacing of frequencies on a
-log-scale, rather than an adaptive scheme. This approach is favourable if a
-wide range of offsets is needed, as the required frequencies change with offset
-and an adaptive frequency selection is therefore often tailored to a single
-offset. Also, it allows for straightforward parallelization over frequencies,
-which is not completely possible with an adaptive scheme.
+time-domain result. The same applies for the reverse operation, the fewer times
+required, the quicker we obtain frequency-domain results. Even though we limit
+our analysis here to frequency-to-time transformations, most of the arguments
+are reversible and applicable to time-to-frequency transformations. We decided
+to use a regular spacing of frequencies on a log-scale, rather than an adaptive
+scheme. This approach is favourable if a wide range of offsets is needed, as
+the required frequencies change with offset and an adaptive frequency selection
+is therefore often tailored to a single offset. Also, it allows for
+straightforward parallelization over frequencies, which is not completely
+possible with an adaptive scheme.
 
@@
 in comparison with the standard FFT: the required input frequencies are equally
-spaced on a logarithmic scale\added{ (natural logarithm in the case of the DLF
-and decimal logarithm in the case of the FFTLog)}, and they only require either
-the real or the imaginary part of the frequency-domain response. We can take
+spaced on a logarithmic scale (natural logarithm in the case of the DLF and
+decimal logarithm in the case of the FFTLog), and they only require either the
+real or the imaginary part of the frequency-domain response. We can take
 advantage of that by using only the imaginary part of the frequency-domain
-response. The imaginary part goes to zero when the frequency goes
-\added{either} to zero or to infinity, with the advantage that knowing the
-endpoints makes it possible to convert the extrapolation of missing frequencies
-into \added{an }interpolation.
+response. The imaginary part goes to zero when the frequency goes either to
+zero or to infinity, with the advantage that knowing the endpoints makes it
+possible to convert the extrapolation of missing frequencies into an
+interpolation.
 
@@
 
-The \added{shown} example model is a marine scenario with 1\,km water depth of
+The shown example model is a marine scenario with 1\,km water depth of
 resistivity $\rho=0.3\,\ohmm$ ($\rho = \sigma^{-1}$) and a 100\,m thick target
@@
 imaginary part in the frequency domain and the right plot the corresponding
-impulse response in the time domain. The red lines are the
-\replaced{semi-analytical responses}{precise results obtained from \empymod}.
-The blue circles indicate the actually computed responses and the black dots
-the frequencies which are interpolated or set to zero. The resulting
+impulse response in the time domain. The red lines are the semi-analytical
+responses. The blue circles indicate the actually computed responses and the
+black dots the frequencies which are interpolated or set to zero. The resulting
 time-domain response has a relative error of less than 1\,\% everywhere except
@@
 and below 0.3\,s. It also shows the oscillating high-frequency part, which is
-hard to interpolate\added{ and we therefore set to zero}. Figure
-\ref{fig:GUI-DLF-lin} is the same as \ref{fig:GUI-FFTLog-lin}, but transformed
-with the DLF method applying the 81-point sine-cosine filter from
+hard to interpolate and we therefore set it to zero. Figure
+\ref{fig:GUI-DLF-lin} is the same as Figure \ref{fig:GUI-FFTLog-lin}, but
+transformed with the DLF method applying the 81-point sine-cosine filter from
 \cite{GEO.09.Key}. The same frequencies were computed as in the FFTLog case and
@@
   \caption{(a) Same as Figure \ref{fig:GUI-FFTLog-lin}, but on a logarithmic
-  scale. (b) Using DLF instead of FFTLog, but computing the same
-  frequenc\replaced{y range}{ies} as for the FFTLog. The resulting time domain
-  response has comparable accuracy (compare the error to the error in Figure
+  scale. (b) Using DLF instead of FFTLog, but computing the same frequency
+  range as for the FFTLog. The resulting time-domain response has comparable
+  accuracy (compare the error to the error in Figure
   \ref{fig:GUI-FFTLog-lin}).}
@@
 wide range of offsets. This is shown in Figure \ref{fig:multi-offset}, where
-the above parameters were used to model the responses at offsets $r=1.5,3,6$,
-and 12\,km. \added{We only need to compute 50\,\% of the required frequencies
-in the case of the FFTLog, and only 14\,\% in the case of the DLF.}
+the same kernels were used to yield the responses at offsets $r=1.5,3,6$, and
+12\,km. We only need to compute 50\,\% of the required frequencies in the case
+of the FFTLog, and only 14\,\% in the case of the DLF.
 %
@@
     $r=1.5,3,6$, and 12\,km using the parameters defined in Figure
-    \ref{fig:GUI-FFTLog-lin}. The coloured \replaced{symbols}{circles} are the
-    actually computed responses, the black dots are the responses which are set
-    to zero \added{(high frequencies)} or interpolated \added{(low
-    frequencies)}. The \replaced{grey}{black} curves are the analytical
-    responses.}
+    \ref{fig:GUI-FFTLog-lin}. The coloured symbols are the actually computed
+    responses, the black dots are the responses which are set to zero (high
+    frequencies) or interpolated (low frequencies). The grey curves are the
+    analytical responses.}
   \label{fig:multi-offset}
@@
 offset the lowest required frequency. So the more one can restrict the
-necessary offset range, the \replaced{fewer}{less} frequencies are needed.
-Another important factor is how to interpolate and extrapolate from the
-computed frequencies to the frequencies required for the Fourier transform. For
-the FFTLog only extrapolation for higher and lower frequencies is required. The
-EM response becomes highly oscillatory for high frequencies, which makes it
-very hard to extrapolate the response to frequencies $f>f_\mathrm{max}$.
-However, if $f_\mathrm{max}$ is chosen judiciously, the importance of higher
-frequencies for the Fourier transform can be neglected and we can set those
-responses to zero. The extrapolation of frequencies $f<f_\mathrm{min}$ can be
-changed to an interpolation by assuming a zero imaginary response at zero
-frequency, and we then use PCHIP to interpolate the missing frequencies (as we
-work on a logarithmic scale we cannot choose 0\,Hz \added{nor 0\,V/m}, but
-instead take $10^{-100}$\,Hz \added{and $10^{-100}$\,V/m}). In the case of the
-DLF method we also have to interpolate in between the computed frequencies, for
-which we found it better to use a cubic spline. As can be seen from Figures
-\ref{fig:GUI-FFTLog-lin} and \ref{fig:GUI-DLF-lin}, using the FFTLog or the DLF
-with the same actually computed frequencies results in very similar responses.
+necessary offset range, the fewer frequencies are needed. Another important
+factor is how to interpolate and extrapolate from the computed frequencies to
+the frequencies required for the Fourier transform. For the FFTLog only
+extrapolation for higher and lower frequencies is required. The EM response
+becomes highly oscillatory for high frequencies, which makes it very hard to
+extrapolate the response to frequencies $f>f_\mathrm{max}$. However, if
+$f_\mathrm{max}$ is chosen judiciously, the importance of higher frequencies
+for the Fourier transform can be neglected and we can set those responses to
+zero. The extrapolation of frequencies $f<f_\mathrm{min}$ can be changed to an
+interpolation by assuming a zero imaginary response at zero frequency, and we
+then use PCHIP to interpolate the missing frequencies (as we work on a
+logarithmic scale we cannot choose 0\,Hz nor 0\,V/m, but instead take
+$10^{-100}$\,Hz and $10^{-100}$\,V/m). In the case of the DLF method we also
+have to interpolate in between the computed frequencies, for which we found it
+better to use a cubic spline. As can be seen by comparing Figure
+\ref{fig:GUI-FFTLog-lin} with Figure \ref{fig:GUI-DLF-lin}, using the FFTLog or
+the DLF with the same actually computed frequencies results in very similar
+responses.
 
@@
 takes much longer than the transform itself. What matters is solely how many
-frequencies are required by it in order to be stable, and how long it takes to
-compute the responses for these frequencies.
-
-\section{Gridding}  % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
-
-The computation grid consists of a core or survey domain $D_\mr{s}$ which
-should contain all source and receiver positions. The survey domain usually has
-no or very little cell stretching $\alpha_\mr{s}$. The minimum cell width is
+frequencies are required by it to achieve the desired precision, and how long
+it takes to compute the responses for these frequencies.
+
+\subsection{Gridding}   % % % % % % % % % % % % % % % % % % % % % % % % % % % %
+
+The computation grid consists of a core or survey domain $D_\mr{s}$ that should
+contain all source and receiver positions. The survey domain usually has no or
+very little cell stretching factor $\alpha_\mr{s}$. The minimum cell width is
 defined as
@@
 resides, and $n_\delta$ is a positive number that defines how many cells there
-should be per skin depth. The actual computation domain $D_\mr{c}$ is usually
+should be per skin depth. The actual computational domain $D_\mr{c}$ is usually
 much bigger than $D_\mr{s}$ in order to avoid artefacts from the perfectly
 electrically conducting boundary condition. It can also have a much higher
-stretching $\alpha_\mr{c}$. In \replaced{the presented examples}{our scheme} we
-have chosen $D_\mr{c}$ such that the distance for the signal diffusing from the
-source to the boundary and back to the receiver closest to the boundary is at
-least two wavelengths, after which the initial signal is reduced to a millionth
-of its strength. The wavelength \added{$\lambda$} to compute $D_\mr{c}$ is
+stretching factor $\alpha_\mr{c}$. In the presented examples we have chosen
+$D_\mr{c}$ such that the distance for the signal diffusing from the source to
+the boundary and back to the receiver closest to the boundary is at least two
+wavelengths, after which the amplitude of the signal is reduced to a millionth
+of its initial strength. The wavelength $\lambda$ to compute $D_\mr{c}$ is
 given by
@@
 \begin{equation}
-  \deleted{2}\lambda = \replaced{2}{4}\pi\delta(f, \sigma_\mr{ave})
-  \ \approx \ \replaced{3162}{6324.7}/\sqrt{f\sigma_\mr{ave}}\, ,
+  \lambda = 2\pi\delta(f, \sigma_\mr{ave})
+  \ \approx \ 3162/\sqrt{f\sigma_\mr{ave}}\, ,
  \label{eq:lambda}
@@
 computational domain is therefore enforced, defining the maximum distance from
-the source to the boundary; this distance is by default 100\,km, but this can
-be reduced in the marine case with increasing water depth. Note that for
-shallow marine and land cases this also applies to the horizontal dimensions,
-not only to the upward $z$-direction and similarly for very resistive
-basements, even in deep water. One way to circumvent this difficulty is the use
-of a primary-secondary formulation, where the primary field, including the air
-wave, is computed with a semi-analytical code for layered media. We do not
-consider this approach here.
+the source to the boundary; this distance is by default set to 100\,km, but
+this can be reduced in the marine case with increasing water depth. Note that
+for shallow marine and land cases this also applies to the horizontal
+dimensions, not only to the upward $z$-direction and similarly for very
+resistive basements, even in deep water. One way to circumvent this difficulty
+is the use of a primary-secondary formulation, where the primary field,
+including the air wave, is computed with a semi-analytical code for layered
+media. We do not consider this approach here.
 
@@
 $\Delta_\mr{min}$ can further be restricted by a user-defined range. Given
-these inputs the adaptive gridding will search for the smallest possible number
-of cells which fulfils these criteria. The \added{implemented} multigrid method
-\deleted{implemented in \emg3d} puts some constraints on the number of cells, of
-which the adaptive gridding takes care (the number of cells have to be powers
-of two multiplied by a low prime, e.g., $\{2,3,5\}\cdot2^n$).
+these inputs, the adaptive gridding will search for the smallest possible
+number of cells which fulfils these criteria. The implemented multigrid method
+puts some constraints on the number of cells, of which the adaptive gridding
+takes care (the number of cells have to be powers of two multiplied by a low
+prime, e.g., $\{2,3,5\}\cdot2^n$).
 
@@
 criteria, for a fixed number of cells. Our adaptive gridding, on the other
-hand, searches for the smallest number of cells that still fulfil the given
+hand, searches for the smallest number of cells that still fulfils the given
 criteria. The number of cells becomes therefore also a function of frequency.
-\added{It is important to note that this is our implementation of an adaptive
-grid, but there are certainly other possibilities. The relevant point for fast
+It is important to note that this is our implementation of an adaptive grid,
+but there are certainly other possibilities. The relevant point for fast
 computations is that the adaptive gridding tries to minimize the number of
-required cells. This is generally best done in a frequency and conductivity
-dependent manner.} To go from the model grid to the computational grid, we use
-the volume-averaging technique on logarithmic resistivities, as used in
-\cite{GEO.07.Plessix}. While this technique ensures that the total resistivity
-in the subsurface remains the same, it does not consider effective-medium
-theory \citep{GEO.03.Davydycheva}, for instance, the apparent anisotropy from a
-stack of finely layered formations of varying resistivity.
+required cells. This is generally best done in a frequency- and
+conductivity-dependent manner. To go from the model grid to the computational
+grid, we use the volume-averaging technique on logarithmic resistivities, as
+used in \cite{GEO.07.Plessix}. While this technique ensures that the total
+resistivity in the subsurface remains the same, it does not consider
+effective-medium theory \citep{GEO.03.Davydycheva}, for instance, the apparent
+anisotropy from a stack of finely layered formations of varying resistivity.
 
@@
 compute from 0.05\,Hz to 20.0\,Hz. The complete frequency range for the
-transform includes 30 frequencies from 0.0002\,Hz to 126.4\,Hz. For the
-adaptive gridding the following inputs were used: $n_\delta=12$, minimum cell
-width must be between 20 and 40\,m, and $\alpha_\mr{s}=1$,
-$\alpha_\mr{c}=[1,1.3]$. This created grids with cell numbers between
-\num{46080} ($80\times24\times24$, for 20.0\,Hz) and \num{128000}
-($80\times40\times40$, for 0.05\,Hz) cells. The run times for each frequency,
-the corresponding number of cells, minimum cell width and computation domain
-stretching factor are listed in Table~\ref{tbl:timefull}. The total run time to
-compute this model was less than two minutes.
+transform, including the frequencies for which we use interpolation, includes
+30 frequencies from 0.0002\,Hz to 126.4\,Hz. For the adaptive gridding the
+following inputs were used: $n_\delta=12$, minimum cell width must be between
+20 and 40\,m, and $\alpha_\mr{s}=1$, $\alpha_\mr{c}=[1,1.3]$. This yields grids
+with cell numbers between \num{46080} ($80\times24\times24$, for 20.0\,Hz) and
+\num{128000} ($80\times40\times40$, for 0.05\,Hz) cells. The run times for each
+frequency, the corresponding number of cells, minimum cell width, and
+computation domain stretching factor are listed in Table~\ref{tbl:timefull}.
+The total run time to compute this model was less than two minutes.
 %
@@
     model. The red lines are the analytical solutions, the blue circles are the
-    actually computed responses\deleted{ with \emg3d}, the black dots are the
-    interpolated responses, and the dashed black line the obtained time-domain
-    response. The errors are clipped for values smaller than 0.1\,\% and bigger
-    than 100\,\%.}
+    actually computed responses, the black dots are the interpolated responses,
+    and the dashed black line the obtained time-domain response. The errors are
+    clipped for values smaller than 0.1\,\% and bigger than 100\,\%.}
   \label{fig:fullspace}
@@
 frequencies. Figure \ref{fig:fullspace}(b) shows the corresponding time-domain
-result, where the dashed black line is the result \replaced{obtained by
-transformation of the frequency-domain response}{from \emg3d}, on
-top of the red line which is the analytical result. The relative error is
-mostly below 1\,\%, except for early times. However, for practical reasons that
-is more than enough. Figure~\ref{fig:fullspace-log} shows the same on a
-logarithmic scale, with times up to 10 seconds. It clearly shows that if later
-times are required, we would need to adjust our Fourier transform parameters.
+result, where the dashed black line is the result obtained by transformation of
+the frequency-domain response, on top of the red line which is the analytical
+result. The relative error is mostly below 1\,\%, except for early times.
+However, for practical reasons that is more than enough.
+Figure~\ref{fig:fullspace-log} shows the same on a logarithmic scale, with
+times up to 10 seconds. It clearly shows that if later times are required, we
+would need to adjust our Fourier transform parameters.
 %
@@
 took less than two minutes in total, where the individual frequencies took
-between \replaced{4}{3} and \replaced{11}{15}\,s to run.
+between 4 and 11\,s to run.
 
@@
   \caption{Electric inline response at an offset of 5\,km for a shallow marine,
-    layered scenario. (a) Frequency-domain, where the blue circles are computed
-    responses and the black dots are interpolated responses or responses set to
-    zero. (b) Time-domain response.}
+    layered scenario. (a) Frequency-domain response, where the blue circles
+    denote computed responses and the black dots interpolated responses or
+    responses set to zero. (b) Time-domain response.}
   \label{fig:marine}
@@
 frequency-domain computations and the same frequencies for the Fourier
-transform. The computation of these frequencies took less than 13 minutes and
-it handles any offset between 3 and 7\,km. It can be seen that the chosen
+transform. The computation of these frequencies took less than 9 minutes and it
+handles any offset between 3 and 7\,km. It can be seen that the chosen
 frequency selection is sufficient for this offset range; again, more low
@@
 spreading. On land and in shallow marine scenarios one has therefore to include
-a sufficiently large computational domain, and the default in our scheme is
+a sufficiently large computational domain. The default in our scheme is
 100\,km. The important point is that this does not only apply to the upward
-$z$-direction, but also to the horizontal directions, as the airwave also
-bounces back horizontally and would continuously emit energy into the
-subsurface if the boundaries are not chosen far enough away from the receivers.
-If models are computed with very resistive layers or models with highly
-resistive basements, this can even apply to deep marine scenarios.
+$z$-direction, but also to the horizontal directions, as the airwave bounces
+back horizontally and would continuously emit energy into the subsurface if the
+boundaries are not chosen far enough away from the receivers. If models are
+computed with very resistive layers or models with highly resistive basements,
+this can even apply to deep marine scenarios.
 
@@
 the previous section; however, for the adaptive gridding in the horizontal
-directions $\rho_\mr{ave}=1\,\ohmm$ was used instead of
+directions, $\rho_\mr{ave}=1\,\ohmm$ was used instead of
 $\rho_\mr{ave}=\num{10000}\,\ohmm$. Having the boundaries too near in the
@@
 It can be difficult to spot these errors in the time-domain result, as the
-response looks plausible and only comparing it with the 1D result reveals that
-it is actually wrong. A possibility to detect such problems for complicated
-cases, where there is no semi-analytical result to compare with, is to compute
-two or more models, moving the boundary. When the responses stop to change, one
-can assume that the boundary is far enough. Another possibility is to look at
-the amplitudes close to the boundaries and ensure that they are small enough.
+response looks plausible. Only a comparison with the 1D result reveals that it
+is actually wrong. A possibility to detect such problems for complicated cases,
+where there is no semi-analytical result to compare with, is to compute two or
+more models, moving the boundary. When the responses stop to change, one can
+assume that the boundary is far enough. Another possibility is to look at the
+amplitudes close to the boundaries and ensure that they are small enough.
 
@@
 
-For the comparison we use the open-source code \simpeg
-\citep{CAG.15.Cockett}\replaced{, which}{. \simpeg} is a framework for modelling
-and inversion of geophysical data such as gravity, magnetics, and CSEM. It has
-Maxwell's equations implemented in both the frequency and time domain. As such
-we can compare our result computed in the frequency domain followed by a
-Fourier transform to a result computed directly in the time domain. A principal
-difference between \simpeg and \emg3d is that the former has various direct
-solvers implemented, whereas the latter is an iterative multigrid solver. The
-3D model is therefore a rather small example in order to be able to run it on
-our test machine, as the memory requirement by the direct solver would
-otherwise be too high. There are not many options out there of open-source
-time-domain 3D codes, \simpeg being the one we found to be suitable. A step-off
-response was chosen as this is the response currently implemented in
-\replaced{it}{\simpeg}.
+For the comparison we use the open-source code \simpeg \citep{CAG.15.Cockett},
+which is a framework for modelling and inversion of geophysical data such as
+gravity, magnetics, and CSEM. It has Maxwell's equations implemented in both
+the frequency and time domain. As such we can compare our result computed in
+the frequency domain followed by a Fourier transform to a result computed
+directly in the time domain. A principal difference between \simpeg and \emg3d
+is that the former has various direct solvers implemented, whereas the latter
+is an iterative multigrid solver. The 3D model is therefore a rather small
+example in order to be able to run it on our test machine, as the memory
+requirement by the direct solver would otherwise be too high. There are not
+many options out there of open-source time-domain 3D CSEM codes, \simpeg being
+the one we found to be suitable. A step-off response was chosen as this is the
+response currently implemented in it.
 
 The model was discretised with $100\times100\times100\,$m cells in the survey
-domain $D_\mr{s}$. For the \replaced{time-domain}{\simpeg} model, 14 cells in
-$x$-direction and 12 cells in $y$- and $z$-directions were used on both sides
-with a stretching of 1.3 for the total computation domain $D_\mr{c}$, which
-yields a mesh of \num{58344} cells. The time-steps start at 0.1\,s and are:
-$21\times0.01\,$s, $23\times0.03\,$s, $21\times0.1\,$s, $23\times0.3\,$s,
-covering exactly the desired range of 0.1--10\,s. For
-\replaced{the frequency-domain model}{\emg3d}, the mesh is generated
-frequency-dependent as in the previous examples, with a maximum stretching of
-$\alpha_\mr{c}=1.5$. This results in meshes between \num{18432} cells for the
-highest frequencies and \num{76800} cells for the lowest frequencies. The
-required frequencies were obtained by using the FFTLog with five points per
-decade, which results in 20 frequencies between 0.001\,Hz and 8\,Hz. The actual
-transform was carried out with the 201-point sine-cosine filter from
-\cite{GEO.09.Key}.
+domain $D_\mr{s}$. For the time-domain model, 14 cells in $x$-direction and 12
+cells in $y$- and $z$-directions were used on both sides with a stretching of
+1.3 for the total computation domain $D_\mr{c}$, which yields a mesh of
+\num{58344} cells. The time-steps start at 0.1\,s and are: $21\times0.01\,$s,
+$23\times0.03\,$s, $21\times0.1\,$s, $23\times0.3\,$s, covering exactly the
+desired range of 0.1--10\,s. For the frequency-domain model, the mesh is
+generated frequency-dependent as in the previous examples, with a maximum
+stretching of $\alpha_\mr{c}=1.5$. This results in meshes between \num{18432}
+cells for the highest frequencies and \num{76800} cells for the lowest
+frequencies. The required frequencies were obtained by using the FFTLog with
+five points per decade, which results in 20 frequencies between 0.001\,Hz and
+8\,Hz. The actual transform was carried out with the 201-point sine-cosine
+filter from \cite{GEO.09.Key}.
 
 The results are shown in Figure~\ref{fig:3d-result}: In (a) the 1D background
-responses and the relative error using the semi-analytical
-result\deleted{obtained from \empymod}, and in (b) the responses including the
-target.
+responses and the relative error using the semi-analytical result, and in (b)
+the responses including the target.
 %
@@
   \caption{Responses for the model outlined in Figure~\ref{fig:3d-model} using
-    \replaced{time-domain}{\simpeg} and \replaced{frequency-domain
-    computations}{\emg3d}, and for the layered background also \replaced{the
-    semi-analytical result}{\empymod}. In the lower plot of (a) the relative
-    error (\%) \added{is shown} in comparison to \replaced{the semi-analytical
-    result}{\empymod is shown}, and in (b) the normalized difference (\%)
-    between the two 3D codes. The results of these plots are clipped for values
-    smaller than 0.1\,\% and bigger than 10\,\%.}
+    time-domain and frequency-domain computations, and for the layered
+    background also the semi-analytical result. In the lower plot of (a) the
+    relative error (\%) is shown in comparison to the semi-analytical result,
+    and in (b) the normalized difference (\%) between the two 3D codes. The
+    results of these plots are clipped for values smaller than 0.1\,\% and
+    bigger than 10\,\%.}
   \label{fig:3d-result}
@@
 %
-The background comparison shows that both 3D codes do an acceptable job having
-a relative error of a few percents at most; the result \replaced{obtained
-through transformation}{for \emg3d} seems to be better at early times.
-\added{The reason is probably the implemented backward Euler in the time-domain
-code which has an error of order one in time.} We cannot compare the errors for
-the response that includes the target for lack of an analytical solution. The
-\added{1D} background model \deleted{from \empymod }is only included to show
-that there is a significant response from the target. We therefore show the
-normalized difference (\replaced{NRMSD}{ND}) between the two
-\replaced{responses $R_1$ and $R_2$}{3D codes} as a percentage, where
-\replaced{$\mr{NRMSD~(\%)} = 200|R_1 - R_2|/(|R_1| + |R_2|)$.}{$\mr{ND}=
-50|(\simpeg-\emg3d)/(\simpeg+\emg3d)|$.} The \replaced{NRMSD}{ND} between the
-two codes is below 1\,\% everywhere except for early times. Both codes took
-\replaced{roughly 4--5}{a little over 7} minutes to compute the two models
+The background comparison shows that both 3D codes do an acceptable job with a
+relative error of a few percents at most; the result obtained through
+transformation seems to be better at early times. The reason is probably the
+implemented backward Euler scheme in the time-domain code that has an error of
+order one in time. We cannot compare the errors for the response that includes
+the target for lack of an analytical solution. The 1D background model is only
+included to show that there is a significant response from the target. We
+therefore show the normalized difference (NRMSD) between the two responses
+$R_1$ and $R_2$ as a percentage, where $\mr{NRMSD~(\%)} = 200|R_1 - R_2|/(|R_1|
++ |R_2|)$. The NRMSD between the two codes is below 1\,\% everywhere except for
+early times. Both codes took roughly 4--5 minutes to compute the two models
 (single thread). However, in this particular comparison, the main difference in
@@
 iterative solver vs. direct solver.
-
-\deleted{\emph{(Deleted last three sentences, as they do not add a lot.)}For a
-test we doubled the number of cells in each direction by using 50\,m for the
-minimum cell width instead of 100\,m, and taking also half of the stretching,
-which results then in a mesh of \num{466752} cells for \simpeg and meshes
-between \num{102400} and \num{614400} for \emg3d. Because the memory of our
-test machine was not large enough for the direct solver of \simpeg, the last
-model had to be run on a cluster. In this scenario, \emg3d was about eight
-times faster (single thread).}
 
@@
 Twenty frequencies or less are usually sufficient for a wide range of offsets.
-\added{The values for lower frequencies can be interpolated using PCHIP knowing
-that the imaginary part goes to zero for zero frequency. The values for higher
+The values for lower frequencies can be interpolated using PCHIP knowing that
+the imaginary part goes to zero for zero frequency. The values for higher
 frequencies can be set to zero, as we can neglect their influence. And values
 for frequencies in-between the computed ones are best obtained with a spline
-interpolation. The actual transform can be made with either the DLF method or
-FFTLog, where the latter one requires usually much fewer frequencies to be
-interpolated.} \deleted{We believe that our proposed improvements to the
-previously published methods make time-domain CSEM modelling with a
-frequency-domain code even more attractive than it was already before.} We have
-demonstrated the idea of our Fourier transform method on CSEM data transformed
-from the frequency to the time domain. However, it could equally be applied to
-the transform from the time to the frequency domain and to other methods with
-similar characteristics. \added{We believe that our proposed improvements to
-the previously published methods makes simulating results in one domain
-obtained through computations in the other domain followed by a transformation
-a viable alternative. The methodology is relatively simple to implement in any
-code.}
+interpolation. The actual transform can be carried out with either the DLF
+method or FFTLog, where the latter one requires usually much fewer frequencies
+to be interpolated. We have demonstrated the idea of our Fourier transform
+method on CSEM data transformed from the frequency domain to the time domain.
+However, it could equally be applied to the transform from the time domain to
+the frequency domain and to other methods with similar characteristics. We
+believe that our proposed improvements to the previously published methods
+makes simulating results in one domain obtained through computations in the
+other domain followed by a transformation a viable alternative. The methodology
+is relatively simple to implement in any code.
 
@@
 MarTERA as part of Horizon 2020, a funding scheme of the European Research Area
-(ERA-NET Cofund, \href{https://www.martera.eu}{martera.eu}).
-An early idea of this manuscript was presented at \cite{EAGE.20.Werthmuller}.
+(ERA-NET Cofund, \href{https://www.martera.eu}{martera.eu}). An early idea of
+this manuscript was presented by \cite{EAGE.20.Werthmuller}.
 \end{acknowledgments}

### 2021-02-22 ManuscriptTime.tex
Initiate Revision II
Point 8

@@
 \documentclass[extra, camera,%
-    % onecolumn,   % comment for two columns
-    % referee,     % uncomment for double lines
-    final,       % uncomment for changes
+    onecolumn,   % comment for two columns
+    referee,     % uncomment for double lines
+    % final,       % uncomment for changes
 ]{gji}
@@
 \newcommand{\ohmm}{\ensuremath{\Omega\,}\text{m}\xspace}
+\newcommand{\rtodo}{{\color{red}\texttt{TODO:}}\xspace}
 
@@
 
+\rtodo \added{Add citation to R.R. (2021) regarding RA and speed-up; limitation
+to switch-on/-off.}
+
+\rtodo \added{Add notion about preference for f-domain for IP problems}
+
 \section{Motivation}  % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
@@
 frequencies. The fewer frequencies required, the quicker we obtain the
-time-domain result. The same applies for the reverse operation, the fewer times
-required, the quicker we obtain frequency-domain results. Even though we limit
-our analysis here to frequency-to-time transformations, most of the arguments
-are reversible and applicable to time-to-frequency transformations. We decided
-to use a regular spacing of frequencies on a log-scale, rather than an adaptive
-scheme. This approach is favourable if a wide range of offsets is needed, as
-the required frequencies change with offset and an adaptive frequency selection
-is therefore often tailored to a single offset. Also, it allows for
-straightforward parallelization over frequencies, which is not completely
-possible with an adaptive scheme.
+time-domain result. \added{This applies primarily for expensive kernels such
+as used in 3D modelling where the computation of the response of a single
+frequency takes much longer than the transformation itself.} The same applies
+for the reverse operation, the fewer times required, the quicker we obtain
+frequency-domain results. Even though we limit our analysis here to
+frequency-to-time transformations, most of the arguments are reversible and
+applicable to time-to-frequency transformations. We decided to use a regular
+spacing of frequencies on a log-scale, rather than an adaptive scheme. This
+approach is favourable if a wide range of offsets is needed, as the required
+frequencies change with offset and an adaptive frequency selection is therefore
+often tailored to a single offset\added{ or a limited range of offset}. Also,
+it allows for straightforward parallelization over frequencies, which is not
+completely possible with an adaptive scheme.
 
@@
 
-There are three key-parameters that have to be defined: Minimum and maximum
-frequency ($f_\mr{min}, f_\mr{max}$) and the number of frequencies per decade
-($n_\mr{dec}^f$). Trial-and-error with a 3D code is very time-intensive.
-However, a simplified, layered model for the required survey setup and a fast
-1D modeller makes it possible to estimate these parameters easily. To this end,
-we created a graphical, interactive tool using \empymod, as shown in
-Figure~\ref{fig:GUI-FFTLog-lin}.
+\rtodo \added{Start this section with the method; fmin, fmax, ndec; below we
+interpolate with PCHIP, above we set to zero, in-between we interpolate with
+cubic spline.}
+\replaced{The actually computed frequencies in our approach are defined by
+three key-parameters}{There are three key-parameters that have to be defined}:
+\replaced{The m}{M}inimum and maximum frequency ($f_\mr{min}, f_\mr{max}$) and
+the number of frequencies per decade ($n_\mr{dec}^f$). \added{This creates a
+frequency range which is much smaller than the frequency range required for the
+transform.} \replaced{A t}{T}rial-and-error \added{approach} with a 3D code is
+very time-intensive. However, a simplified, layered model for the required
+survey setup and a fast 1D modeller makes it possible to estimate these
+parameters easily. \replaced{An example of this is shown in
+Figure~\ref{fig:FFTLog-log}.}{To this end, we created a graphical, interactive
+tool using \empymod, as shown in Figure~\ref{fig:FFTLog-log}.}
 %
@@
   \centering
-  \includegraphics[width=\cwidth]{01-GUI-FFTLog-lin}
-  \caption{Interactive frequency selection for a user-provided layered model
-    (the shown model parameters are described in the text). This example shows
-    the impulse response at an offset of 5\,km, for which it uses FFTLog from
-    0.001\,Hz to 10\,Hz with five frequencies per decade. Note that the error
-    is clipped for values smaller than 0.01\,\% and bigger than 100\,\%.}
-  \label{fig:GUI-FFTLog-lin}
+  \includegraphics[width=\cwidth]{01-FFTLog-log}
+  \caption{\replaced{Example}{Interactive} frequency selection for a
+    \replaced{simple}{user-provided} layered model (the shown model parameters
+    are described in the text). This example shows the impulse response at an
+    offset of 5\,km, for which it uses FFTLog from 0.001\,Hz to 10\,Hz with
+    five frequencies per decade. \deleted{Note that the error is clipped for
+    values smaller than 0.01\,\% and bigger than 100\,\%.}}
+  \label{fig:FFTLog-log}
 \end{figure}
 %
-
-The shown example model is a marine scenario with 1\,km water depth of
-resistivity $\rho=0.3\,\ohmm$ ($\rho = \sigma^{-1}$) and a 100\,m thick target
-of 100\,\ohmm at 1\,km below the seafloor in a background of 1\,\ohmm. The
-source is 50\,m above the seafloor, the receiver is on the seafloor, and the
-response is the inline $x$-directed $E$-field. The left subplot shows the
-imaginary part in the frequency domain and the right plot the corresponding
-impulse response in the time domain. The red lines are the semi-analytical
-responses. The blue circles indicate the actually computed responses and the
-black dots the frequencies which are interpolated or set to zero. The resulting
-time-domain response has a relative error of less than 1\,\% everywhere except
-for very early times.
-
-Figure \ref{fig:GUI-FFTLog-log} shows exactly the same on a logarithmic scale,
-without the interactive GUI-related elements. It can be seen that with the
-chosen frequencies the time-domain response starts to divert above about 100\,s
-and below 0.3\,s. It also shows the oscillating high-frequency part, which is
-hard to interpolate and we therefore set it to zero. Figure
-\ref{fig:GUI-DLF-lin} is the same as Figure \ref{fig:GUI-FFTLog-lin}, but
+\replaced{It }{The shown example model} is a marine scenario with 1\,km water
+depth of resistivity $\rho=0.3\,\ohmm$ ($\rho = \sigma^{-1}$) and a 100\,m
+thick target of 100\,\ohmm at 1\,km below the seafloor in a background of
+1\,\ohmm. The source is 50\,m above the seafloor, the receiver is on the
+seafloor, and the response is the inline $x$-directed $E$-field. The left
+subplot shows the imaginary part in the frequency domain and the right plot the
+corresponding impulse response in the time domain. The red lines are the
+semi-analytical responses. The blue circles indicate the actually computed
+responses and the black dots the frequencies which are interpolated or set to
+zero. The resulting time-domain response has a relative error of less than
+1\,\% everywhere except for very early times.
+
+Figure \ref{fig:FFTLog-lin} shows exactly the same on a
+\replaced{linear}{logarithmic} scale\replaced{for the time and amplitude
+axes}{, without the interactive GUI-related elements}. It can be seen that with
+the chosen frequencies the time-domain response starts to divert above about
+100\,s and below 0.3\,s. It also shows the oscillating high-frequency part,
+which is hard to interpolate and we therefore set it to zero.
+Figure~\ref{fig:DLF-lin} is the same as Figure \ref{fig:FFTLog-lin}, but
 transformed with the DLF method applying the 81-point sine-cosine filter from
@@
   \centering
-  \subfloat[]{{\includegraphics[width=0.5\fwidth]{02a-GUI-FFTLog-log}}%
-  \label{fig:GUI-FFTLog-log}}\hfill
-  \subfloat[]{{\includegraphics[width=0.5\fwidth]{02b-GUI-DLF-lin}}%
-  \label{fig:GUI-DLF-lin}}\\
-  \caption{(a) Same as Figure \ref{fig:GUI-FFTLog-lin}, but on a logarithmic
+  \subfloat[]{{\includegraphics[width=0.5\fwidth]{02a-FFTLog-lin}}%
+  \label{fig:FFTLog-lin}}\hfill
+  \subfloat[]{{\includegraphics[width=0.5\fwidth]{02b-DLF-lin}}%
+  \label{fig:DLF-lin}}\\
+  \caption{(a) Same as Figure \ref{fig:FFTLog-log}, but on a logarithmic
   scale. (b) Using DLF instead of FFTLog, but computing the same frequency
   range as for the FFTLog. The resulting time-domain response has comparable
-  accuracy (compare the error to the error in Figure
-  \ref{fig:GUI-FFTLog-lin}).}
-  \label{fig:GUI-log-lin}
+  accuracy\deleted{ (compare the error to the error in Figure
+    \ref{fig:FFTLog-log})}.}
+  \label{fig:lin-lin}
 \end{figure*}
@@
     $r=1.5,3,6$, and 12\,km using the parameters defined in Figure
-    \ref{fig:GUI-FFTLog-lin}. The coloured symbols are the actually computed
+    \ref{fig:FFTLog-lin}. The coloured symbols are the actually computed
     responses, the black dots are the responses which are set to zero (high
@@
 better to use a cubic spline. As can be seen by comparing Figure
-\ref{fig:GUI-FFTLog-lin} with Figure \ref{fig:GUI-DLF-lin}, using the FFTLog or
+\ref{fig:FFTLog-lin} with Figure \ref{fig:DLF-lin}, using the FFTLog or
 the DLF with the same actually computed frequencies results in very similar
@@
 \subsection{Gridding}   % % % % % % % % % % % % % % % % % % % % % % % % % % % %
+
+\rtodo \added{Mention again that important for the transform are accurate
+results; these gridding as what we implemented in the 3D code.}
 
@@
     actually computed responses, the black dots are the interpolated responses,
-    and the dashed black line the obtained time-domain response. The errors are
-    clipped for values smaller than 0.1\,\% and bigger than 100\,\%.}
+    and the dashed black line the obtained time-domain response. \deleted{The
+    errors are clipped for values smaller than 0.1\,\% and bigger than
+    100\,\%.}}
   \label{fig:fullspace}
@@
     relative error (\%) is shown in comparison to the semi-analytical result,
-    and in (b) the normalized difference (\%) between the two 3D codes. The
-    results of these plots are clipped for values smaller than 0.1\,\% and
-    bigger than 10\,\%.}
+    and in (b) the normalized difference (\%) between the two 3D codes.
+    \deleted{The results of these plots are clipped for values smaller than
+    0.1\,\% and bigger than 10\,\%.}}
   \label{fig:3d-result}
@@
 iterative solver vs. direct solver.
+
+\subsection{\added{Induced Polarization}} % % % % % % % % % % % % % % % % % % %
 

### 2021-02-24 ManuscriptTime.tex
EOD

@@
 
+\rtodo \added{Extra example: No-IP vs IP; land, 1D}
+
 \section{Motivation}  % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
@@
 opposite direction. We present a methodology which significantly reduces the
-amount and range of the required frequencies, translating into a significant
-reduction in computation time. The required computation grids for low
-frequencies (in our case in the order of 0.001\,Hz) and high frequencies
+amount and range of the required frequencies \added{or times}, translating into
+a significant reduction in computation time. The required computation grids for
+low frequencies (in our case in the order of 0.001\,Hz) and high frequencies
 (around 100\,Hz) are hugely different. Low frequencies can be computed on a
@@
 are much more limited in reach. An adaptive gridding scheme is therefore
-required. We build our approach upon \cite{GEO.07.Plessix}, who presented such
-an adaptive gridding for multi-frequency (and multi-source) CSEM modelling.
-
-An adaptive grid in the frequency domain is naturally based on the skin depth,
-which is the distance after which the amplitude of the electromagnetic field is
-decayed by $1/e\approx 37\,\%$. The skin depth $\delta$ is a function of
-conductivity and frequency, and for the diffusive approximation of Maxwell's
-equations in an isotropic, homogeneous medium is given by \citep[e.g.,][
-equation~1.53]{B.SEG.88.Ward}
+required\replaced{, which}{. We build our approach upon
+\mbox{\cite{GEO.07.Plessix}}, who presented such an adaptive gridding for
+multi-frequency (and multi-source) CSEM modelling.
+
+An adaptive grid in the frequency domain} is naturally based on the skin depth,
+\deleted{which is} the distance after which the amplitude of the
+electromagnetic field is decayed by $1/e\approx 37\,\%$. The skin depth
+$\delta$ is a function of conductivity and frequency, and for the diffusive
+approximation of Maxwell's equations in an isotropic, homogeneous medium is
+given by \citep[e.g.,][ equation~1.53]{B.SEG.88.Ward}
 %
@@
 
-\cite{GEO.07.Plessix} define the minimum cell width $\Delta_\mathrm{min}$ as a
-fraction of the minimum skin depth $\delta_\mathrm{min}$, where
+\added{We build our approach upon} \cite{GEO.07.Plessix}\added{, who presented
+such an adaptive gridding for multi-frequency (and multi-source) CSEM
+modelling. They} define the minimum cell width $\Delta_\mathrm{min}$ as
+a fraction of the minimum skin depth $\delta_\mathrm{min}$, where
 $\delta_\mathrm{min}$ should be 2--3 times bigger than $\Delta_\mathrm{min}$.
@@
 applied to the transform from time-domain to frequency-domain data for a code
-that solves Maxwell's equations in the time domain. A variation of our method
-was implemented, for instance, by \custem \citep{GEO.19.Rochlitz} to obtain
-frequency-domain data from time-domain computations (Rochlitz, private
-communication, 2020).
+that solves Maxwell's equations in the time domain. \deleted{A variation of our
+method was implemented, for instance, by \custem \mbox{\citep{GEO.19.Rochlitz}}
+to obtain frequency-domain data from time-domain computations (Rochlitz,
+private communication, 2020).}
 
@@
 
-For the numerical computations we use the open-source (Apache License 2.0) code
-\emg3d \citep{JOSS.19.Werthmuller}, a multigrid solver based on
-\cite{GP.06.Mulder}, which can be used as a preconditioner for Krylov subspace
-solvers or as a solver on its own. The multigrid approach works fine for the
-diffusive approximation of Maxwell's equations, which assumes that
+\added{The requirement for any transform is a robust solver of sufficient
+precision.} \replaced{We use f}{F}or the numerical computations \deleted{we
+use} the open-source (Apache License 2.0) code\added{s \empymod
+\mbox{\citep{GEO.17.Werthmuller}} and} \emg3d
+\citep{JOSS.19.Werthmuller}\replaced{.}{,} \added{The former computes
+semi-analytical responses of layered models, the latter is} a multigrid solver
+based on \cite{GP.06.Mulder}, which can be used as a preconditioner for Krylov
+subspace solvers or as a solver on its own. The multigrid approach works fine
+for the diffusive approximation of Maxwell's equations, which assumes that
 $\omega\varepsilon \ll \sigma$, where $\varepsilon$ is electric permittivity
@@
 parallelization over frequencies is straight forward, as the computations are
-independent. For the semi-analytical computations of layered models we use the
-open-source (Apache License 2.0) code \empymod \citep{GEO.17.Werthmuller}.
+independent. \deleted{For the semi-analytical computations of layered models we
+use the open-source (Apache License 2.0) code \empymod
+\mbox{\citep{GEO.17.Werthmuller}}.}
 
@@
 
-\rtodo \added{Mention again that important for the transform are accurate
-results; these gridding as what we implemented in the 3D code.}
+\added{The proposed Fourier transformation requires a robust solver that can
+compute precise results over a wide range of frequencies (times) to obtain
+time-domain (frequency-domain) responses. In our case the solver uses a
+stretched, regular grid and computes the electric fields in the frequency
+domain. This setup is the target of the following gridding recommendations.
+Whilst they will look differently for other mesh types or a time-domain code,
+some conclusion will still hold.}
 
@@
 including the air wave, is computed with a semi-analytical code for layered
-media. We do not consider this approach here.
+media. We do not consider this approach here. \added{Note that grid stretching
+for complex-valued diffusion fields is essentially what is done for wave fields
+with the perfectly matched layer (PML). PML is an absorbing layer to avoid
+scattering from the boundary by letting the field die out to zero, which is
+achieved by introducing a complex factor which causes damping. As
+electromagnetic diffusion fields are damped fields by themselves it suffices to
+stretch the grid.}
 
@@
 \subsection{\added{Induced Polarization}} % % % % % % % % % % % % % % % % % % %
+
+\added{mention independence of spatial complexity, why we also show 1D
+examples.}
 

### 2021-02-26 ManuscriptTime.tex
EOW

@@
 
+\added{Note to reviewers: This entire section is new, without being all
+marked in green.}
+
 \added{mention independence of spatial complexity, why we also show 1D
 examples.}
+
+The Cole-Cole model \citep[CCM, ][]{JCP.41.Cole}, written in terms of
+conductivities instead of electric permittivities as by the original authors,
+is given by
+%
+\begin{equation}
+  \sigma(\omega) = \sigma_\infty + \frac{\sigma_0 - \sigma_\infty}{1 +
+    (\rm{i}\omega\tau)^c}\ ,
+  \label{eq:ccm}
+\end{equation}
+%
+where $\sigma_0$ and $\sigma_\infty$ refer to the low-frequency and
+high-frequency conductivity values, respectively, $\tau$ is the central
+relaxation time, and $c$ the CCM exponent describing the broadness of the
+relaxation time distribution. Note that this model slightly differs from the
+one phrased in terms of resistivities given by \cite{GEO.78.Pelton}, see, e.g.,
+\cite{GJI.13.Tarasov}.
+%
+\begin{figure}
+  \centering
+  \includegraphics[width=\fwidth]{11-cole-cole-model}
+  \caption{}
+  \label{fig:CCM}
+\end{figure}
+%
+
+It required only 30 frequencies from 1.5e-8 to 8.3\,Hz, instead of the 244
+frequencies from 1.5e-8 to 6.8e6\,Hz required by the digital filter (201\,pt
+filter from Key, 2012). Besides the fact that the responses for the higher
+frequencies can be set to zero it is also sufficient to only compute every
+5th of the required frequencies.
+
+low frequencies cannot be 'interpolated'; however, other tricks could be
+implemented. The DC values (so low frequencies or late times) is the same
+independent of the IP effect. This makes it possible to estimate the DC value
+with the IP-free model which does not require very low frequencies, and use
+consequently this DC value to interpolate the low frequencies.
 

### 2021-02-27 ManuscriptTime.tex
land ip

@@
 \section{Introduction}  % % % % % % % % % % % % % % % % % % % % % % % % % % % %
+
+\rtodo 1 \added{\emph{The independence from spatial model complexity is in my
+  opinion debatable, because of a potential issue with the accuracy of phase
+  information. It’s my experience that a high degree of imaginary-field
+  component accuracy is needed for a proper f-t transform. --
+  I also needed a high level of numerical accuracy in the f-domain, that is,
+  very fine meshes and tight solver tolerances. An aspect which in my opinion
+  is not addressed sufficiently in the current manuscript.}: Add somewhere that
+  high accurate results are required}.
+
+\rtodo 2 \added{\emph{In my first revision, I actually did not make a request
+  for making the literature review "more extensive". Instead, I wrote that the
+  authors "should spend more time researching the EM literature that is related
+  to the actual topic." As it was and still is, half of the introduction
+  (current manuscript, page 2 to p. 3, line 46) is a rather broad/general EM
+  overview. Given the fact that the paper connects directly to Plessix et al.
+  (2007) and Mulder et al. (2008), I would have expected a more targeted
+  introduction towards the specific advances of this manuscript – instead of
+  talking about a choice of only loosely connected and more general topics like
+  Lebedev grids, tetrahedral grids, integral equation, direct solvers, etc. I
+  like that the introduction is generally concise. My impression remains that
+  the introduction can have the same length but be more targeted. Other readers
+  may also see it that way. I won’t need another explanation about your
+  potentially differing opinion. This is just a suggestion for improvement.
+  Feel free to ignore.}}
+
+\rtodo 3 Reply; Include comments from earlier review; papers that were
+referred; difficulties in late/early times; IP effects. Refer to steel-casing
+effects (lindsey), but state we look here at non-Debye cole models.
+Add also the papers seogi mentioned re IP effects.
+\emph{We tried to improve the introduction and focus more on the actual topic.
+  Amongst other, we address the topic of steel casing and non-Debye Cole-Cole
+models.}
 
@@
 consequently this DC value to interpolate the low frequencies.
+
+for c>0.6, 17 frequencies from 1.2e-4 to 8.3\,Hz would be sufficient.
+
+%    'src': (0, 0, 0.01),
+%    'rec': (900, 0, 0.01),
+%    'depth': [0, 300, 400],
+%    res = 2e14, 1, 1, 1
+%    layer: cond_8 = 1.25, cond_0 = 1 => m = 0.2
+%    tau: 1
+%    fmin=1e-4,
+%    fmax=5e2,
+%    signal=1,        # <= Step-On!
+%    every_x_freq=6,  # Only every 6th value
+
+%  time        [s] :  0.01 - 10000 : 301  [min-max; #]
+%  Fourier         :  DLF (Sine-Filter)
+%    > Filter      :  Key 601 CosSin (2009)
+%    > DLF type    :  Lagged Convolution
+%  Req. freq  [Hz] :  6.67463E-18 - 4.00754E+13 : 747  [min-max; #]
+%  Calc. freq [Hz] :  0.000155601 - 424.87 : 27  [min-max; #]
+
+% Comparison: Full DLF; cosine-filter step-off, 747 frequencies
+% Our method: 27 freqs; DC(no IP) - step-on; DC at 1000 s
+%    (no more frequencies were required)
+%    important to see that the DC value (low freq, late t; is the same IP/noIP)
+
+The actual transform was carried out with the 601-point sine-cosine filter from
+\cite{GEO.09.Key}.
+
+Key 601\,pt filter is very long; seems to be by far the best for low c values.
+Designing a new filter specific for IP problems might be the best approach
+
+DC: 1000s; the same amount of frequencies is required to compute times up to
+10s or up to 1000s
+
+\cite{GJI.17.Commer}
+
+\begin{itemize}
+  \item Designing a DLF for IP effects
+  \item DC value is the same for the IP or non-IP model, so one could use
+    the non-IP response and do interpolation
+\end{itemize}
 

### 2021-03-20 ManuscriptTime.tex
references
WOI

@@
 effects (lindsey), but state we look here at non-Debye cole models.
-Add also the papers seogi mentioned re IP effects.
-\emph{We tried to improve the introduction and focus more on the actual topic.
-  Amongst other, we address the topic of steel casing and non-Debye Cole-Cole
-models.}
+Add also the papers seogi mentioned re IP effects. \cite{CAG.19.Heagy}
+\cite{EXP.20.Kang} \emph{We tried to improve the introduction and focus more on
+  the actual topic. Amongst other, we address the topic of steel casing and
+  non-Debye Cole-Cole models.}
 
@@
 
+\rtodo \added{REDO and send to R.R}
+
 \rtodo \added{Add citation to R.R. (2021) regarding RA and speed-up; limitation
@@
 opposite direction. We present a methodology which significantly reduces the
-amount and range of the required frequencies \added{or times}, translating into
-a significant reduction in computation time. The required computation grids for
-low frequencies (in our case in the order of 0.001\,Hz) and high frequencies
-(around 100\,Hz) are hugely different. Low frequencies can be computed on a
-coarser grid, but they require a much larger domain in order to avoid boundary
-effects. High frequencies, on the other hand, require denser gridding, but they
-are much more limited in reach. An adaptive gridding scheme is therefore
-required\replaced{, which}{. We build our approach upon
+amount and range of the required frequencies\replaced{ through a combination of
+extrapolation, interpolation, and setting the responses for high frequencies to
+zero. This translates}{, translating} into a significant reduction in
+computation time.
+\replaced{The number of required frequencies is the most important aspect for reducing computation time.
+The second most important aspect is the computation grid, as t}{T}he required \deleted{computation} grids
+for low frequencies (in our case in the order of 0.001\,Hz) and high
+frequencies (around 100\,Hz) are hugely different. Low frequencies can be
+computed on a coarser grid, but they require a much larger domain in order to
+avoid boundary effects. High frequencies, on the other hand, require denser
+gridding, but they are much more limited in reach. An adaptive gridding scheme
+is therefore required\replaced{, which}{. We build our approach upon
 \mbox{\cite{GEO.07.Plessix}}, who presented such an adaptive gridding for
@@
 approximation of Maxwell's equations in an isotropic, homogeneous medium is
-given by \citep[e.g.,][ equation~1.53]{B.SEG.88.Ward}
+given by \citep[e.g.,][\mbox{\deleted{ equation~1.53}}]{B.SEG.88.Ward}
 %
@@
 minimum skin depth is therefore defined by the conductivity of seawater.
-However, this yields quite large cells for low frequencies, so special care has
-to be taken around the source by defining a maximum allowed
-$\Delta_\mathrm{min}$. The grid dimension, on the other hand, is defined as a
-function of the skin depth for the average conductivity of the background,
-$\delta_\mathrm{ave}$. They use four times $\delta_\mathrm{ave}$ for the $x$-,
-$y$-, and downward $z$-directions, and a fixed 50\,km for the upward
-$z$-direction to account for the airwave. To reduce the number of cells, it is
-desirable to introduce stretching, at least in the buffer zone outside of the
-area where source and receivers are located.
+However, this \replaced{can yield}{yields} quite large cells for low
+frequencies, so special care has to be taken around the source by defining a
+maximum allowed $\Delta_\mathrm{min}$. The grid dimension, on the other hand,
+is defined as a function of \deleted{the} skin depth for the average
+conductivity of the background, $\delta_\mathrm{ave}$. They use four times
+$\delta_\mathrm{ave}$ for the $x$-, $y$-, and downward $z$-directions, and a
+fixed 50\,km for the upward $z$-direction to account for the airwave. To reduce
+the number of cells, it is desirable to introduce stretching, at least in the
+buffer zone outside of the area where source and receivers are located.
 
@@
 \added{The requirement for any transform is a robust solver of sufficient
-precision.} \replaced{We use f}{F}or the numerical computations \deleted{we
-use} the open-source (Apache License 2.0) code\added{s \empymod
-\mbox{\citep{GEO.17.Werthmuller}} and} \emg3d
-\citep{JOSS.19.Werthmuller}\replaced{.}{,} \added{The former computes
-semi-analytical responses of layered models, the latter is} a multigrid solver
-based on \cite{GP.06.Mulder}, which can be used as a preconditioner for Krylov
-subspace solvers or as a solver on its own. The multigrid approach works fine
-for the diffusive approximation of Maxwell's equations, which assumes that
-$\omega\varepsilon \ll \sigma$, where $\varepsilon$ is electric permittivity
-(F/m). The remaining system to solve in the frequency domain is then given by
-the second-order differential equation of the electric field,
+precision over a wide range of frequencies.} \replaced{We use f}{F}or the
+numerical computations \deleted{we use} the open-source (Apache License 2.0)
+code\added{s \empymod} \added{\mbox{\citep{GEO.17.Werthmuller}}} \added{and}
+\emg3d \citep{JOSS.19.Werthmuller}\replaced{.}{,} \added{The former computes
+semi-analytical responses of layered models, the latter is} a \added{3D}
+multigrid solver based on \cite{GP.06.Mulder}, which can be used as a
+preconditioner for Krylov subspace solvers or as a solver on its own. The
+multigrid approach works fine for the diffusive approximation of Maxwell's
+equations, which assumes that $\omega\varepsilon \ll \sigma$, where
+$\varepsilon$ is electric permittivity (F/m). The remaining system to solve in
+the frequency domain is then given by the second-order differential equation of
+the electric field,
 %
@@
 makes it feasible to run even big models on standard computers, without the
-need for big clusters. All the examples in this article are run on a laptop
-with an i7-6600U\,CPU\,@\,2.6\,GHz (x4) and 16\,GB of memory, using Ubuntu
-20.04 and Python 3.8. All examples were run using a single thread, but
-parallelization over frequencies is straight forward, as the computations are
-independent. \deleted{For the semi-analytical computations of layered models we
-use the open-source (Apache License 2.0) code \empymod
+need for big clusters. All \deleted{the} examples in this article are run on a
+laptop with an i7-6600U\,CPU\,@\,2.6\,GHz (x4) and 16\,GB of memory, using
+Ubuntu 20.04 and Python 3.8. \replaced{They}{All examples} were run using a
+single thread, but parallelization over frequencies is straight forward, as the
+computations are independent. \deleted{For the semi-analytical computations of
+layered models we use the open-source (Apache License 2.0) code \empymod
 \mbox{\citep{GEO.17.Werthmuller}}.}
@@
 \added{mention independence of spatial complexity, why we also show 1D
-examples.}
+examples; however, sufficiently accurate 3D results are an obvious
+pre-requisite.}
 

### 2021-03-30 ManuscriptTime.tex
Edits

@@
 
-\rtodo 1 \added{\emph{The independence from spatial model complexity is in my
+\added{BEGIN TEMPORARY ADDED v v v v}
+
+\rtodo 1 \emph{The independence from spatial model complexity is in my
   opinion debatable, because of a potential issue with the accuracy of phase
@@
   is not addressed sufficiently in the current manuscript.}: Add somewhere that
-  high accurate results are required}.
-
-\rtodo 2 \added{\emph{In my first revision, I actually did not make a request
+  high accurate results are required.
+
+\rtodo 2 \emph{In my first revision, I actually did not make a request
   for making the literature review "more extensive". Instead, I wrote that the
@@
   (current manuscript, page 2 to p. 3, line 46) is a rather broad/general EM
-  overview. Given the fact that the paper connects directly to Plessix et al.
-  (2007) and Mulder et al. (2008), I would have expected a more targeted
-  introduction towards the specific advances of this manuscript – instead of
-  talking about a choice of only loosely connected and more general topics like
-  Lebedev grids, tetrahedral grids, integral equation, direct solvers, etc. I
-  like that the introduction is generally concise. My impression remains that
-  the introduction can have the same length but be more targeted. Other readers
-  may also see it that way. I won’t need another explanation about your
-  potentially differing opinion. This is just a suggestion for improvement.
-  Feel free to ignore.}}
+  overview. Given the fact that the paper \textbf{connects directly to Plessix
+  et al. (2007) and Mulder et al. (2008)}, I would have expected a \textbf{more
+  targeted introduction towards the specific advances of this manuscript} –
+  instead of talking about a choice of only loosely connected and more general
+  topics like Lebedev grids, tetrahedral grids, integral equation, direct
+  solvers, etc. I like that the introduction is generally concise. My
+  impression remains that the introduction can have the same length but be more
+  targeted. Other readers may also see it that way. I won’t need another
+  explanation about your potentially differing opinion. This is just a
+  suggestion for improvement. Feel free to ignore.}
 
@@
   non-Debye Cole-Cole models.}
+
+\added{END TEMPORARY ADDED n n n}
 
@@
 frequencies change with offset and an adaptive frequency selection is therefore
-often tailored to a single offset\added{ or a limited range of offset}. Also,
+often tailored to a single offset\added{ or a limited range of offsets}. Also,
 it allows for straightforward parallelization over frequencies, which is not
@@
 
-\rtodo \added{Start this section with the method; fmin, fmax, ndec; below we
-interpolate with PCHIP, above we set to zero, in-between we interpolate with
-cubic spline.}
+\hrule
+
+\added{Any Fourier transform needs a certain range of frequencies to obtain
+time-domain results. Our approach is not to minimize the required for the
+transform, but the number of frequencies for which we actually have to compute
+responses. We do this in three zones: Above an upper threshold fmax we set the
+frequency-domain response to zero. Below a lower threshold fmin we extrapolate
+the frequency-domain response. As we know that the imaginary response goes
+towards zero for zero frequency, we can turn it into an interpolation, for
+which we use PCHIP. And in-between the thresholds we use cubic spline
+interpolation in order to only have to actually compute a few distinct
+frequencies. Using this approach we found that computing the responses for
+15--25 frequencies is usually good enough to obtain time-domain responses for a
+large range of offsets; even though the actual transform might need up to
+hundreds of frequencies. The required frequencies depend naturally on the
+conductivities and on the chosen acquisition type and layout. A trial-and-error
+\dots}
+
+\hrule
+
 \replaced{The actually computed frequencies in our approach are defined by

### 2021-04-01 ManuscriptTime.tex
On it

@@
 zero. This translates}{, translating} into a significant reduction in
-computation time.
-\replaced{The number of required frequencies is the most important aspect for reducing computation time.
-The second most important aspect is the computation grid, as t}{T}he required \deleted{computation} grids
+computation time. \replaced{The number of required frequencies is not the only
+important point for reducing computation time. Another important aspect is the computation grid, as t}{T}he required \deleted{computation} grids
 for low frequencies (in our case in the order of 0.001\,Hz) and high
@@
 
+\added{Note that while we use these two algorithms for the numerical examples
+any solver that provides sufficient accurate and precise frequency-domain
+responses could be used.}
+
 \subsection{Frequency selection}  % % % % % % % % % % % % % % % % % % % % % % %
@@
 frequencies. The fewer frequencies required, the quicker we obtain the
-time-domain result. \added{This applies primarily for expensive kernels such
-as used in 3D modelling where the computation of the response of a single
-frequency takes much longer than the transformation itself.} The same applies
-for the reverse operation, the fewer times required, the quicker we obtain
-frequency-domain results. Even though we limit our analysis here to
-frequency-to-time transformations, most of the arguments are reversible and
-applicable to time-to-frequency transformations. We decided to use a regular
-spacing of frequencies on a log-scale, rather than an adaptive scheme. This
-approach is favourable if a wide range of offsets is needed, as the required
-frequencies change with offset and an adaptive frequency selection is therefore
-often tailored to a single offset\added{ or a limited range of offsets}. Also,
-it allows for straightforward parallelization over frequencies, which is not
-completely possible with an adaptive scheme.
+time-domain result. \added{This is true in general but applies in particular
+to expensive kernels such as is the case in 3D modelling, where the computation
+of the response of a single frequency takes much longer than the transformation
+itself.} The same applies for the reverse operation, the fewer times required,
+the quicker we obtain frequency-domain results. Even though we limit our
+analysis here to frequency-to-time transformations, most of the arguments are
+reversible and applicable to time-to-frequency transformations. We decided to
+use a regular spacing of frequencies on a log-scale, rather than an adaptive
+scheme. This approach is favourable if a wide range of offsets is needed, as
+the required frequencies change with offset and an adaptive frequency selection
+is therefore often tailored to a single offset\added{ or a limited range of
+offsets}. Also, it allows for straightforward parallelization over frequencies,
+which is not completely possible with an adaptive scheme.
 
@@
 
-\hrule
-
-\added{Any Fourier transform needs a certain range of frequencies to obtain
-time-domain results. Our approach is not to minimize the required for the
-transform, but the number of frequencies for which we actually have to compute
-responses. We do this in three zones: Above an upper threshold fmax we set the
-frequency-domain response to zero. Below a lower threshold fmin we extrapolate
-the frequency-domain response. As we know that the imaginary response goes
-towards zero for zero frequency, we can turn it into an interpolation, for
-which we use PCHIP. And in-between the thresholds we use cubic spline
-interpolation in order to only have to actually compute a few distinct
-frequencies. Using this approach we found that computing the responses for
-15--25 frequencies is usually good enough to obtain time-domain responses for a
-large range of offsets; even though the actual transform might need up to
-hundreds of frequencies. The required frequencies depend naturally on the
-conductivities and on the chosen acquisition type and layout. A trial-and-error
-\dots}
-
-\hrule
-
-\replaced{The actually computed frequencies in our approach are defined by
-three key-parameters}{There are three key-parameters that have to be defined}:
-\replaced{The m}{M}inimum and maximum frequency ($f_\mr{min}, f_\mr{max}$) and
-the number of frequencies per decade ($n_\mr{dec}^f$). \added{This creates a
-frequency range which is much smaller than the frequency range required for the
-transform.} \replaced{A t}{T}rial-and-error \added{approach} with a 3D code is
-very time-intensive. However, a simplified, layered model for the required
-survey setup and a fast 1D modeller makes it possible to estimate these
-parameters easily. \replaced{An example of this is shown in
+\replaced{Any Fourier transform needs a certain range of frequencies to obtain
+time-domain results. Our approach is not to minimize the required frequencies
+for the transform, but to minimize the number of frequencies for which we
+actually have to compute responses. We do this in three zones: Above an upper
+threshold $f_\mr{max}$ we set the frequency-domain response to zero. Below a
+lower threshold $f_\mr{min}$ we extrapolate the frequency-domain response. As
+we know that the imaginary response goes towards zero for zero frequency, we
+can turn it into an interpolation, for which we use PCHIP. And in-between the
+thresholds we use cubic spline interpolation in order to only have to actually
+compute a few distinct frequencies per decade ($n_\mr{dec}^f$). Using this
+approach we found that computing the responses for 15--25 frequencies is
+usually good enough to obtain time-domain responses for a large range of
+offsets; even though the actual transform might need up to hundreds of
+frequencies.}{There are three key-parameters that have to be defined: Minimum
+and maximum frequency ($f_\mr{min}, f_\mr{max}$) and the number of frequencies
+per decade ($n_\mr{dec}^f$).}
+
+\replaced{The required frequencies depend naturally on the conductivities and
+on the chosen acquisition type and layout. A t}{T}rial-and-error approach with a
+3D code is very time-intensive. However, a simplified, layered model for the
+required survey setup and a fast 1D modeller makes it possible to estimate
+these parameters easily. \replaced{An example of this is shown in
 Figure~\ref{fig:FFTLog-log}.}{To this end, we created a graphical, interactive
@@
 %
-\replaced{It }{The shown example model} is a marine scenario with 1\,km water
-depth of resistivity $\rho=0.3\,\ohmm$ ($\rho = \sigma^{-1}$) and a 100\,m
-thick target of 100\,\ohmm at 1\,km below the seafloor in a background of
-1\,\ohmm. The source is 50\,m above the seafloor, the receiver is on the
-seafloor, and the response is the inline $x$-directed $E$-field. The left
-subplot shows the imaginary part in the frequency domain and the right plot the
-corresponding impulse response in the time domain. The red lines are the
-semi-analytical responses. The blue circles indicate the actually computed
-responses and the black dots the frequencies which are interpolated or set to
-zero. The resulting time-domain response has a relative error of less than
-1\,\% everywhere except for very early times.
+\replaced{It shows the responses for}{The shown example model is} a
+marine scenario with 1\,km water depth of resistivity $\rho=0.3\,\ohmm$ ($\rho
+= \sigma^{-1}$) and a 100\,m thick target of 100\,\ohmm at 1\,km below the
+seafloor in a background of 1\,\ohmm. The source is 50\,m above the seafloor,
+the receiver is on the seafloor, and the response is the inline $x$-directed
+$E$-field. The left subplot shows the imaginary part in the frequency domain
+and the right plot the corresponding impulse response in the time domain. The
+red lines are the semi-analytical responses. The blue circles indicate the
+actually computed responses and the black dots the frequencies which are
+interpolated or set to zero. The resulting time-domain response has a relative
+error of less than 1\,\% everywhere except for very early times.
 
 Figure \ref{fig:FFTLog-lin} shows exactly the same on a
-\replaced{linear}{logarithmic} scale\replaced{for the time and amplitude
+\replaced{linear}{logarithmic} scale \replaced{for the time and amplitude
 axes}{, without the interactive GUI-related elements}. It can be seen that with
-the chosen frequencies the time-domain response starts to divert above about
-100\,s and below 0.3\,s. It also shows the oscillating high-frequency part,
-which is hard to interpolate and we therefore set it to zero.
-Figure~\ref{fig:DLF-lin} is the same as Figure \ref{fig:FFTLog-lin}, but
-transformed with the DLF method applying the 81-point sine-cosine filter from
-\cite{GEO.09.Key}. The same frequencies were computed as in the FFTLog case and
-the missing ones interpolated. The error of the corresponding time-domain
-response is comparable, so either FFTLog or DLF can be used.
+the chosen\added{, computed} frequencies \added{(blue dots)} the time-domain
+response starts to divert above about 100\,s and below 0.3\,s. It also shows
+the oscillating high-frequency part, which is hard to interpolate and we
+therefore set it to zero. Figure~\ref{fig:DLF-lin} is the same as Figure
+\ref{fig:FFTLog-lin}, but transformed with the DLF method applying the 81-point
+sine-cosine filter from \cite{GEO.09.Key}. The same frequencies were computed
+as in the FFTLog case and the missing ones interpolated. The error of the
+corresponding time-domain response is comparable, so either FFTLog or DLF can
+be used.
 %
@@
 required for the shown offset of 5\,km, a few of the lower and higher
-frequencies could be left out. With this frequency-selection we can model a
-wide range of offsets. This is shown in Figure \ref{fig:multi-offset}, where
-the same kernels were used to yield the responses at offsets $r=1.5,3,6$, and
-12\,km. We only need to compute 50\,\% of the required frequencies in the case
-of the FFTLog, and only 14\,\% in the case of the DLF.
+frequencies could be left out. With this frequency-selection\added{, however,}
+we can model a wide range of offsets. This is shown in Figure
+\ref{fig:multi-offset}, where the same kernels were used to yield the responses
+at offsets $r=1.5,3,6$, and 12\,km. We only need to compute 50\,\% of the
+required frequencies in the case of the FFTLog, and only 14\,\% in the case of
+the DLF.
 %
@@
 time-domain (frequency-domain) responses. In our case the solver uses a
-stretched, regular grid and computes the electric fields in the frequency
+regular, stretched grid and computes the electric fields in the frequency
 domain. This setup is the target of the following gridding recommendations.
@@
 should be per skin depth. The actual computational domain $D_\mr{c}$ is usually
-much bigger than $D_\mr{s}$ in order to avoid artefacts from the perfectly
-electrically conducting boundary condition. It can also have a much higher
-stretching factor $\alpha_\mr{c}$. In the presented examples we have chosen
-$D_\mr{c}$ such that the distance for the signal diffusing from the source to
-the boundary and back to the receiver closest to the boundary is at least two
-wavelengths, after which the amplitude of the signal is reduced to a millionth
-of its initial strength. The wavelength $\lambda$ to compute $D_\mr{c}$ is
-given by
+much bigger than $D_\mr{s}$ in order to avoid artefacts from the
+\deleted{perfectly electrically conducting} boundary condition. It can also
+have a much higher stretching factor $\alpha_\mr{c}$. In the presented examples
+we have chosen $D_\mr{c}$ such that the distance for the signal diffusing from
+the source to the boundary and back to the receiver closest to the boundary is
+at least two wavelengths, after which the amplitude of the signal is reduced to
+a millionth of its initial strength. The wavelength $\lambda$ to compute
+$D_\mr{c}$ is given by
 %
@@
 the source to the boundary; this distance is by default set to 100\,km, but
-this can be reduced in the marine case with increasing water depth. Note that
-for shallow marine and land cases this also applies to the horizontal
-dimensions, not only to the upward $z$-direction and similarly for very
+this can be reduced in the marine case with increasing water depth.
+\deleted{Note that for shallow marine and land cases t}\added{T}his also
+applies to the horizontal dimensions, not only to the upward
+$z$-direction\added{,} and \replaced{it equally applies}{similarly} for very
 resistive basements, even in deep water. One way to circumvent this difficulty
@@
 \section{Numerical examples}  % % % % % % % % % % % % % % % % % % % % % % % % %
+
+\todo[inline]{HERE HERE HERE}
 
@@
 
+\subsection{\added{Induced Polarization}} % % % % % % % % % % % % % % % % % % %
+
+\added{$\Downarrow\qquad$ Note to reviewers: The entire section 4.3 below is
+new, without being all marked in green. $\qquad\Downarrow$}
+
+\added{Mention independence of spatial complexity, why we also show 1D
+examples; however, sufficiently accurate 3D results are an obvious
+pre-requisite.}
+
+The Cole-Cole model \citep[CCM, ][]{JCP.41.Cole}, written in terms of
+conductivities instead of electric permittivities as by the original authors,
+is given by
+%
+\begin{equation}
+  \sigma(\omega) = \sigma_\infty + \frac{\sigma_0 - \sigma_\infty}{1 +
+    (\rm{i}\omega\tau)^c}\ ,
+  \label{eq:ccm}
+\end{equation}
+%
+where $\sigma_0$ and $\sigma_\infty$ refer to the low-frequency and
+high-frequency conductivity values, respectively, $\tau$ is the central
+relaxation time, and $c$ the CCM exponent describing the broadness of the
+relaxation time distribution. Note that this model slightly differs from the
+one phrased in terms of resistivities given by \cite{GEO.78.Pelton}, see, e.g.,
+\cite{GJI.13.Tarasov}.
+%
+\begin{figure}
+  \centering
+  \includegraphics[width=\fwidth]{11-cole-cole-model}
+  \caption{}
+  \label{fig:CCM}
+\end{figure}
+%
+
+It required only 30 frequencies from 1.5e-8 to 8.3\,Hz, instead of the 244
+frequencies from 1.5e-8 to 6.8e6\,Hz required by the digital filter (201\,pt
+filter from Key, 2012). Besides the fact that the responses for the higher
+frequencies can be set to zero it is also sufficient to only compute every
+5th of the required frequencies.
+
+low frequencies cannot be 'interpolated'; however, other tricks could be
+implemented. The DC values (so low frequencies or late times) is the same
+independent of the IP effect. This makes it possible to estimate the DC value
+with the IP-free model which does not require very low frequencies, and use
+consequently this DC value to interpolate the low frequencies.
+
+for c>0.6, 17 frequencies from 1.2e-4 to 8.3\,Hz would be sufficient.
+
+%    'src': (0, 0, 0.01),
+%    'rec': (900, 0, 0.01),
+%    'depth': [0, 300, 400],
+%    res = 2e14, 1, 1, 1
+%    layer: cond_8 = 1.25, cond_0 = 1 => m = 0.2
+%    tau: 1
+%    fmin=1e-4,
+%    fmax=5e2,
+%    signal=1,        # <= Step-On!
+%    every_x_freq=6,  # Only every 6th value
+
+%  time        [s] :  0.01 - 10000 : 301  [min-max; #]
+%  Fourier         :  DLF (Sine-Filter)
+%    > Filter      :  Key 601 CosSin (2009)
+%    > DLF type    :  Lagged Convolution
+%  Req. freq  [Hz] :  6.67463E-18 - 4.00754E+13 : 747  [min-max; #]
+%  Calc. freq [Hz] :  0.000155601 - 424.87 : 27  [min-max; #]
+
+% Comparison: Full DLF; cosine-filter step-off, 747 frequencies
+% Our method: 27 freqs; DC(no IP) - step-on; DC at 1000 s
+%    (no more frequencies were required)
+%    important to see that the DC value (low freq, late t; is the same IP/noIP)
+
+The actual transform was carried out with the 601-point sine-cosine filter from
+\cite{GEO.09.Key}.
+
+Key 601\,pt filter is very long; seems to be by far the best for low c values.
+Designing a new filter specific for IP problems might be the best approach
+
+DC: 1000s; the same amount of frequencies is required to compute times up to
+10s or up to 1000s
+
+\cite{GJI.17.Commer}
+
+\begin{itemize}
+  \item Designing a DLF for IP effects
+  \item DC value is the same for the IP or non-IP model, so one could use
+    the non-IP response and do interpolation
+\end{itemize}
+
+\added{$\Uparrow\qquad$ Note to reviewers: The entire section 4.3 above is new,
+without being all marked in green. $\qquad\Uparrow$}
+
 \subsection{Horizontal extent of the computation domain}  % % % % % % % % % % %
@@
 iterative solver vs. direct solver.
-
-\subsection{\added{Induced Polarization}} % % % % % % % % % % % % % % % % % % %
-
-\added{Note to reviewers: This entire section is new, without being all
-marked in green.}
-
-\added{mention independence of spatial complexity, why we also show 1D
-examples; however, sufficiently accurate 3D results are an obvious
-pre-requisite.}
-
-The Cole-Cole model \citep[CCM, ][]{JCP.41.Cole}, written in terms of
-conductivities instead of electric permittivities as by the original authors,
-is given by
-%
-\begin{equation}
-  \sigma(\omega) = \sigma_\infty + \frac{\sigma_0 - \sigma_\infty}{1 +
-    (\rm{i}\omega\tau)^c}\ ,
-  \label{eq:ccm}
-\end{equation}
-%
-where $\sigma_0$ and $\sigma_\infty$ refer to the low-frequency and
-high-frequency conductivity values, respectively, $\tau$ is the central
-relaxation time, and $c$ the CCM exponent describing the broadness of the
-relaxation time distribution. Note that this model slightly differs from the
-one phrased in terms of resistivities given by \cite{GEO.78.Pelton}, see, e.g.,
-\cite{GJI.13.Tarasov}.
-%
-\begin{figure}
-  \centering
-  \includegraphics[width=\fwidth]{11-cole-cole-model}
-  \caption{}
-  \label{fig:CCM}
-\end{figure}
-%
-
-It required only 30 frequencies from 1.5e-8 to 8.3\,Hz, instead of the 244
-frequencies from 1.5e-8 to 6.8e6\,Hz required by the digital filter (201\,pt
-filter from Key, 2012). Besides the fact that the responses for the higher
-frequencies can be set to zero it is also sufficient to only compute every
-5th of the required frequencies.
-
-low frequencies cannot be 'interpolated'; however, other tricks could be
-implemented. The DC values (so low frequencies or late times) is the same
-independent of the IP effect. This makes it possible to estimate the DC value
-with the IP-free model which does not require very low frequencies, and use
-consequently this DC value to interpolate the low frequencies.
-
-for c>0.6, 17 frequencies from 1.2e-4 to 8.3\,Hz would be sufficient.
-
-%    'src': (0, 0, 0.01),
-%    'rec': (900, 0, 0.01),
-%    'depth': [0, 300, 400],
-%    res = 2e14, 1, 1, 1
-%    layer: cond_8 = 1.25, cond_0 = 1 => m = 0.2
-%    tau: 1
-%    fmin=1e-4,
-%    fmax=5e2,
-%    signal=1,        # <= Step-On!
-%    every_x_freq=6,  # Only every 6th value
-
-%  time        [s] :  0.01 - 10000 : 301  [min-max; #]
-%  Fourier         :  DLF (Sine-Filter)
-%    > Filter      :  Key 601 CosSin (2009)
-%    > DLF type    :  Lagged Convolution
-%  Req. freq  [Hz] :  6.67463E-18 - 4.00754E+13 : 747  [min-max; #]
-%  Calc. freq [Hz] :  0.000155601 - 424.87 : 27  [min-max; #]
-
-% Comparison: Full DLF; cosine-filter step-off, 747 frequencies
-% Our method: 27 freqs; DC(no IP) - step-on; DC at 1000 s
-%    (no more frequencies were required)
-%    important to see that the DC value (low freq, late t; is the same IP/noIP)
-
-The actual transform was carried out with the 601-point sine-cosine filter from
-\cite{GEO.09.Key}.
-
-Key 601\,pt filter is very long; seems to be by far the best for low c values.
-Designing a new filter specific for IP problems might be the best approach
-
-DC: 1000s; the same amount of frequencies is required to compute times up to
-10s or up to 1000s
-
-\cite{GJI.17.Commer}
-
-\begin{itemize}
-  \item Designing a DLF for IP effects
-  \item DC value is the same for the IP or non-IP model, so one could use
-    the non-IP response and do interpolation
-\end{itemize}
 

### 2021-04-03 ManuscriptTime.tex
Draft RevII

@@
     referee,     % uncomment for double lines
-    % final,       % uncomment for changes
+    final,       % uncomment for changes
 ]{gji}
@@
 \section{Introduction}  % % % % % % % % % % % % % % % % % % % % % % % % % % % %
-
-\added{BEGIN TEMPORARY ADDED v v v v}
-
-\rtodo 1 \emph{The independence from spatial model complexity is in my
-  opinion debatable, because of a potential issue with the accuracy of phase
-  information. It’s my experience that a high degree of imaginary-field
-  component accuracy is needed for a proper f-t transform. --
-  I also needed a high level of numerical accuracy in the f-domain, that is,
-  very fine meshes and tight solver tolerances. An aspect which in my opinion
-  is not addressed sufficiently in the current manuscript.}: Add somewhere that
-  high accurate results are required.
-
-\rtodo 2 \emph{In my first revision, I actually did not make a request
-  for making the literature review "more extensive". Instead, I wrote that the
-  authors "should spend more time researching the EM literature that is related
-  to the actual topic." As it was and still is, half of the introduction
-  (current manuscript, page 2 to p. 3, line 46) is a rather broad/general EM
-  overview. Given the fact that the paper \textbf{connects directly to Plessix
-  et al. (2007) and Mulder et al. (2008)}, I would have expected a \textbf{more
-  targeted introduction towards the specific advances of this manuscript} –
-  instead of talking about a choice of only loosely connected and more general
-  topics like Lebedev grids, tetrahedral grids, integral equation, direct
-  solvers, etc. I like that the introduction is generally concise. My
-  impression remains that the introduction can have the same length but be more
-  targeted. Other readers may also see it that way. I won’t need another
-  explanation about your potentially differing opinion. This is just a
-  suggestion for improvement. Feel free to ignore.}
-
-\rtodo 3 Reply; Include comments from earlier review; papers that were
-referred; difficulties in late/early times; IP effects. Refer to steel-casing
-effects (lindsey), but state we look here at non-Debye cole models.
-Add also the papers seogi mentioned re IP effects. \cite{CAG.19.Heagy}
-\cite{EXP.20.Kang} \emph{We tried to improve the introduction and focus more on
-  the actual topic. Amongst other, we address the topic of steel casing and
-  non-Debye Cole-Cole models.}
-
-\added{END TEMPORARY ADDED n n n}
 
@@
 transform if time-domain data are required \citep[e.g., ][]{GEO.15.Hunziker}.
-
 CSEM codes for 2D and 3D computations, on the other hand, mostly compute their
-responses directly in the required domain, either frequency or time. There is a
-wealth of 3D electromagnetic codes, with a first boom in the 1990's;
-\cite{B.SEG.99.Oristaglio} provide a good review of the state of art at that
-time. A second wave occurred during the CSEM boom in the hydrocarbon industry
-in the early 21th century, see, e.g., \cite{SG.05.Avdeev} and
-\cite{SG.10.Borner} for comprehensive overviews. Recent publications after
-these review articles span the width of numerical methods from integral
-equation codes \citep{MGS.17.Kruglyakov} and differential equation methods such
-as finite differences \citep{GEO.10.Mittet, GEO.11.Zaslavsky, CAG.13.Sommer},
-finite elements \citep{GJI.13.Grayver}, and finite volumes
-\citep{GEO.14.Jahandari}. Besides the different methods there are also many
-different types of discretisation. Commonly used, besides rectilinear Yee grids
-\citep{IEEE.66.Yee} as used by our code, are Lebedev grids
-\citep{CMMP.64.Lebedev} but also unstructured grids using, for instance,
-tetrahedra \citep{CAG.17.Cai}. The computational requirement to model CSEM data
-also depends heavily on the technique used to solve the system. There are
-direct solvers \citep{GEO.15.Grayver}, indirect solvers
-\citep{GJI.15.Jaysaval}, and more recently also hybrid solvers
-\citep{GEO.18.Liu}.
+responses directly in the required domain, either frequency or time.
+
+\added{Electromagnetic methods in geophysics span a wide range of model
+scenarios and acquisition layouts which have all different implications. A
+well-known case is for instance the land or shallow marine case versus the deep
+marine case, where in the former case a dominating airwave has to be accurately
+modelled which can can be completely ignored in the latter case. Recent areas
+of particular interest in time-domain modelling that pose numerical challenges
+are, for instance, simulating the fields through steel-cased wells
+\mbox{\citep{CAG.19.Heagy}} or the effects of induced polarization
+\mbox{\citep{EXP.20.Kang}}. The former is challenging because of the high
+conductivity contrasts requiring very detailed meshing, the latter is
+challenging for time-domain modelling because the models are frequency
+dependent.}
+
+\deleted{There is a wealth of 3D electromagnetic codes, with a first boom in the
+1990's; \mbox{\cite{B.SEG.99.Oristaglio}} provide a good review of the state of
+art at that time. A second wave occurred during the CSEM boom in the
+hydrocarbon industry in the early 21th century, see, e.g.,
+\mbox{\cite{SG.05.Avdeev}} and \mbox{\cite{SG.10.Borner}} for comprehensive
+overviews. Recent publications after these review articles span the width of
+numerical methods from integral equation codes \mbox{\citep{MGS.17.Kruglyakov}}
+and differential equation methods such as finite differences
+\mbox{\citep{GEO.10.Mittet, GEO.11.Zaslavsky, CAG.13.Sommer}}, finite elements
+\mbox{\citep{GJI.13.Grayver}}, and finite volumes
+\mbox{\citep{GEO.14.Jahandari}}. Besides the different methods there are also
+many different types of discretisation. Commonly used, besides rectilinear Yee
+grids \mbox{\citep{IEEE.66.Yee}} as used by our code, are Lebedev grids
+\mbox{\citep{CMMP.64.Lebedev}} but also unstructured grids using, for instance,
+tetrahedra \mbox{\citep{CAG.17.Cai}}. The computational requirement to model
+CSEM data also depends heavily on the technique used to solve the system. There
+are direct solvers \mbox{\citep{GEO.15.Grayver}}, indirect solvers
+\mbox{\citep{GJI.15.Jaysaval}}, and more recently also hybrid solvers
+\mbox{\citep{GEO.18.Liu}}.}
 
@@
 data with a frequency-domain code, as shown by \cite{GEO.08.Mulder}. They
-compared the computational complexity of explicit time-stepping schemes (Du
-Fort-Frankel method as used in, e.g., \cite{GEO.04.Commer, GEO.07.Maao});
-implicit schemes \citep[e.g., ][]{GEO.04.Haber}, and matrix exponentials and
-Lanczos reduction schemes \cite[e.g., ][]{RS.94.Druskin} with the Fourier
-transform of frequency-domain solutions. Their conclusion was that the Fourier
-transform method can be favourable if the number of required frequencies is
-small relative to the number of grid points in each coordinate. The conditions
-for fast computation of time-domain data with a frequency-domain code are: a
+compared the computational complexity \added{of different time-domain methods
+with the computational complexity of frequency-domain computation followed by a
+Fourier transform:} \deleted{of} explicit time-stepping schemes (Du
+Fort-Frankel method as used in, e.g., \cite{GEO.04.Commer,
+GEO.07.Maao})\replaced{ and}{;} implicit schemes \citep[e.g.,
+][]{GEO.04.Haber}\added{ have a complexity of $O(n^4)$}, \deleted{and} matrix
+exponentials and Lanczos reduction schemes \cite[e.g., ][]{RS.94.Druskin}
+\added{have a complexity of $O(n^4\sqrt{\log n})$}\replaced{, and}{with the}
+Fourier transform of frequency-domain solutions \added{have a complexity $O(n_f
+n^3)$}; \added{$n$ is the number of grid points in each direction, and $n_f$
+the number of frequencies}. Their conclusion was that the Fourier transform
+method can be favourable if the number of required frequencies is small
+relative to the number of grid points in each coordinate. The conditions for
+fast computation of time-domain data with a frequency-domain code are: a
 sufficiently powerful solver, appropriate frequency selection and
@@
 these results but improve the run time from hours to minutes. The main reasons
-for this significant speed-up are an adaptive, frequency-dependent gridding
-scheme that minimizes the required cells in each dimension, and a logarithmic
-Fourier transform such as digital linear filters \citep[DLF, ][]{GP.71.Ghosh}
-or the logarithmic fast Fourier transform \citep[FFTLog, ][]{RAS.00.Hamilton}
-to go from the frequency to the time domain. The latter makes it also possible
-to only use the imaginary part of the frequency-domain response, which has
-advantages when it comes to interpolation.
+for this significant speed-up are \added{the further reduction of required
+frequencies by introducing lower and upper thresholds of numerically important
+frequencies,} an adaptive, frequency-dependent gridding scheme that minimizes
+the required cells in each dimension, and a logarithmic Fourier transform such
+as digital linear filters \citep[DLF, ][]{GP.71.Ghosh} or the logarithmic fast
+Fourier transform \citep[FFTLog, ][]{RAS.00.Hamilton} to go from the frequency
+to the time domain. The latter makes it also possible to only use the imaginary
+part of the frequency-domain response, which has advantages when it comes to
+interpolation.
 
@@
 demonstrate the efficiency of the approach with some numerical results.
-
-\rtodo \added{REDO and send to R.R}
-
-\rtodo \added{Add citation to R.R. (2021) regarding RA and speed-up; limitation
-to switch-on/-off.}
-
-\rtodo \added{Add notion about preference for f-domain for IP problems}
-
-\rtodo \added{Extra example: No-IP vs IP; land, 1D}
 
@@
 applied to the transform from time-domain to frequency-domain data for a code
-that solves Maxwell's equations in the time domain. \deleted{A variation of our
-method was implemented, for instance, by \custem \mbox{\citep{GEO.19.Rochlitz}}
-to obtain frequency-domain data from time-domain computations (Rochlitz,
-private communication, 2020).}
+that solves Maxwell's equations in the time domain. \added{Our recommendations
+for speeding up the Fourier transform for expensive kernels are independent of
+spatial complexity, as diffusive EM responses are smooth functions of both
+frequency and time. However, the intrinsic requirement is an accurate solver
+over a sufficient wide range of frequencies or times.} \deleted{A variation of
+our method was implemented, for instance, by \custem
+\mbox{\citep{GEO.19.Rochlitz}} to obtain frequency-domain data from time-domain
+computations (Rochlitz, private communication, 2020).}
 
@@
 \added{The proposed Fourier transformation requires a robust solver that can
-compute precise results over a wide range of frequencies (times) to obtain
+compute accurate results over a wide range of frequencies (times) to obtain
 time-domain (frequency-domain) responses. In our case the solver uses a
@@
 
-\todo[inline]{HERE HERE HERE}
-
 \subsection{Homogeneous space}  % % % % % % % % % % % % % % % % % % % % % % % %
@@
 
-Figure~\ref{fig:marine} shows the result for an offset of 5\,km, in (a) the
-frequency and (b) the time domain. The recovered response with the 3D code
-captures the airwave (first peak) and the subsurface (second peak) very
-accurately. At later times the error starts to increase. We would need to
-compute a few additional lower frequencies if we want to improve it.
+Figure~\ref{fig:marine} shows the result for an offset of 5\,km, \added{(a)} in
+\deleted{(a)} the frequency \added{domain} and (b) \added{in} the time domain.
+The recovered response with the 3D code captures the airwave (first peak) and
+the subsurface (second peak) very accurately. At later times the error starts
+to increase. We would need to compute a few additional lower frequencies if we
+want to improve it.
 %
@@
 %
-
-\subsection{\added{Induced Polarization}} % % % % % % % % % % % % % % % % % % %
-
-\added{$\Downarrow\qquad$ Note to reviewers: The entire section 4.3 below is
-new, without being all marked in green. $\qquad\Downarrow$}
-
-\added{Mention independence of spatial complexity, why we also show 1D
-examples; however, sufficiently accurate 3D results are an obvious
-pre-requisite.}
-
-The Cole-Cole model \citep[CCM, ][]{JCP.41.Cole}, written in terms of
-conductivities instead of electric permittivities as by the original authors,
-is given by
-%
-\begin{equation}
-  \sigma(\omega) = \sigma_\infty + \frac{\sigma_0 - \sigma_\infty}{1 +
-    (\rm{i}\omega\tau)^c}\ ,
-  \label{eq:ccm}
-\end{equation}
-%
-where $\sigma_0$ and $\sigma_\infty$ refer to the low-frequency and
-high-frequency conductivity values, respectively, $\tau$ is the central
-relaxation time, and $c$ the CCM exponent describing the broadness of the
-relaxation time distribution. Note that this model slightly differs from the
-one phrased in terms of resistivities given by \cite{GEO.78.Pelton}, see, e.g.,
-\cite{GJI.13.Tarasov}.
-%
-\begin{figure}
-  \centering
-  \includegraphics[width=\fwidth]{11-cole-cole-model}
-  \caption{}
-  \label{fig:CCM}
-\end{figure}
-%
-
-It required only 30 frequencies from 1.5e-8 to 8.3\,Hz, instead of the 244
-frequencies from 1.5e-8 to 6.8e6\,Hz required by the digital filter (201\,pt
-filter from Key, 2012). Besides the fact that the responses for the higher
-frequencies can be set to zero it is also sufficient to only compute every
-5th of the required frequencies.
-
-low frequencies cannot be 'interpolated'; however, other tricks could be
-implemented. The DC values (so low frequencies or late times) is the same
-independent of the IP effect. This makes it possible to estimate the DC value
-with the IP-free model which does not require very low frequencies, and use
-consequently this DC value to interpolate the low frequencies.
-
-for c>0.6, 17 frequencies from 1.2e-4 to 8.3\,Hz would be sufficient.
-
-%    'src': (0, 0, 0.01),
-%    'rec': (900, 0, 0.01),
-%    'depth': [0, 300, 400],
-%    res = 2e14, 1, 1, 1
-%    layer: cond_8 = 1.25, cond_0 = 1 => m = 0.2
-%    tau: 1
-%    fmin=1e-4,
-%    fmax=5e2,
-%    signal=1,        # <= Step-On!
-%    every_x_freq=6,  # Only every 6th value
-
-%  time        [s] :  0.01 - 10000 : 301  [min-max; #]
-%  Fourier         :  DLF (Sine-Filter)
-%    > Filter      :  Key 601 CosSin (2009)
-%    > DLF type    :  Lagged Convolution
-%  Req. freq  [Hz] :  6.67463E-18 - 4.00754E+13 : 747  [min-max; #]
-%  Calc. freq [Hz] :  0.000155601 - 424.87 : 27  [min-max; #]
-
-% Comparison: Full DLF; cosine-filter step-off, 747 frequencies
-% Our method: 27 freqs; DC(no IP) - step-on; DC at 1000 s
-%    (no more frequencies were required)
-%    important to see that the DC value (low freq, late t; is the same IP/noIP)
-
-The actual transform was carried out with the 601-point sine-cosine filter from
-\cite{GEO.09.Key}.
-
-Key 601\,pt filter is very long; seems to be by far the best for low c values.
-Designing a new filter specific for IP problems might be the best approach
-
-DC: 1000s; the same amount of frequencies is required to compute times up to
-10s or up to 1000s
-
-\cite{GJI.17.Commer}
-
-\begin{itemize}
-  \item Designing a DLF for IP effects
-  \item DC value is the same for the IP or non-IP model, so one could use
-    the non-IP response and do interpolation
-\end{itemize}
-
-\added{$\Uparrow\qquad$ Note to reviewers: The entire section 4.3 above is new,
-without being all marked in green. $\qquad\Uparrow$}
 
@@
 is actually wrong. A possibility to detect such problems for complicated cases,
-where there is no semi-analytical result to compare with, is to compute two or
-more models, moving the boundary. When the responses stop to change, one can
-assume that the boundary is far enough. Another possibility is to look at the
-amplitudes close to the boundaries and ensure that they are small enough.
+where there is no semi-analytical result to compare with, is to \replaced{carry
+out a mesh-convergence test by computing}{compute} two or more models, moving
+the boundary. When the responses stop to change, one can assume that the
+boundary is far enough. Another possibility is to look at the amplitudes close
+to the boundaries and ensure that they are small enough.
+
+\subsection{\added{Induced Polarization}} % % % % % % % % % % % % % % % % % % %
+
+\added{The proposed Fourier-transform approach works independently of spatial
+complexity, as diffusive electromagnetic fields are smooth functions of both
+frequency and time. An interesting test it to see whether this applies as
+well for dispersive media where the model parameters are frequency dependent,
+such as is the case in induced polarization. To test this we use the Cole-Cole
+model \mbox{\citep[CCM, ][]{JCP.41.Cole}} which, written in terms of
+conductivities instead of electric permittivities as by the original authors,
+is given by}
+%
+\begin{equation}
+\added{
+  \sigma(\omega) = \sigma_\infty + \frac{\sigma_0 - \sigma_\infty}{1 +
+    (\rm{i}\omega\tau)^c}\ ,
+}
+  \label{eq:ccm}
+\end{equation}
+%
+\added{where $\sigma_0$ and $\sigma_\infty$ refer to the low-frequency and
+high-frequency conductivity values, respectively, $\tau$ is the central
+relaxation time (s), and $c$ the CCM exponent describing the broadness of the
+relaxation time distribution. Note that this model slightly differs from the
+one phrased in terms of resistivities given by \mbox{\cite{GEO.78.Pelton}},
+see, e.g., \mbox{\cite{GJI.13.Tarasov}}. }
+
+\added{\mbox{\cite{GJI.17.Commer}}, have shown that time-domain IP modelling
+can be difficult, particularly for low $c$-values. Modelling
+frequency-dependent models in the frequency domain is for any given frequency
+naturally no different than modelling a non frequency-dependent model. However,
+the interesting point is to see how the Fourier transform behaves to obtain
+time-domain data from frequency-domain responses. As an example we modelled the
+response for a land case, air above a halfspace of 1\,S/m. Within the halfspace
+is a dispersive, 100\,m thick layer at a depth of 300\,m, with
+$\sigma_0=1.0\,$S/m and $\sigma_\infty=1.25\,$S/m; $\tau=1$, and we vary
+$c$ from 0 (no IP) to 1.0, 0.75, 0.5, and 0.25. The x-directed step-off source
+and the receiver at an inline offset of 900\,m are place at the surface. The
+real and imaginary frequency-domain responses for this model are shown in the
+left column of Figure~\ref{fig:CCM}. It can be seen that the effect of the
+IP layer in the frequency domain is that the real part starts to deviate from
+the DC value at lower frequencies. The imaginary part decays slower towards
+zero when decreasing the frequency, and it deviates from a linear function of 
+frequency (on log-log scale) at lower frequencies as well. This indicates that
+we need lower frequency content for the Fourier transform the smaller $c$
+becomes.}
+%
+\begin{figure}
+  \centering
+  \includegraphics[width=\fwidth]{11-cole-cole-model}
+  \caption{\added{IP example for different $c$ values, frequency-domain
+  response on the left and time-domain response on the right. The frequencies
+  indicated by the blue vertical bars in the lower-left figure were used for
+  our Fourier transform approach, requiring only the computation of 27
+  frequencies instead of the full 747 frequencies required by the digital
+linear filter method.}}
+  \label{fig:CCM}
+\end{figure}
+%
+
+\added{For the Fourier transform we used the 601-point sine-cosine filter from
+\mbox{\cite{GEO.09.Key}}. This is a very long filter, but it seemed to be by
+far the best filter for low $c$-values. However, designing a new filter
+particularly for IP problems might be a better approach. The selected frequency
+thresholds were chosen at $f_\mr{min}=10^{-4}\,$Hz and $f_\mr{max}=500\,$Hz,
+and within the thresholds we only computed every 6th frequency. Using these
+values only 27 frequencies have to be computed, instead of 747 frequencies
+required by the actual transform. To obtain the step-off response we computed
+the step-on response and subtracted it from the DC value at 1000\,s (without
+the need for any additional frequencies).}
 

### 2021-04-07 ManuscriptTime.tex
Revision II marked changes

@@
     referee,     % uncomment for double lines
-    final,       % uncomment for changes
+    % final,       % uncomment for changes
 ]{gji}
@@
 
-\title[Transforming expensive EM kernels]{Fast Fourier transformation of
-electromagnetic data for computationally expensive kernels}
+\title[Transforming expensive EM kernels]{Fast Fourier transformation
+of electromagnetic data for computationally expensive kernels}
 
@@
 three key points: solver, method and implementation of the Fourier
-transformation, and gridding.
+transform\deleted{ation}, and gridding.
 %
@@
 transform if time-domain data are required \citep[e.g., ][]{GEO.15.Hunziker}.
-CSEM codes for 2D and 3D computations, on the other hand, mostly compute their
-responses directly in the required domain, either frequency or time.
+CSEM codes for 2D and 3D computations, on the other hand,
+\replaced{often}{mostly} compute their responses directly in the required
+domain, either frequency or time.
 
 \added{Electromagnetic methods in geophysics span a wide range of model
-scenarios and acquisition layouts which have all different implications. A
-well-known case is for instance the land or shallow marine case versus the deep
-marine case, where in the former case a dominating airwave has to be accurately
-modelled which can can be completely ignored in the latter case. Recent areas
-of particular interest in time-domain modelling that pose numerical challenges
-are, for instance, simulating the fields through steel-cased wells
-\mbox{\citep{CAG.19.Heagy}} or the effects of induced polarization
-\mbox{\citep{EXP.20.Kang}}. The former is challenging because of the high
-conductivity contrasts requiring very detailed meshing, the latter is
+scenarios and acquisition layouts, each with its own modelling-related
+implications. A well-known example is the aforementioned land or shallow marine
+case versus the deep marine case, where in the former case a dominating airwave
+has to be accurately modelled, whereas it can be completely ignored in the
+latter case. Recent areas of particular interest in time-domain modelling that
+pose numerical challenges are, for instance, simulating the fields through
+steel-cased wells \mbox{\citep{CAG.19.Heagy}} or the effects of induced
+polarization \mbox{\citep{EXP.20.Kang}}. The former is challenging because of
+the high conductivity contrasts requiring very detailed meshing, the latter is
 challenging for time-domain modelling because the models are frequency
@@
 Fort-Frankel method as used in, e.g., \cite{GEO.04.Commer,
-GEO.07.Maao})\replaced{ and}{;} implicit schemes \citep[e.g.,
-][]{GEO.04.Haber}\added{ have a complexity of $O(n^4)$}, \deleted{and} matrix
-exponentials and Lanczos reduction schemes \cite[e.g., ][]{RS.94.Druskin}
-\added{have a complexity of $O(n^4\sqrt{\log n})$}\replaced{, and}{with the}
-Fourier transform of frequency-domain solutions \added{have a complexity $O(n_f
-n^3)$}; \added{$n$ is the number of grid points in each direction, and $n_f$
-the number of frequencies}. Their conclusion was that the Fourier transform
-method can be favourable if the number of required frequencies is small
-relative to the number of grid points in each coordinate. The conditions for
-fast computation of time-domain data with a frequency-domain code are: a
-sufficiently powerful solver, appropriate frequency selection and
+GEO.07.Maao})\replaced{ and}{;} implicit schemes (e.g.,
+\cite{GEO.04.Haber}\added{; \mbox{\cite{GEO.10.Um}}}) \added{ have a complexity
+of $O(n^4)$}, \deleted{and} matrix exponentials and Lanczos reduction schemes
+\cite[e.g., ][]{RS.94.Druskin} \added{have a complexity of $O(n^4\sqrt{\log
+n})$}\replaced{, and}{with the} Fourier transform of frequency-domain solutions
+\added{have a complexity $O(n_f n^3)$}; \added{$n$ is the number of grid points
+in each direction, and $n_f$ the number of frequencies}. Their conclusion was
+that the Fourier transform method can be favourable if the number of required
+frequencies is small relative to the number of grid points in each coordinate.
+The conditions for fast computation of time-domain data with a frequency-domain
+code are: a sufficiently powerful solver, appropriate frequency selection and
 interpolation, and an automated gridding, for which they used the
@@
 
-While we use the finite-integration technique \citep{AEU.77.Weiland} in the
-frequency domain to demonstrate our method, it is important to note that our
-ideas can be used with any solver and is not bound to one or another frequency-
-or time-domain solver. The ideas for the frequency selection can be applied to
-time selection, and time-dependent, adaptive gridding exists as well
-\citep[e.g., ][]{RS.06.Commer}. In the next section, we briefly review the
-methodology as introduced by \cite{GEO.07.Plessix} and \cite{GEO.08.Mulder} and
-highlight their advantages and shortcomings. This is followed by an outline of
+\deleted{While we use the finite-integration technique
+\mbox{\citep{AEU.77.Weiland}} in the frequency domain to demonstrate our
+method, it is important to note that our ideas can be used with any solver and
+is not bound to one or another frequency- or time-domain solver. The ideas for
+the frequency selection can be applied to time selection, and time-dependent,
+adaptive gridding exists as well \mbox{\citep[e.g., ][]{RS.06.Commer}.}} In the
+next section, we briefly review the methodology as introduced by
+\cite{GEO.07.Plessix} and \cite{GEO.08.Mulder} and highlight their advantages
+and shortcomings. This is followed by an outline of \added{the methodology,}
 our changes to the Fourier transform and the adaptive gridding. Finally, we
 demonstrate the efficiency of the approach with some numerical results.
+\added{These compromise of a homogeneous space, a layered model, a model which
+includes dispersive media (induced polarization), and a three-dimensional
+model.}
 
@@
 domain can be desirable, as both domains have advantages and disadvantages. One
-way to achieve this is to implement Maxwell's equations in both domains.
-Another approach is to have Maxwell's equations only implemented in one domain,
-and use Fourier transforms to go to the other. However, this approach can be
-costly, as many frequencies over a wide range are required to go from the
-frequency domain to the time domain, or many times over a wide range for the
-opposite direction. We present a methodology which significantly reduces the
-amount and range of the required frequencies\replaced{ through a combination of
-extrapolation, interpolation, and setting the responses for high frequencies to
-zero. This translates}{, translating} into a significant reduction in
-computation time. \replaced{The number of required frequencies is not the only
-important point for reducing computation time. Another important aspect is the computation grid, as t}{T}he required \deleted{computation} grids
-for low frequencies (in our case in the order of 0.001\,Hz) and high
-frequencies (around 100\,Hz) are hugely different. Low frequencies can be
-computed on a coarser grid, but they require a much larger domain in order to
-avoid boundary effects. High frequencies, on the other hand, require denser
-gridding, but they are much more limited in reach. An adaptive gridding scheme
-is therefore required\replaced{, which}{. We build our approach upon
-\mbox{\cite{GEO.07.Plessix}}, who presented such an adaptive gridding for
-multi-frequency (and multi-source) CSEM modelling.
+way to achieve this is to implement Maxwell's equations in both domains
+\added{as is done, for instance, in \texttt{SimPEG}
+\mbox{\citep{CAG.15.Cockett}} or in \texttt{custEM}
+\mbox{\citep{GJI.21.Rochlitz}}}. Another approach is to have Maxwell's equations
+only implemented in one domain, and use Fourier transforms to go to the other.
+However, this approach can be costly, as many frequencies over a wide range are
+required to go from the frequency domain to the time domain, or many times over
+a wide range for the opposite direction. We present a methodology which
+significantly reduces the amount and range of the required
+frequencies\replaced{ through a combination of extrapolation, interpolation,
+and setting the responses for high frequencies to zero. This translates}{,
+translating} into a significant reduction in computation time. \replaced{The
+number of required frequencies is not the only important point for reducing
+computation time. Another important aspect is the computation grid, as t}{T}he
+required \deleted{computation} grids for low frequencies (in our case in the
+order of 0.001\,Hz) and high frequencies (around 100\,Hz) are hugely different.
+Low frequencies can be computed on a coarser grid, but they require a much
+larger domain in order to avoid boundary effects. High frequencies, on the
+other hand, require denser gridding, but they are much more limited in reach.
+An adaptive gridding scheme is therefore required\replaced{, which}{. We build
+our approach upon \mbox{\cite{GEO.07.Plessix}}, who presented such an adaptive
+gridding for multi-frequency (and multi-source) CSEM modelling.
 
@@
 \deleted{which is} the distance after which the amplitude of the
-electromagnetic field is decayed by $1/e\approx 37\,\%$. The skin depth
-$\delta$ is a function of conductivity and frequency, and for the diffusive
-approximation of Maxwell's equations in an isotropic, homogeneous medium is
-given by \citep[e.g.,][\mbox{\deleted{ equation~1.53}}]{B.SEG.88.Ward}
+electromagnetic field \replaced{has}{is} decayed by $1/e\approx 37\,\%$. The
+skin depth $\delta$ is a function of conductivity and frequency, and for the
+diffusive approximation of Maxwell's equations in an isotropic, homogeneous
+medium is given by \citep[e.g.,][\mbox{\deleted{
+equation~1.53}}]{B.SEG.88.Ward}
 %
@@
 We present improvements to both the adaptive gridding and the transform from
-frequency domain to time domain, which makes time-domain modelling with a
-frequency-domain code viable in comparison with time-domain codes. It is
-important to note that while we show the transform from frequency to time
-domain, the conclusions regarding the Fourier transform can equally well be
-applied to the transform from time-domain to frequency-domain data for a code
-that solves Maxwell's equations in the time domain. \added{Our recommendations
-for speeding up the Fourier transform for expensive kernels are independent of
+frequency domain to time domain \added{using the finite-integration technique
+\mbox{\citep{AEU.77.Weiland}}}, which makes time-domain modelling with a
+frequency-domain code viable in comparison with time-domain codes.
+It is important to note that \added{our ideas can be used with any solver and
+is not bound to one or another frequency- or time-domain solver. The ideas for
+the frequency selection can be applied to time selection, and time-dependent,
+adaptive gridding exists as well \mbox{\citep[e.g., ][]{RS.06.Commer}.}}
+\deleted{while we show the transform from frequency to time domain, the
+conclusions regarding the Fourier transform can equally well be applied to the
+transform from time-domain to frequency-domain data for a code that solves
+Maxwell's equations in the time domain.} \added{Our recommendations for
+speeding up the Fourier transform for expensive kernels are independent of
 spatial complexity, as diffusive EM responses are smooth functions of both
@@
 \added{The requirement for any transform is a robust solver of sufficient
-precision over a wide range of frequencies.} \replaced{We use f}{F}or the
+accuracy over a wide range of frequencies.} \replaced{We use f}{F}or the
 numerical computations \deleted{we use} the open-source (Apache License 2.0)
@@
 
-\added{Note that while we use these two algorithms for the numerical examples
-any solver that provides sufficient accurate and precise frequency-domain
-responses could be used.}
+\added{Note that while we use these two algorithms for the numerical examples,
+any solver that provides sufficiently accurate frequency-domain responses could
+be used.}
 
@@
 for the transform, but to minimize the number of frequencies for which we
-actually have to compute responses. We do this in three zones: Above an upper
-threshold $f_\mr{max}$ we set the frequency-domain response to zero. Below a
-lower threshold $f_\mr{min}$ we extrapolate the frequency-domain response. As
+actually have to compute responses. We do this in three zones: above an upper
+threshold $f_\mr{max}$, we set the frequency-domain response to zero. Below a
+lower threshold $f_\mr{min}$, we extrapolate the frequency-domain response. As
 we know that the imaginary response goes towards zero for zero frequency, we
 can turn it into an interpolation, for which we use PCHIP. And in-between the
-thresholds we use cubic spline interpolation in order to only have to actually
+thresholds, we use cubic spline interpolation in order to only have to actually
 compute a few distinct frequencies per decade ($n_\mr{dec}^f$). Using this
-approach we found that computing the responses for 15--25 frequencies is
+approach, we found that computing the responses for 15--25 frequencies is
 usually good enough to obtain time-domain responses for a large range of
@@
 
-\added{The proposed Fourier transformation requires a robust solver that can
-compute accurate results over a wide range of frequencies (times) to obtain
-time-domain (frequency-domain) responses. In our case the solver uses a
-regular, stretched grid and computes the electric fields in the frequency
-domain. This setup is the target of the following gridding recommendations.
-Whilst they will look differently for other mesh types or a time-domain code,
-some conclusion will still hold.}
+\added{The proposed Fourier transform requires a robust solver that can compute
+accurate results over a wide range of frequencies (times) to obtain time-domain
+(frequency-domain) responses. In our case the solver uses a regular, stretched
+grid and computes the electric fields in the frequency domain. This setup is
+the target of the following gridding recommendations. Whilst they will look
+differently for other mesh types or a time-domain code, some conclusion will
+still hold.}
 
@@
 applies to the horizontal dimensions, not only to the upward
-$z$-direction\added{,} and \replaced{it equally applies}{similarly} for very
+$z$-direction\added{,} and \replaced{it equally applies to}{similarly for} very
 resistive basements, even in deep water. One way to circumvent this difficulty
@@
 with the perfectly matched layer (PML). PML is an absorbing layer to avoid
-scattering from the boundary by letting the field die out to zero, which is
-achieved by introducing a complex factor which causes damping. As
-electromagnetic diffusion fields are damped fields by themselves it suffices to
-stretch the grid.}
+scattering from the boundary by letting the field decay to zero, which is
+achieved by introducing a complex factor that causes damping. As
+electromagnetic diffusion fields are damped fields by themselves, it suffices
+to stretch the grid.}
 
@@
 frequency and time. An interesting test it to see whether this applies as
-well for dispersive media where the model parameters are frequency dependent,
-such as is the case in induced polarization. To test this we use the Cole-Cole
-model \mbox{\citep[CCM, ][]{JCP.41.Cole}} which, written in terms of
-conductivities instead of electric permittivities as by the original authors,
-is given by}
+well for dispersive media where the model parameters depend on frequency, as is
+the case in induced polarization. To test this we use the Cole-Cole model
+\mbox{\citep[CCM, ][]{JCP.41.Cole}} which, written in terms of conductivities
+instead of electric permittivities as by the original authors, is given by}
 %
@@
 
-\added{\mbox{\cite{GJI.17.Commer}}, have shown that time-domain IP modelling
+\added{\mbox{\cite{GJI.17.Commer}} have shown that time-domain IP modelling
 can be difficult, particularly for low $c$-values. Modelling
-frequency-dependent models in the frequency domain is for any given frequency
-naturally no different than modelling a non frequency-dependent model. However,
-the interesting point is to see how the Fourier transform behaves to obtain
-time-domain data from frequency-domain responses. As an example we modelled the
-response for a land case, air above a halfspace of 1\,S/m. Within the halfspace
-is a dispersive, 100\,m thick layer at a depth of 300\,m, with
-$\sigma_0=1.0\,$S/m and $\sigma_\infty=1.25\,$S/m; $\tau=1$, and we vary
-$c$ from 0 (no IP) to 1.0, 0.75, 0.5, and 0.25. The x-directed step-off source
-and the receiver at an inline offset of 900\,m are place at the surface. The
-real and imaginary frequency-domain responses for this model are shown in the
-left column of Figure~\ref{fig:CCM}. It can be seen that the effect of the
-IP layer in the frequency domain is that the real part starts to deviate from
-the DC value at lower frequencies. The imaginary part decays slower towards
-zero when decreasing the frequency, and it deviates from a linear function of 
-frequency (on log-log scale) at lower frequencies as well. This indicates that
-we need lower frequency content for the Fourier transform the smaller $c$
-becomes.}
+frequency-dependent models in the frequency domain is, for any given frequency,
+not different from modelling a non frequency-dependent model. However, the
+interesting point is to see how the Fourier transform behaves to obtain
+time-domain data from frequency-domain responses. As an example, we modelled
+the response for a land case with air above a halfspace of 1\,S/m. Within the
+halfspace is a dispersive, 100\,m thick layer at a depth of 300\,m, with
+$\sigma_0=1.0\,$S/m and $\sigma_\infty=1.25\,$S/m; $\tau=1$, and we vary $c$
+from 0 (no IP) to 1.0, 0.75, 0.5, and 0.25. The x-directed step-off source and
+the receiver at an inline offset of 900\,m are place at the surface. The real
+and imaginary frequency-domain responses for this model are shown in the left
+column of Figure~\ref{fig:CCM}. It can be seen that the effect of the IP layer
+in the frequency domain is that the real part starts to deviate from the DC
+value at lower frequencies. The imaginary part decays slower towards zero when
+decreasing the frequency, and it deviates from a power-law function at lower
+frequencies as well. This indicates that we need lower frequency content for
+the Fourier transform the smaller $c$ becomes.}
 %
@@
 
-\added{For the Fourier transform we used the 601-point sine-cosine filter from
+\added{For the Fourier transform, we used the 601-point sine-cosine filter from
 \mbox{\cite{GEO.09.Key}}. This is a very long filter, but it seemed to be by
@@
 and within the thresholds we only computed every 6th frequency. Using these
-values only 27 frequencies have to be computed, instead of 747 frequencies
+values only 27 frequencies have to be computed, instead of the 747 frequencies
 required by the actual transform. To obtain the step-off response we computed
@@
 
-The final example consists of a resistive, three-dimensional block embedded in
-the lower of two halfspaces, as depicted in Figure~\ref{fig:3d-model}. The
-target has resistivity $\rho_\mr{tg}=100\,\ohmm$, the upper halfspace
-corresponds to seawater with $\rho_\mr{sea}=0.3\,\ohmm$, and the lower
-halfspace is the background with $\rho_\mr{bg}=1\,\ohmm$.
+The \replaced{last}{final} example consists of a resistive, three-dimensional
+block embedded in the lower of two halfspaces, as depicted in
+Figure~\ref{fig:3d-model}. The target has resistivity
+$\rho_\mr{tg}=100\,\ohmm$, the upper halfspace corresponds to seawater with
+$\rho_\mr{sea}=0.3\,\ohmm$, and the lower halfspace is the background with
+$\rho_\mr{bg}=1\,\ohmm$.
 %
@@
 
-For the comparison we use the open-source code \simpeg \citep{CAG.15.Cockett},
-which is a framework for modelling and inversion of geophysical data such as
-gravity, magnetics, and CSEM. It has Maxwell's equations implemented in both
-the frequency and time domain. As such we can compare our result computed in
-the frequency domain followed by a Fourier transform to a result computed
-directly in the time domain. A principal difference between \simpeg and \emg3d
-is that the former has various direct solvers implemented, whereas the latter
-is an iterative multigrid solver. The 3D model is therefore a rather small
-example in order to be able to run it on our test machine, as the memory
-requirement by the direct solver would otherwise be too high. There are not
-many options out there of open-source time-domain 3D CSEM codes, \simpeg being
-the one we found to be suitable. A step-off response was chosen as this is the
-response currently implemented in it.
+For the comparison\added{,} we use the open-source code \simpeg
+\citep{CAG.15.Cockett}, which is a framework for modelling and inversion of
+geophysical data such as gravity, magnetics, and CSEM. It has Maxwell's
+equations implemented in both the frequency and time domain. As such we can
+compare our result computed in the frequency domain followed by a Fourier
+transform to a result computed directly in the time domain. A principal
+difference between \simpeg and \emg3d is that the former has various direct
+solvers implemented, whereas the latter is an iterative multigrid solver. The
+3D model is therefore a rather small example in order to be able to run it on
+our test machine, as the memory requirement by the direct solver would
+otherwise be too high. There are not many options out there of open-source
+time-domain 3D CSEM codes, \simpeg being the one we found to be suitable. A
+step-off response was chosen as this is the response currently implemented in
+it.
 
@@
 other domain followed by a transformation a viable alternative. The methodology
-is relatively simple to implement in any code.
+is relatively simple to implement \added{and has therefore the potential to
+expand the capability} \replaced{of}{in} any \added{existing} code \added{to
+an additional dimension}.
 
@@
 \begin{acknowledgments} % % % % % % % % % % % % % % % % % % % % % % % % % % % %
+\added{We would like to thank the editor Ute Weckmann and assistant editor
+Louise Alexander as well as the reviewers Michael Commer and Chester Weiss for
+many helpful comments, which improved this manuscript considerably.}
 This research was conducted within the Gitaro.JIM project funded through

### 2021-04-07 notebooks/freqselect.py
Revision II - changes (#3)

@@ -156,8 +156,8 @@ class InteractiveFrequency(emg3d.utils.Fourier):
         ax3.set_ylabel('Rel. Error (%)')
         ax3.set_xlabel('Frequency (Hz)')
         ax4.set_xlabel('Time (s)')
-        ax3.axhline(1, c='k')
-        ax4.axhline(1, c='k')
+        ax3.axhline(1, c='0.4')
+        ax4.axhline(1, c='0.4')
 
         # Add instances
         self.fig = fig
@@ -363,9 +363,9 @@ class InteractiveFrequency(emg3d.utils.Fourier):
         t_int = self.freq2time(f_calc, self.model['rec'][0])
 
         # Calculate the errors.
-        f_error = np.clip(100*abs((self.reim(f_int)-self.reim(f_req)) /
-                                  self.reim(f_req)), 0.01, 100)
-        t_error = np.clip(100*abs((t_int-self.t_base)/self.t_base), 0.01, 100)
+        f_error = 100*abs((self.reim(f_int)-self.reim(f_req)) /
+                          self.reim(f_req))
+        t_error = 100*abs((t_int-self.t_base)/self.t_base)
 
         # Clear existing handles
         self.clear_handle(['f_int', 't_int', 'f_inti', 'f_inte', 't_inte'])
@@ -375,11 +375,11 @@ class InteractiveFrequency(emg3d.utils.Fourier):
                 self.freq_req, self.reim(f_int), 'k.', ms=4)
         self.h_f_int, = self.axs[0].plot(
                 self.freq_calc, self.reim(f_calc), 'C0.', ms=8)
-        self.h_f_inte, = self.axs[2].plot(self.freq_req, f_error, 'k.')
+        self.h_f_inte, = self.axs[2].plot(self.freq_req, f_error, 'k')
 
         # Plot time-domain result
         self.h_t_int, = self.axs[1].plot(self.time, t_int, 'k--')
-        self.h_t_inte, = self.axs[3].plot(self.time, t_error, 'k.')
+        self.h_t_inte, = self.axs[3].plot(self.time, t_error, 'k')
 
         # Update legend
         self.print_legend()

### 2021-04-07 ManuscriptTime.tex
Revision II - clean

@@
 \documentclass[extra, camera,%
-    onecolumn,   % comment for two columns
-    referee,     % uncomment for double lines
-    % final,       % uncomment for changes
+    % onecolumn,   % comment for two columns
+    % referee,     % uncomment for double lines
+    final,       % uncomment for changes
 ]{gji}
@@
 The speed of such modelling that is obtained through a transform is defined by
-three key points: solver, method and implementation of the Fourier
-transform\deleted{ation}, and gridding.
+three key points: solver, method and implementation of the Fourier transform,
+and gridding.
 %
@@
 transform if time-domain data are required \citep[e.g., ][]{GEO.15.Hunziker}.
-CSEM codes for 2D and 3D computations, on the other hand,
-\replaced{often}{mostly} compute their responses directly in the required
-domain, either frequency or time.
-
-\added{Electromagnetic methods in geophysics span a wide range of model
-scenarios and acquisition layouts, each with its own modelling-related
-implications. A well-known example is the aforementioned land or shallow marine
-case versus the deep marine case, where in the former case a dominating airwave
-has to be accurately modelled, whereas it can be completely ignored in the
-latter case. Recent areas of particular interest in time-domain modelling that
-pose numerical challenges are, for instance, simulating the fields through
-steel-cased wells \mbox{\citep{CAG.19.Heagy}} or the effects of induced
-polarization \mbox{\citep{EXP.20.Kang}}. The former is challenging because of
-the high conductivity contrasts requiring very detailed meshing, the latter is
-challenging for time-domain modelling because the models are frequency
-dependent.}
-
-\deleted{There is a wealth of 3D electromagnetic codes, with a first boom in the
-1990's; \mbox{\cite{B.SEG.99.Oristaglio}} provide a good review of the state of
-art at that time. A second wave occurred during the CSEM boom in the
-hydrocarbon industry in the early 21th century, see, e.g.,
-\mbox{\cite{SG.05.Avdeev}} and \mbox{\cite{SG.10.Borner}} for comprehensive
-overviews. Recent publications after these review articles span the width of
-numerical methods from integral equation codes \mbox{\citep{MGS.17.Kruglyakov}}
-and differential equation methods such as finite differences
-\mbox{\citep{GEO.10.Mittet, GEO.11.Zaslavsky, CAG.13.Sommer}}, finite elements
-\mbox{\citep{GJI.13.Grayver}}, and finite volumes
-\mbox{\citep{GEO.14.Jahandari}}. Besides the different methods there are also
-many different types of discretisation. Commonly used, besides rectilinear Yee
-grids \mbox{\citep{IEEE.66.Yee}} as used by our code, are Lebedev grids
-\mbox{\citep{CMMP.64.Lebedev}} but also unstructured grids using, for instance,
-tetrahedra \mbox{\citep{CAG.17.Cai}}. The computational requirement to model
-CSEM data also depends heavily on the technique used to solve the system. There
-are direct solvers \mbox{\citep{GEO.15.Grayver}}, indirect solvers
-\mbox{\citep{GJI.15.Jaysaval}}, and more recently also hybrid solvers
-\mbox{\citep{GEO.18.Liu}}.}
+CSEM codes for 2D and 3D computations, on the other hand, often compute their
+responses directly in the required domain, either frequency or time.
+
+Electromagnetic methods in geophysics span a wide range of model scenarios and
+acquisition layouts, each with its own modelling-related implications. A
+well-known example is the aforementioned land or shallow marine case versus the
+deep marine case, where in the former case a dominating airwave has to be
+accurately modelled, whereas it can be completely ignored in the latter case.
+Recent areas of particular interest in time-domain modelling that pose
+numerical challenges are, for instance, simulating the fields through
+steel-cased wells \citep{CAG.19.Heagy} or the effects of induced polarization
+\citep{EXP.20.Kang}. The former is challenging because of the high conductivity
+contrasts requiring very detailed meshing, the latter is challenging for
+time-domain modelling because the models are frequency dependent.
 
@@
 data with a frequency-domain code, as shown by \cite{GEO.08.Mulder}. They
-compared the computational complexity \added{of different time-domain methods
-with the computational complexity of frequency-domain computation followed by a
-Fourier transform:} \deleted{of} explicit time-stepping schemes (Du
-Fort-Frankel method as used in, e.g., \cite{GEO.04.Commer,
-GEO.07.Maao})\replaced{ and}{;} implicit schemes (e.g.,
-\cite{GEO.04.Haber}\added{; \mbox{\cite{GEO.10.Um}}}) \added{ have a complexity
-of $O(n^4)$}, \deleted{and} matrix exponentials and Lanczos reduction schemes
-\cite[e.g., ][]{RS.94.Druskin} \added{have a complexity of $O(n^4\sqrt{\log
-n})$}\replaced{, and}{with the} Fourier transform of frequency-domain solutions
-\added{have a complexity $O(n_f n^3)$}; \added{$n$ is the number of grid points
-in each direction, and $n_f$ the number of frequencies}. Their conclusion was
-that the Fourier transform method can be favourable if the number of required
+compared the computational complexity of different time-domain methods with the
+computational complexity of frequency-domain computation followed by a Fourier
+transform: explicit time-stepping schemes (Du Fort-Frankel method as used in,
+e.g., \cite{GEO.04.Commer, GEO.07.Maao}) and implicit schemes \citep[e.g.,
+][]{GEO.04.Haber,GEO.10.Um} have a complexity of $O(n^4)$, matrix exponentials
+and Lanczos reduction schemes \cite[e.g., ][]{RS.94.Druskin} have a complexity
+of $O(n^4\sqrt{\log n})$, and Fourier transform of frequency-domain solutions
+have a complexity $O(n_f n^3)$; $n$ is the number of grid points in each
+direction, and $n_f$ the number of frequencies. Their conclusion was that the
+Fourier transform method can be favourable if the number of required
 frequencies is small relative to the number of grid points in each coordinate.
@@
 these results but improve the run time from hours to minutes. The main reasons
-for this significant speed-up are \added{the further reduction of required
-frequencies by introducing lower and upper thresholds of numerically important
-frequencies,} an adaptive, frequency-dependent gridding scheme that minimizes
-the required cells in each dimension, and a logarithmic Fourier transform such
-as digital linear filters \citep[DLF, ][]{GP.71.Ghosh} or the logarithmic fast
-Fourier transform \citep[FFTLog, ][]{RAS.00.Hamilton} to go from the frequency
-to the time domain. The latter makes it also possible to only use the imaginary
-part of the frequency-domain response, which has advantages when it comes to
+for this significant speed-up are the further reduction of required frequencies
+by introducing lower and upper thresholds of numerically important frequencies,
+an adaptive, frequency-dependent gridding scheme that minimizes the required
+cells in each dimension, and a logarithmic Fourier transform such as digital
+linear filters \citep[DLF, ][]{GP.71.Ghosh} or the logarithmic fast Fourier
+transform \citep[FFTLog, ][]{RAS.00.Hamilton} to go from the frequency to the
+time domain. The latter makes it also possible to only use the imaginary part
+of the frequency-domain response, which has advantages when it comes to
 interpolation.
 
-\deleted{While we use the finite-integration technique
-\mbox{\citep{AEU.77.Weiland}} in the frequency domain to demonstrate our
-method, it is important to note that our ideas can be used with any solver and
-is not bound to one or another frequency- or time-domain solver. The ideas for
-the frequency selection can be applied to time selection, and time-dependent,
-adaptive gridding exists as well \mbox{\citep[e.g., ][]{RS.06.Commer}.}} In the
-next section, we briefly review the methodology as introduced by
+In the next section, we briefly review the methodology as introduced by
 \cite{GEO.07.Plessix} and \cite{GEO.08.Mulder} and highlight their advantages
-and shortcomings. This is followed by an outline of \added{the methodology,}
-our changes to the Fourier transform and the adaptive gridding. Finally, we
-demonstrate the efficiency of the approach with some numerical results.
-\added{These compromise of a homogeneous space, a layered model, a model which
-includes dispersive media (induced polarization), and a three-dimensional
-model.}
+and shortcomings. This is followed by an outline of the methodology, our
+changes to the Fourier transform and the adaptive gridding. Finally, we
+demonstrate the efficiency of the approach with some numerical results. These
+compromise of a homogeneous space, a layered model, a model which includes
+dispersive media (induced polarization), and a three-dimensional model.
 
@@
 domain can be desirable, as both domains have advantages and disadvantages. One
-way to achieve this is to implement Maxwell's equations in both domains
-\added{as is done, for instance, in \texttt{SimPEG}
-\mbox{\citep{CAG.15.Cockett}} or in \texttt{custEM}
-\mbox{\citep{GJI.21.Rochlitz}}}. Another approach is to have Maxwell's equations
-only implemented in one domain, and use Fourier transforms to go to the other.
-However, this approach can be costly, as many frequencies over a wide range are
-required to go from the frequency domain to the time domain, or many times over
-a wide range for the opposite direction. We present a methodology which
-significantly reduces the amount and range of the required
-frequencies\replaced{ through a combination of extrapolation, interpolation,
-and setting the responses for high frequencies to zero. This translates}{,
-translating} into a significant reduction in computation time. \replaced{The
-number of required frequencies is not the only important point for reducing
-computation time. Another important aspect is the computation grid, as t}{T}he
-required \deleted{computation} grids for low frequencies (in our case in the
-order of 0.001\,Hz) and high frequencies (around 100\,Hz) are hugely different.
-Low frequencies can be computed on a coarser grid, but they require a much
-larger domain in order to avoid boundary effects. High frequencies, on the
-other hand, require denser gridding, but they are much more limited in reach.
-An adaptive gridding scheme is therefore required\replaced{, which}{. We build
-our approach upon \mbox{\cite{GEO.07.Plessix}}, who presented such an adaptive
-gridding for multi-frequency (and multi-source) CSEM modelling.
-
-An adaptive grid in the frequency domain} is naturally based on the skin depth,
-\deleted{which is} the distance after which the amplitude of the
-electromagnetic field \replaced{has}{is} decayed by $1/e\approx 37\,\%$. The
-skin depth $\delta$ is a function of conductivity and frequency, and for the
-diffusive approximation of Maxwell's equations in an isotropic, homogeneous
-medium is given by \citep[e.g.,][\mbox{\deleted{
-equation~1.53}}]{B.SEG.88.Ward}
+way to achieve this is to implement Maxwell's equations in both domains as is
+done, for instance, in \texttt{SimPEG} \citep{CAG.15.Cockett} or in
+\texttt{custEM} \citep{GJI.21.Rochlitz}. Another approach is to have Maxwell's
+equations only implemented in one domain, and use Fourier transforms to go to
+the other. However, this approach can be costly, as many frequencies over a
+wide range are required to go from the frequency domain to the time domain, or
+many times over a wide range for the opposite direction. We present a
+methodology which significantly reduces the amount and range of the required
+frequencies through a combination of extrapolation, interpolation, and setting
+the responses for high frequencies to zero. This translates into a significant
+reduction in computation time. The number of required frequencies is not the
+only important point for reducing computation time. Another important aspect is
+the computation grid, as the required grids for low frequencies (in our case in
+the order of 0.001\,Hz) and high frequencies (around 100\,Hz) are hugely
+different. Low frequencies can be computed on a coarser grid, but they require
+a much larger domain in order to avoid boundary effects. High frequencies, on
+the other hand, require denser gridding, but they are much more limited in
+reach. An adaptive gridding scheme is therefore required, which is naturally
+based on the skin depth, the distance after which the amplitude of the
+electromagnetic field has decayed by $1/e\approx 37\,\%$. The skin depth
+$\delta$ is a function of conductivity and frequency, and for the diffusive
+approximation of Maxwell's equations in an isotropic, homogeneous medium is
+given by \citep[e.g.,][]{B.SEG.88.Ward}
 %
@@
 
-\added{We build our approach upon} \cite{GEO.07.Plessix}\added{, who presented
-such an adaptive gridding for multi-frequency (and multi-source) CSEM
-modelling. They} define the minimum cell width $\Delta_\mathrm{min}$ as
-a fraction of the minimum skin depth $\delta_\mathrm{min}$, where
-$\delta_\mathrm{min}$ should be 2--3 times bigger than $\Delta_\mathrm{min}$.
-The cells have to be smallest around the source; in the marine case, the
-minimum skin depth is therefore defined by the conductivity of seawater.
-However, this \replaced{can yield}{yields} quite large cells for low
-frequencies, so special care has to be taken around the source by defining a
-maximum allowed $\Delta_\mathrm{min}$. The grid dimension, on the other hand,
-is defined as a function of \deleted{the} skin depth for the average
-conductivity of the background, $\delta_\mathrm{ave}$. They use four times
+We build our approach upon \cite{GEO.07.Plessix}, who presented such an
+adaptive gridding for multi-frequency (and multi-source) CSEM modelling. They
+define the minimum cell width $\Delta_\mathrm{min}$ as a fraction of the
+minimum skin depth $\delta_\mathrm{min}$, where $\delta_\mathrm{min}$ should be
+2--3 times bigger than $\Delta_\mathrm{min}$. The cells have to be smallest
+around the source; in the marine case, the minimum skin depth is therefore
+defined by the conductivity of seawater. However, this can yield quite large
+cells for low frequencies, so special care has to be taken around the source by
+defining a maximum allowed $\Delta_\mathrm{min}$. The grid dimension, on the
+other hand, is defined as a function of skin depth for the average conductivity
+of the background, $\delta_\mathrm{ave}$. They use four times
 $\delta_\mathrm{ave}$ for the $x$-, $y$-, and downward $z$-directions, and a
@@
 We present improvements to both the adaptive gridding and the transform from
-frequency domain to time domain \added{using the finite-integration technique
-\mbox{\citep{AEU.77.Weiland}}}, which makes time-domain modelling with a
-frequency-domain code viable in comparison with time-domain codes.
-It is important to note that \added{our ideas can be used with any solver and
-is not bound to one or another frequency- or time-domain solver. The ideas for
-the frequency selection can be applied to time selection, and time-dependent,
-adaptive gridding exists as well \mbox{\citep[e.g., ][]{RS.06.Commer}.}}
-\deleted{while we show the transform from frequency to time domain, the
-conclusions regarding the Fourier transform can equally well be applied to the
-transform from time-domain to frequency-domain data for a code that solves
-Maxwell's equations in the time domain.} \added{Our recommendations for
+frequency domain to time domain using the finite-integration technique
+\citep{AEU.77.Weiland}, which makes time-domain modelling with a
+frequency-domain code viable in comparison with time-domain codes. It is
+important to note that our ideas can be used with any solver and is not bound
+to one or another frequency- or time-domain solver. The ideas for the frequency
+selection can be applied to time selection, and time-dependent, adaptive
+gridding exists as well \citep[e.g., ][]{RS.06.Commer}. Our recommendations for
 speeding up the Fourier transform for expensive kernels are independent of
@@
 frequency and time. However, the intrinsic requirement is an accurate solver
-over a sufficient wide range of frequencies or times.} \deleted{A variation of
-our method was implemented, for instance, by \custem
-\mbox{\citep{GEO.19.Rochlitz}} to obtain frequency-domain data from time-domain
-computations (Rochlitz, private communication, 2020).}
+over a sufficient wide range of frequencies or times.
 
@@
 
-\added{The requirement for any transform is a robust solver of sufficient
-accuracy over a wide range of frequencies.} \replaced{We use f}{F}or the
-numerical computations \deleted{we use} the open-source (Apache License 2.0)
-code\added{s \empymod} \added{\mbox{\citep{GEO.17.Werthmuller}}} \added{and}
-\emg3d \citep{JOSS.19.Werthmuller}\replaced{.}{,} \added{The former computes
-semi-analytical responses of layered models, the latter is} a \added{3D}
-multigrid solver based on \cite{GP.06.Mulder}, which can be used as a
-preconditioner for Krylov subspace solvers or as a solver on its own. The
-multigrid approach works fine for the diffusive approximation of Maxwell's
-equations, which assumes that $\omega\varepsilon \ll \sigma$, where
-$\varepsilon$ is electric permittivity (F/m). The remaining system to solve in
-the frequency domain is then given by the second-order differential equation of
-the electric field,
+The requirement for any transform is a robust solver of sufficient accuracy
+over a wide range of frequencies. We use for the numerical computations the
+open-source (Apache License 2.0) codes \empymod \citep{GEO.17.Werthmuller} and
+\emg3d \citep{JOSS.19.Werthmuller}. The former computes semi-analytical
+responses of layered models, the latter is a 3D multigrid solver based on
+\cite{GP.06.Mulder}, which can be used as a preconditioner for Krylov subspace
+solvers or as a solver on its own. The multigrid approach works fine for the
+diffusive approximation of Maxwell's equations, which assumes that
+$\omega\varepsilon \ll \sigma$, where $\varepsilon$ is electric permittivity
+(F/m). The remaining system to solve in the frequency domain is then given by
+the second-order differential equation of the electric field,
 %
@@
 makes it feasible to run even big models on standard computers, without the
-need for big clusters. All \deleted{the} examples in this article are run on a
-laptop with an i7-6600U\,CPU\,@\,2.6\,GHz (x4) and 16\,GB of memory, using
-Ubuntu 20.04 and Python 3.8. \replaced{They}{All examples} were run using a
-single thread, but parallelization over frequencies is straight forward, as the
-computations are independent. \deleted{For the semi-analytical computations of
-layered models we use the open-source (Apache License 2.0) code \empymod
-\mbox{\citep{GEO.17.Werthmuller}}.}
-
-\added{Note that while we use these two algorithms for the numerical examples,
-any solver that provides sufficiently accurate frequency-domain responses could
-be used.}
+need for big clusters. All examples in this article are run on a laptop with an
+i7-6600U\,CPU\,@\,2.6\,GHz (x4) and 16\,GB of memory, using Ubuntu 20.04 and
+Python 3.8. They were run using a single thread, but parallelization over
+frequencies is straight forward, as the computations are independent.
+
+Note that while we use these two algorithms for the numerical examples, any
+solver that provides sufficiently accurate frequency-domain responses could be
+used.
 
@@
 frequencies. The fewer frequencies required, the quicker we obtain the
-time-domain result. \added{This is true in general but applies in particular
-to expensive kernels such as is the case in 3D modelling, where the computation
-of the response of a single frequency takes much longer than the transformation
-itself.} The same applies for the reverse operation, the fewer times required,
+time-domain result. This is true in general but applies in particular to
+expensive kernels such as is the case in 3D modelling, where the computation of
+the response of a single frequency takes much longer than the transformation
+itself. The same applies for the reverse operation, the fewer times required,
 the quicker we obtain frequency-domain results. Even though we limit our
@@
 the required frequencies change with offset and an adaptive frequency selection
-is therefore often tailored to a single offset\added{ or a limited range of
-offsets}. Also, it allows for straightforward parallelization over frequencies,
-which is not completely possible with an adaptive scheme.
+is therefore often tailored to a single offset or a limited range of offsets.
+Also, it allows for straightforward parallelization over frequencies, which is
+not completely possible with an adaptive scheme.
 
@@
 
-\replaced{Any Fourier transform needs a certain range of frequencies to obtain
+Any Fourier transform needs a certain range of frequencies to obtain
 time-domain results. Our approach is not to minimize the required frequencies
@@
 offsets; even though the actual transform might need up to hundreds of
-frequencies.}{There are three key-parameters that have to be defined: Minimum
-and maximum frequency ($f_\mr{min}, f_\mr{max}$) and the number of frequencies
-per decade ($n_\mr{dec}^f$).}
-
-\replaced{The required frequencies depend naturally on the conductivities and
-on the chosen acquisition type and layout. A t}{T}rial-and-error approach with a
-3D code is very time-intensive. However, a simplified, layered model for the
-required survey setup and a fast 1D modeller makes it possible to estimate
-these parameters easily. \replaced{An example of this is shown in
-Figure~\ref{fig:FFTLog-log}.}{To this end, we created a graphical, interactive
-tool using \empymod, as shown in Figure~\ref{fig:FFTLog-log}.}
+frequencies.
+
+The required frequencies depend naturally on the conductivities and on the
+chosen acquisition type and layout. A trial-and-error approach with a 3D code
+is very time-intensive. However, a simplified, layered model for the required
+survey setup and a fast 1D modeller makes it possible to estimate these
+parameters easily. An example of this is shown in Figure~\ref{fig:FFTLog-log}.
 %
@@
   \includegraphics[width=\cwidth]{01-FFTLog-log}
-  \caption{\replaced{Example}{Interactive} frequency selection for a
-    \replaced{simple}{user-provided} layered model (the shown model parameters
-    are described in the text). This example shows the impulse response at an
-    offset of 5\,km, for which it uses FFTLog from 0.001\,Hz to 10\,Hz with
-    five frequencies per decade. \deleted{Note that the error is clipped for
-    values smaller than 0.01\,\% and bigger than 100\,\%.}}
+  \caption{Example frequency selection for a simple layered model (the shown
+    model parameters are described in the text). This example shows the impulse
+    response at an offset of 5\,km, for which it uses FFTLog from 0.001\,Hz to
+    10\,Hz with five frequencies per decade.}
   \label{fig:FFTLog-log}
@@
 %
-\replaced{It shows the responses for}{The shown example model is} a
-marine scenario with 1\,km water depth of resistivity $\rho=0.3\,\ohmm$ ($\rho
-= \sigma^{-1}$) and a 100\,m thick target of 100\,\ohmm at 1\,km below the
-seafloor in a background of 1\,\ohmm. The source is 50\,m above the seafloor,
-the receiver is on the seafloor, and the response is the inline $x$-directed
-$E$-field. The left subplot shows the imaginary part in the frequency domain
-and the right plot the corresponding impulse response in the time domain. The
-red lines are the semi-analytical responses. The blue circles indicate the
-actually computed responses and the black dots the frequencies which are
-interpolated or set to zero. The resulting time-domain response has a relative
-error of less than 1\,\% everywhere except for very early times.
-
-Figure \ref{fig:FFTLog-lin} shows exactly the same on a
-\replaced{linear}{logarithmic} scale \replaced{for the time and amplitude
-axes}{, without the interactive GUI-related elements}. It can be seen that with
-the chosen\added{, computed} frequencies \added{(blue dots)} the time-domain
-response starts to divert above about 100\,s and below 0.3\,s. It also shows
-the oscillating high-frequency part, which is hard to interpolate and we
-therefore set it to zero. Figure~\ref{fig:DLF-lin} is the same as Figure
-\ref{fig:FFTLog-lin}, but transformed with the DLF method applying the 81-point
-sine-cosine filter from \cite{GEO.09.Key}. The same frequencies were computed
-as in the FFTLog case and the missing ones interpolated. The error of the
-corresponding time-domain response is comparable, so either FFTLog or DLF can
-be used.
+It shows the responses for a marine scenario with 1\,km water depth of
+resistivity $\rho=0.3\,\ohmm$ ($\rho = \sigma^{-1}$) and a 100\,m thick target
+of 100\,\ohmm at 1\,km below the seafloor in a background of 1\,\ohmm. The
+source is 50\,m above the seafloor, the receiver is on the seafloor, and the
+response is the inline $x$-directed $E$-field. The left subplot shows the
+imaginary part in the frequency domain and the right plot the corresponding
+impulse response in the time domain. The red lines are the semi-analytical
+responses. The blue circles indicate the actually computed responses and the
+black dots the frequencies which are interpolated or set to zero. The resulting
+time-domain response has a relative error of less than 1\,\% everywhere except
+for very early times.
+
+Figure \ref{fig:FFTLog-lin} shows exactly the same on a linear scale for the
+time and amplitude axes. It can be seen that with the chosen, computed
+frequencies (blue dots) the time-domain response starts to divert above about
+100\,s and below 0.3\,s. It also shows the oscillating high-frequency part,
+which is hard to interpolate and we therefore set it to zero.
+Figure~\ref{fig:DLF-lin} is the same as Figure \ref{fig:FFTLog-lin}, but
+transformed with the DLF method applying the 81-point sine-cosine filter from
+\cite{GEO.09.Key}. The same frequencies were computed as in the FFTLog case and
+the missing ones interpolated. The error of the corresponding time-domain
+response is comparable, so either FFTLog or DLF can be used.
 %
@@
   \label{fig:DLF-lin}}\\
-  \caption{(a) Same as Figure \ref{fig:FFTLog-log}, but on a logarithmic
-  scale. (b) Using DLF instead of FFTLog, but computing the same frequency
-  range as for the FFTLog. The resulting time-domain response has comparable
-  accuracy\deleted{ (compare the error to the error in Figure
-    \ref{fig:FFTLog-log})}.}
+  \caption{(a) Same as Figure \ref{fig:FFTLog-log}, but on a logarithmic scale.
+  (b) Using DLF instead of FFTLog, but computing the same frequency range as
+  for the FFTLog. The resulting time-domain response has comparable accuracy.}
   \label{fig:lin-lin}
@@
 required for the shown offset of 5\,km, a few of the lower and higher
-frequencies could be left out. With this frequency-selection\added{, however,}
-we can model a wide range of offsets. This is shown in Figure
-\ref{fig:multi-offset}, where the same kernels were used to yield the responses
-at offsets $r=1.5,3,6$, and 12\,km. We only need to compute 50\,\% of the
-required frequencies in the case of the FFTLog, and only 14\,\% in the case of
-the DLF.
+frequencies could be left out. With this frequency-selection, however, we can
+model a wide range of offsets. This is shown in Figure \ref{fig:multi-offset},
+where the same kernels were used to yield the responses at offsets $r=1.5,3,6$,
+and 12\,km. We only need to compute 50\,\% of the required frequencies in the
+case of the FFTLog, and only 14\,\% in the case of the DLF.
 %
@@
 better to use a cubic spline. As can be seen by comparing Figure
-\ref{fig:FFTLog-lin} with Figure \ref{fig:DLF-lin}, using the FFTLog or
-the DLF with the same actually computed frequencies results in very similar
-responses.
+\ref{fig:FFTLog-lin} with Figure \ref{fig:DLF-lin}, using the FFTLog or the DLF
+with the same actually computed frequencies results in very similar responses.
 
@@
 
-\added{The proposed Fourier transform requires a robust solver that can compute
+The proposed Fourier transform requires a robust solver that can compute
 accurate results over a wide range of frequencies (times) to obtain time-domain
@@
 differently for other mesh types or a time-domain code, some conclusion will
-still hold.}
+still hold.
 
@@
 should be per skin depth. The actual computational domain $D_\mr{c}$ is usually
-much bigger than $D_\mr{s}$ in order to avoid artefacts from the
-\deleted{perfectly electrically conducting} boundary condition. It can also
-have a much higher stretching factor $\alpha_\mr{c}$. In the presented examples
-we have chosen $D_\mr{c}$ such that the distance for the signal diffusing from
-the source to the boundary and back to the receiver closest to the boundary is
-at least two wavelengths, after which the amplitude of the signal is reduced to
-a millionth of its initial strength. The wavelength $\lambda$ to compute
-$D_\mr{c}$ is given by
+much bigger than $D_\mr{s}$ in order to avoid artefacts from the boundary
+condition. It can also have a much higher stretching factor $\alpha_\mr{c}$. In
+the presented examples we have chosen $D_\mr{c}$ such that the distance for the
+signal diffusing from the source to the boundary and back to the receiver
+closest to the boundary is at least two wavelengths, after which the amplitude
+of the signal is reduced to a millionth of its initial strength. The wavelength
+$\lambda$ to compute $D_\mr{c}$ is given by
 %
@@
 the source to the boundary; this distance is by default set to 100\,km, but
-this can be reduced in the marine case with increasing water depth.
-\deleted{Note that for shallow marine and land cases t}\added{T}his also
-applies to the horizontal dimensions, not only to the upward
-$z$-direction\added{,} and \replaced{it equally applies to}{similarly for} very
-resistive basements, even in deep water. One way to circumvent this difficulty
-is the use of a primary-secondary formulation, where the primary field,
-including the air wave, is computed with a semi-analytical code for layered
-media. We do not consider this approach here. \added{Note that grid stretching
-for complex-valued diffusion fields is essentially what is done for wave fields
-with the perfectly matched layer (PML). PML is an absorbing layer to avoid
-scattering from the boundary by letting the field decay to zero, which is
-achieved by introducing a complex factor that causes damping. As
+this can be reduced in the marine case with increasing water depth. This also
+applies to the horizontal dimensions, not only to the upward $z$-direction, and
+it equally applies to very resistive basements, even in deep water. One way to
+circumvent this difficulty is the use of a primary-secondary formulation, where
+the primary field, including the air wave, is computed with a semi-analytical
+code for layered media. We do not consider this approach here. Note that grid
+stretching for complex-valued diffusion fields is essentially what is done for
+wave fields with the perfectly matched layer (PML). PML is an absorbing layer
+to avoid scattering from the boundary by letting the field decay to zero, which
+is achieved by introducing a complex factor that causes damping. As
 electromagnetic diffusion fields are damped fields by themselves, it suffices
-to stretch the grid.}
+to stretch the grid.
 
@@
     actually computed responses, the black dots are the interpolated responses,
-    and the dashed black line the obtained time-domain response. \deleted{The
-    errors are clipped for values smaller than 0.1\,\% and bigger than
-    100\,\%.}}
+    and the dashed black line the obtained time-domain response.}
   \label{fig:fullspace}
@@
 
-Figure~\ref{fig:marine} shows the result for an offset of 5\,km, \added{(a)} in
-\deleted{(a)} the frequency \added{domain} and (b) \added{in} the time domain.
-The recovered response with the 3D code captures the airwave (first peak) and
-the subsurface (second peak) very accurately. At later times the error starts
-to increase. We would need to compute a few additional lower frequencies if we
-want to improve it.
+Figure~\ref{fig:marine} shows the result for an offset of 5\,km, (a) in the
+frequency domain and (b) in the time domain. The recovered response with the 3D
+code captures the airwave (first peak) and the subsurface (second peak) very
+accurately. At later times the error starts to increase. We would need to
+compute a few additional lower frequencies if we want to improve it.
 %
@@
 is actually wrong. A possibility to detect such problems for complicated cases,
-where there is no semi-analytical result to compare with, is to \replaced{carry
-out a mesh-convergence test by computing}{compute} two or more models, moving
-the boundary. When the responses stop to change, one can assume that the
-boundary is far enough. Another possibility is to look at the amplitudes close
-to the boundaries and ensure that they are small enough.
-
-\subsection{\added{Induced Polarization}} % % % % % % % % % % % % % % % % % % %
-
-\added{The proposed Fourier-transform approach works independently of spatial
+where there is no semi-analytical result to compare with, is to carry out a
+mesh-convergence test by computing two or more models, moving the boundary.
+When the responses stop to change, one can assume that the boundary is far
+enough. Another possibility is to look at the amplitudes close to the
+boundaries and ensure that they are small enough.
+
+\subsection{Induced Polarization} % % % % % % % % % % % % % % % % % % % % % % %
+
+The proposed Fourier-transform approach works independently of spatial
 complexity, as diffusive electromagnetic fields are smooth functions of both
-frequency and time. An interesting test it to see whether this applies as
-well for dispersive media where the model parameters depend on frequency, as is
-the case in induced polarization. To test this we use the Cole-Cole model
-\mbox{\citep[CCM, ][]{JCP.41.Cole}} which, written in terms of conductivities
-instead of electric permittivities as by the original authors, is given by}
+frequency and time. An interesting test it to see whether this applies as well
+for dispersive media where the model parameters depend on frequency, as is the
+case in induced polarization. To test this we use the Cole-Cole model
+\citep[CCM, ][]{JCP.41.Cole} which, written in terms of conductivities instead
+of electric permittivities as by the original authors, is given by
 %
 \begin{equation}
-\added{
   \sigma(\omega) = \sigma_\infty + \frac{\sigma_0 - \sigma_\infty}{1 +
     (\rm{i}\omega\tau)^c}\ ,
-}
   \label{eq:ccm}
@@
 %
-\added{where $\sigma_0$ and $\sigma_\infty$ refer to the low-frequency and
+where $\sigma_0$ and $\sigma_\infty$ refer to the low-frequency and
 high-frequency conductivity values, respectively, $\tau$ is the central
@@
 relaxation time distribution. Note that this model slightly differs from the
-one phrased in terms of resistivities given by \mbox{\cite{GEO.78.Pelton}},
-see, e.g., \mbox{\cite{GJI.13.Tarasov}}. }
-
-\added{\mbox{\cite{GJI.17.Commer}} have shown that time-domain IP modelling
-can be difficult, particularly for low $c$-values. Modelling
-frequency-dependent models in the frequency domain is, for any given frequency,
-not different from modelling a non frequency-dependent model. However, the
-interesting point is to see how the Fourier transform behaves to obtain
-time-domain data from frequency-domain responses. As an example, we modelled
-the response for a land case with air above a halfspace of 1\,S/m. Within the
-halfspace is a dispersive, 100\,m thick layer at a depth of 300\,m, with
-$\sigma_0=1.0\,$S/m and $\sigma_\infty=1.25\,$S/m; $\tau=1$, and we vary $c$
-from 0 (no IP) to 1.0, 0.75, 0.5, and 0.25. The x-directed step-off source and
-the receiver at an inline offset of 900\,m are place at the surface. The real
-and imaginary frequency-domain responses for this model are shown in the left
-column of Figure~\ref{fig:CCM}. It can be seen that the effect of the IP layer
-in the frequency domain is that the real part starts to deviate from the DC
-value at lower frequencies. The imaginary part decays slower towards zero when
+one phrased in terms of resistivities given by \cite{GEO.78.Pelton}, see, e.g.,
+\cite{GJI.13.Tarasov}.
+
+\cite{GJI.17.Commer} have shown that time-domain IP modelling can be difficult,
+particularly for low $c$-values. Modelling frequency-dependent models in the
+frequency domain is, for any given frequency, not different from modelling a
+non frequency-dependent model. However, the interesting point is to see how the
+Fourier transform behaves to obtain time-domain data from frequency-domain
+responses. As an example, we modelled the response for a land case with air
+above a halfspace of 1\,S/m. Within the halfspace is a dispersive, 100\,m thick
+layer at a depth of 300\,m, with $\sigma_0=1.0\,$S/m and
+$\sigma_\infty=1.25\,$S/m; $\tau=1$, and we vary $c$ from 0 (no IP) to 1.0,
+0.75, 0.5, and 0.25. The x-directed step-off source and the receiver at an
+inline offset of 900\,m are place at the surface. The real and imaginary
+frequency-domain responses for this model are shown in the left column of
+Figure~\ref{fig:CCM}. It can be seen that the effect of the IP layer in the
+frequency domain is that the real part starts to deviate from the DC value at
+lower frequencies. The imaginary part decays slower towards zero when
 decreasing the frequency, and it deviates from a power-law function at lower
 frequencies as well. This indicates that we need lower frequency content for
-the Fourier transform the smaller $c$ becomes.}
-%
-\begin{figure}
+the Fourier transform the smaller $c$ becomes.
+%
+\begin{figure*}
   \centering
   \includegraphics[width=\fwidth]{11-cole-cole-model}
-  \caption{\added{IP example for different $c$ values, frequency-domain
-  response on the left and time-domain response on the right. The frequencies
-  indicated by the blue vertical bars in the lower-left figure were used for
-  our Fourier transform approach, requiring only the computation of 27
+  \caption{IP example for different $c$ values, frequency-domain response on
+    the left and time-domain response on the right. The frequencies indicated
+    by the blue vertical bars in the lower-left figure were used for our
+    Fourier transform approach, requiring only the computation of 27
     frequencies instead of the full 747 frequencies required by the digital
-linear filter method.}}
+    linear filter method.}
   \label{fig:CCM}
-\end{figure}
-%
-
-\added{For the Fourier transform, we used the 601-point sine-cosine filter from
-\mbox{\cite{GEO.09.Key}}. This is a very long filter, but it seemed to be by
-far the best filter for low $c$-values. However, designing a new filter
-particularly for IP problems might be a better approach. The selected frequency
-thresholds were chosen at $f_\mr{min}=10^{-4}\,$Hz and $f_\mr{max}=500\,$Hz,
-and within the thresholds we only computed every 6th frequency. Using these
-values only 27 frequencies have to be computed, instead of the 747 frequencies
-required by the actual transform. To obtain the step-off response we computed
-the step-on response and subtracted it from the DC value at 1000\,s (without
-the need for any additional frequencies).}
+\end{figure*}
+%
+
+For the Fourier transform, we used the 601-point sine-cosine filter from
+\cite{GEO.09.Key}. This is a very long filter, but it seemed to be by far the
+best filter for low $c$-values. However, designing a new filter particularly
+for IP problems might be a better approach. The selected frequency thresholds
+were chosen at $f_\mr{min}=10^{-4}\,$Hz and $f_\mr{max}=500\,$Hz, and within
+the thresholds we only computed every 6th frequency. Using these values only 27
+frequencies have to be computed, instead of the 747 frequencies required by the
+actual transform. To obtain the step-off response we computed the step-on
+response and subtracted it from the DC value at 1000\,s (without the need for
+any additional frequencies).
 
@@
 
-The \replaced{last}{final} example consists of a resistive, three-dimensional
-block embedded in the lower of two halfspaces, as depicted in
-Figure~\ref{fig:3d-model}. The target has resistivity
-$\rho_\mr{tg}=100\,\ohmm$, the upper halfspace corresponds to seawater with
-$\rho_\mr{sea}=0.3\,\ohmm$, and the lower halfspace is the background with
-$\rho_\mr{bg}=1\,\ohmm$.
+The last example consists of a resistive, three-dimensional block embedded in
+the lower of two halfspaces, as depicted in Figure~\ref{fig:3d-model}. The
+target has resistivity $\rho_\mr{tg}=100\,\ohmm$, the upper halfspace
+corresponds to seawater with $\rho_\mr{sea}=0.3\,\ohmm$, and the lower
+halfspace is the background with $\rho_\mr{bg}=1\,\ohmm$.
 %
@@
 
-For the comparison\added{,} we use the open-source code \simpeg
-\citep{CAG.15.Cockett}, which is a framework for modelling and inversion of
-geophysical data such as gravity, magnetics, and CSEM. It has Maxwell's
-equations implemented in both the frequency and time domain. As such we can
-compare our result computed in the frequency domain followed by a Fourier
-transform to a result computed directly in the time domain. A principal
-difference between \simpeg and \emg3d is that the former has various direct
-solvers implemented, whereas the latter is an iterative multigrid solver. The
-3D model is therefore a rather small example in order to be able to run it on
-our test machine, as the memory requirement by the direct solver would
-otherwise be too high. There are not many options out there of open-source
-time-domain 3D CSEM codes, \simpeg being the one we found to be suitable. A
-step-off response was chosen as this is the response currently implemented in
-it.
+For the comparison, we use the open-source code \simpeg \citep{CAG.15.Cockett},
+which is a framework for modelling and inversion of geophysical data such as
+gravity, magnetics, and CSEM. It has Maxwell's equations implemented in both
+the frequency and time domain. As such we can compare our result computed in
+the frequency domain followed by a Fourier transform to a result computed
+directly in the time domain. A principal difference between \simpeg and \emg3d
+is that the former has various direct solvers implemented, whereas the latter
+is an iterative multigrid solver. The 3D model is therefore a rather small
+example in order to be able to run it on our test machine, as the memory
+requirement by the direct solver would otherwise be too high. There are not
+many options out there of open-source time-domain 3D CSEM codes, \simpeg being
+the one we found to be suitable. A step-off response was chosen as this is the
+response currently implemented in it.
 
@@
     and in (b) the normalized difference (\%) between the two 3D codes.
-    \deleted{The results of these plots are clipped for values smaller than
-    0.1\,\% and bigger than 10\,\%.}}
+    }
   \label{fig:3d-result}
@@
 other domain followed by a transformation a viable alternative. The methodology
-is relatively simple to implement \added{and has therefore the potential to
-expand the capability} \replaced{of}{in} any \added{existing} code \added{to
-an additional dimension}.
+is relatively simple to implement and has therefore the potential to expand the
+capability of any existing code to an additional dimension.
 
@@
 \begin{acknowledgments} % % % % % % % % % % % % % % % % % % % % % % % % % % % %
-\added{We would like to thank the editor Ute Weckmann and assistant editor
-Louise Alexander as well as the reviewers Michael Commer and Chester Weiss for
-many helpful comments, which improved this manuscript considerably.}
-This research was conducted within the Gitaro.JIM project funded through
-MarTERA as part of Horizon 2020, a funding scheme of the European Research Area
-(ERA-NET Cofund, \href{https://www.martera.eu}{martera.eu}). An early idea of
-this manuscript was presented by \cite{EAGE.20.Werthmuller}.
+  We would like to thank the editor Ute Weckmann and assistant editor Louise
+  Alexander as well as the reviewers Michael Commer and Chester Weiss for many
+  helpful comments, which improved this manuscript considerably. This research
+  was conducted within the Gitaro.JIM project funded through MarTERA as part of
+  Horizon 2020, a funding scheme of the European Research Area (ERA-NET Cofund,
+  \href{https://www.martera.eu}{martera.eu}). An early idea of this manuscript
+  was presented by \cite{EAGE.20.Werthmuller}.
 \end{acknowledgments}

### 2021-04-23 ManuscriptTime.tex
Changes for Accepted version (#4)
- Add Zenodo DOI.
- Add png figures.

@@
     % referee,     % uncomment for double lines
-    final,       % uncomment for changes
+    % final,       % uncomment for changes
 ]{gji}
@@
 % Change default colour of «changes» package
-\definecolor{alizarin}{rgb}{0.82, 0.1, 0.26}
-\definecolor{mygreen}{rgb}{0, .5, 0}
-\setaddedmarkup{\color{mygreen}{\uline{#1}}}
-\setdeletedmarkup{\color{alizarin}{\uwave{#1}}}
+% \definecolor{alizarin}{rgb}{0.82, 0.1, 0.26}
+% \definecolor{mygreen}{rgb}{0, .5, 0}
+% \setaddedmarkup{\color{mygreen}{\uline{#1}}}
+% \setdeletedmarkup{\color{alizarin}{\uwave{#1}}}
 
@@
 \newcommand{\custem}{\texttt{custEM}\xspace}
+\newcommand{\ohmm}{\ensuremath{\Omega\,}\text{m}\xspace}
 
@@
 % \newcommand{\etodo}[1]{\todo[color=blue!40!white]{#1}}
-\newcommand{\dul}[1]{\underline{\underline{#1}}}
-\newcommand{\ul}[1]{\underline{#1}}
-\newcommand{\rmk}[1]{{\color{red}{#1}}}
-\newcommand{\ohmm}{\ensuremath{\Omega\,}\text{m}\xspace}
-\newcommand{\rtodo}{{\color{red}\texttt{TODO:}}\xspace}
+% \newcommand{\dul}[1]{\underline{\underline{#1}}}
+% \newcommand{\ul}[1]{\underline{#1}}
+% \newcommand{\rmk}[1]{{\color{red}{#1}}}
+% \newcommand{\rtodo}{{\color{red}\texttt{TODO:}}\xspace}
 
@@
 \volume{VVV}
-\pubyear{202Y}
+\pubyear{2021}
 
@@
 The data that support the findings of this study are openly available at Zenodo
-at \href{https://zenodo.org/badge/DOI/10.5281/zenodo.???????}{doi:
-10.5281/zenodo.???????}. The data includes the scripts and instructions to
-reproduce all results and figures.
-
-\emph{Note to editors and reviewers: The data are NOT YET on Zenodo, we will
-upload there the final versions and link the right DOI for the final
-manuscript. Until then everything can be found on GitHub at
-\href{https://github.com/empymod/article-TDEM}%
-{github.com/empymod/article-TDEM}.}
+at \href{https://doi.org/10.5281/zenodo.4710820}{doi: 10.5281/zenodo.4710820}.
+The data includes the scripts and instructions to reproduce all results and
+figures.
 

### 2021-05-25 ManuscriptTime.tex
Add link to original article

@@
 
-\title[Transforming expensive EM kernels]{Fast Fourier transformation
+\title[Transforming expensive EM kernels]{Fast Fourier transform
 of electromagnetic data for computationally expensive kernels}
@@
 %
-Three-dimensional controlled-source electromagnetic data is often computed
+Three-dimensional controlled-source electromagnetic data are often computed
 directly in the domain of interest, either in the frequency domain or in the
@@
 time-consuming as is the case in three-dimensional electromagnetic modelling.
-The speed of such modelling that is obtained through a transform is defined by
-three key points: solver, method and implementation of the Fourier transform,
-and gridding.
+The speed of modelling obtained through such a transform is defined by three
+key points: solver, method and implementation of the Fourier transform, and
+gridding.
 %
@@
 cells should be combined with optimising for computational speed. Looking
-carefully at these points results in much smaller computation times with
+carefully at these points resulted in much smaller computation times with
 speedup factors of ten or more over previous methods. A computation in one
@@
 transform if time-domain data are required \citep[e.g., ][]{GEO.15.Hunziker}.
-CSEM codes for 2D and 3D computations, on the other hand, often compute their
-responses directly in the required domain, either frequency or time.
+CSEM codes for arbitrary 2D and 3D computations, on the other hand, often
+compute their responses directly in the required domain, either frequency or
+time.
 
@@
 and Lanczos reduction schemes \cite[e.g., ][]{RS.94.Druskin} have a complexity
-of $O(n^4\sqrt{\log n})$, and Fourier transform of frequency-domain solutions
+of $O(n^4\sqrt{\log n})$, and Fourier transforms of frequency-domain solutions
 have a complexity $O(n_f n^3)$; $n$ is the number of grid points in each
@@
 demonstrate the efficiency of the approach with some numerical results. These
-compromise of a homogeneous space, a layered model, a model which includes
+comprise in of a homogeneous space, a layered model, a model which includes
 dispersive media (induced polarization), and a three-dimensional model.
@@
 the target of the following gridding recommendations. Whilst they will look
-differently for other mesh types or a time-domain code, some conclusion will
+differently for other mesh types or a time-domain code, some conclusions will
 still hold.
@@
 contain all source and receiver positions. The survey domain usually has no or
-very little cell stretching factor $\alpha_\mr{s}$. The minimum cell width is
+a very small cell stretching factor $\alpha_\mr{s}$. The minimum cell width is
 defined as
@@
 0.75, 0.5, and 0.25. The x-directed step-off source and the receiver at an
-inline offset of 900\,m are place at the surface. The real and imaginary
+inline offset of 900\,m are placed on the surface. The real and imaginary
 frequency-domain responses for this model are shown in the left column of
@@
 frequencies as well. This indicates that we need lower frequency content for
-the Fourier transform the smaller $c$ becomes.
+the Fourier transform, the smaller $c$ becomes.
 %
